### 2020-08-21 README.Rmd
markdownify; add citations; add figures

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Missing data in sports science: Current practice in football research and recommendations for reporting missingness

<!-- badges: start -->
<!-- badges: end -->

_David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)_

1: Griffith University, Menzies Health Institute Queensland, The Hopkins
Centre, Brisbane, Australia.

2: University of New South Wales, Department of Statistics, School of
Mathematics and Statistics, Sydney, Australia

3: Monash University, Department of Econometrics and Business Statistics,
Melbourne, Australia

4: Australian Centre of Excellence for Mathematical and Statistical
Frontiers (ACEMS), Melbourne, Australia

# Abstract

Missing data are often unavoidable. The reason values go missing, along
with decisions made in how missing data are handled (deleted or
imputed), can have a profound effect on the validity, usability and
accuracy of a study's results. In this article, we aimed to: estimate
the proportion of studies in football research that included a missing
data statement, highlight several improvable practices in relation to
missing data, and provide recommendations for exploring, visualising and
reporting missingness. Football related articles, published in 2019 were
studied. A survey of 137 articles, sampled at random, was conducted to
determine whether a missing data was included. As expected, the
proportion of studies in football research including a missing data
statement was low, at only X%; which suggests that this aspect of data
analysis is seldom considered by researchers. We suspect that this
result is consistent with the wider sports science literature. We
recommend researchers describe the number and percentage of missing
values, including when there are no missing values. Exploratory analysis
should be conducted to explore missing values, and visualisations
describing missingness overall should be provided in the paper or at
least supplementary materials. Missing values should almost always be
imputed, and imputation methods should be explored to ensure they are
appropriately representative.

# To recreate the paper

_instructions incoming_

### 2020-08-21 paper/supplementaries.Rmd
markdownify; add citations; add figures

**Supplement 1.** \[fix ordering\]

For example, if a players' player ratings\--a statistic derived from
many variables in the dataset\--are unavailable for one, or several,
Australian Football matches, despite the individual participating in the
match and the other variables being recorded.

MCAR refers to missingness that is unrelated to any observed or
unobserved data. Missing Australian Football player ratings could be
considered MCAR, if the missing ratings are not systematically related
to other variables. MAR is where missingness is related to some observed
data. Data MAR could arise if\... \^\^Rob's out on the full
example\--players who are injured are less likely to fill out wellness
surveys than those who are fully fit.\^\^. MNAR is when missingness is
related to something unobservable. For example, data MNAR could arise
when, if studying doping, a player was to remove their head and body
hair, to avoid providing a hair sample for testing
[\[4\]](https://www.zotero.org/google-docs/?4Ildjf). In this case, data
is missing (the hair sample), and this information is not captured by
any other variable that was collected (hair length was not recorded).

**Supplement 2. Systematic search**

The systematic search was performed in accordance with the reporting
requirements of the PRISMA statement (ref). The search strategy is
described below. The search protocol was not registered, and no funding
was provided to support the undertaking of the search.

**Objective**

This review aimed to estimate the proportion of articles published on
the topic area of football in 2019 that reported on missing data.

**Method**

Football related articles, published in 2019 were studied. Three major
sports medicine databases were searched using the key terms and limits,
on May 20, 2020. The databases were: SPORTDiscus, Embase and Cinahl. A
list of search terms and complete search statements are provided in
Table X.

Included articles were: (a) published in 2019; (b) had an accessible
English abstract; (c) had an accessible full text, in English; and (d)
included quantitative data. Studies containing original data, including
case studies (with time-series data) and n=1 trials, were considered.
The search results were imported into EndNote (version 9.3.2).
Duplicates were removed, and titles and abstracts were screened against
the inclusion criteria. Of the articles identified for full text
screening, 10% were sampled, at random, for survey. Because \~30% of the
papers from the initial random sample were not eligible for inclusion
(e.g., full-texts not in English, conference abstracts, qualitative
studies), an additional 5% random sample was taken\--with the
expectation that 30% of the articles from the additional sample would
also be ineligible\--so to preserve a 10% screen of the total papers
published on football related topics in 2019. Figure X summarises the
search process.

Articles were surveyed to determine: (a) whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; (b) where data were missing, whether data were considered
MCAR, MAR or MNAR; (c) whether imputation was used, and the imputation
method; and (d) whether articles that did not include a missing data
statement shared their data. Two authors independently extracted the
data. Agreement was X%; discrepancies were resolved by NJT. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution (ref),via the 'binom' package (ref) in R (version
3.5.1).

**Results**

The search returned a total of 3,863 articles (Figure X). Of the 199
articles screened, 136 were deemed eligible, representing \~10% of total
articles published on football related studies in 2019 (Figure X). The
proportion of articles that included a missing data statement (or
acknowledged missing data in a table/figure) was 5.9% \[95% confidence
interval; X% to X%\] or 8/136. A breakdown of the extracted information
is provided in Table X.

**Summary**

A survey of 136 articles published in 2019, on football related topics
showed that only one in 20 reported missing data. This result suggests
that missing data are rarely considered by sport science researchers.

**References**

\[Dave use plos one reference\] Moher D. Preferred Reporting Items for
Systematic Reviews and Meta-Analyses: The PRISMA Statement. Ann Intern
Med. 2009;151(4):264--269.
[[https://doi.org/10.7326/0003-4819-151-4-200908180-00135]{.ul}](https://doi.org/10.7326/0003-4819-151-4-200908180-00135).

Clopper CJ, Pearson ES. The use of confidence or fiducial limits
illustrated in the case of the binomial. Biometrika.
1934;26(4):404--413.
[[https://doi.org/10.1093/biomet/26.4.404]{.ul}](https://doi.org/10.1093/biomet/26.4.404)

Dorai-Raj S. binom: Binomial confidence intervals for several
parameterizations. R Package 'binom' version 1.1-1. 2014.
<https://cran.r-project.org/web/packages/binom/binom.pdf>

Search details

  Sportdiscus                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          Embase                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       Cinahl
  -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
  exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       (exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR \'performance analysis\' OR ((\'performance\'/exp OR performance) AND (\'analysis\'/exp OR analysis)) OR \'performance\'/exp OR performance OR \'sport psychology\'/exp OR \'sport psychology\' OR ((\'sport\'/exp OR sport) AND (\'psychology\'/exp OR psychology)) OR \'coaching\'/exp OR coaching) AND \[2019-2019\]/py                                                                                       exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching
  football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      (football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR \'australian rules footbal\' OR ((\'australian\'/exp OR australian) AND rules AND footbal) OR afl OR \'american football\'/exp OR \'american football\' OR ((\'american\'/exp OR american) AND (\'football\'/exp OR football)) OR \'gridiron football\' OR (gridiron AND (\'football\'/exp OR football)) OR \'touch football\' OR ((\'touch\'/exp OR touch) AND (\'football\'/exp OR football))) AND \[2019-2019\]/py   football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football
  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         \#1 AND \#2 AND \[english\]/lim AND 2019:py                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed
  Articles: 1,372                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      Articles: 1,756                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              Articles: 735
  https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=sph&bquery=(((football)+OR+(soccer)+OR+(indoor+soccer)+OR+(futsal)+OR+(rugby+league)+OR+(rugby+union)+OR+(gaelic+football)+OR+(australian+rules+football)+OR+(AFL)+OR+(american+football)+OR+(gridiron+football)+OR+(touch+football)))+AND+(((exercise+science)+OR+(sports+science)+OR+(physiology)+OR+(biomechanics)+OR+(nutrition)+OR+(training)+OR+(testing)+OR+(sports+medicine)+OR+(performance+analysis)+OR+(performance)+OR+(sport+psychology)+OR+(coaching)))&cli0=DT1&clv0=201901-201912&cli1=AA1&clv1=Y&cli2=RV&clv2=Y&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=c8h&bquery=(+exercise+science+OR+sports+science+OR+physiology+OR+biomechanics+OR+nutrition+OR+training+OR+testing+OR+sports+medicine+OR+performance+analysis+OR+performance+OR+sport+psychology+OR+coaching+)+AND+(+football+OR+soccer+OR+indoor+soccer+OR+futsal+OR+rugby+league+OR+rugby+union+OR+gaelic+football+OR+australian+rules+football+OR+AFL+OR+american+football+OR+gridiron+football+OR+touch+football+)&cli0=DT1&clv0=201901-201912&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site

Search results

+---------------------------------+-----------------------------------+
| Imported                        | 3,863                             |
+=================================+===================================+
| After duplicates removed        | 1,954                             |
+---------------------------------+-----------------------------------+
| After title and abstract screen | 1,373                             |
+---------------------------------+-----------------------------------+
| Random sample full-text         | 199 (needed for 10% sample)       |
+---------------------------------+-----------------------------------+
| Full-text included              | 137                               |
+---------------------------------+-----------------------------------+
| Full-text excluded              | Full-text papers excluded (n=62)  |
|                                 | with reasons:                     |
|                                 |                                   |
|                                 | Qualitative study (n=)            |
|                                 |                                   |
|                                 | Not in english (n=)               |
|                                 |                                   |
|                                 | Unable to access full-text (n=)   |
+---------------------------------+-----------------------------------+

**Figure 1.** PRISMA flow chart.

Rob's idea \[case study section\]:

-   Including a spectrum of worst practice - best practice

    -   Worst

        -   Missing data can be observed in how statistical methods are
            > reported

            -   e.g., missing meters gained

    -   less worse

        -   Missing data reported, but no handling stated

            -   e.g., 17% of data is missing, nothing mentioned about
                > deletion/imputation\
                > [[https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ]{.ul}](https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ)

            -   e.g., N = 18, and statistical analysis reports N = 14
                > (Robin, tweet, Dave to request paper)

    -   neutral?

        -   There is missing data, and this is reported?

    -   good

        -   Missing data reported, handling is reported, might be bad
            > imputation like mean imputation / data not shared.

    -   best

        -   Missing data reported, handling is reported, appropriate
            > imputation, data shared.

Appendix of techniques for dealing with missing values

-   scripts for exploring missingness

-   How to unpack the missingness in the paper text

```{=html}
<!-- -->
```
-   Ways to visualise missing data for reviewers, and readers of paper
    > to get a feel of how much to take away.

    -   So the meters gained example, I think a vis could be something
        > like the figure below. Also a way to visualise missing values
        > in athlete studies over time, if for whatever reason on
        > X1,x7,x35 day couldn't do a measurement on athlete 1.
        > \^\^instead of google sheet something ggplot\^\^

        -   vis_miss from visdat/naniar:

![](media/image2.png){width="6.267716535433071in"
height="0.8888888888888888in"}

\^\^Background in sports and exercise science\^\^

-   Missing data (or the absence of missing data) is often not reported

-   In general, imputation is not used

-   Missing data is never really discussed

-   Data are usually fit with ANOVA's, therefore values are subbed in,
    > e.g., mean. (really common practice)

**Notes/examples**

Example 1:
[[https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141]{.ul}](https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141)

Problem---The most important predictor, meters gained relative isn\'t
available in the vast majority of the dataset (2001-2008 for example)
(which is where the \'\'prediction accuracy\'\' was the worst. This
isn\'t mentioned anywhere in the paper, that their most important
predictor isn\'t available. So as a reviewer, unless data is provided,
or code shared. You are relying on domain knowledge to know this.

The supplementary materials which provides coefficients for the GLMS
fitted

![](media/image13.png){width="6.270833333333333in"
height="5.888888888888889in"}

No meters gained? This is weird right, talk about a circle, we have
meters gained most important, but GLM perform similar to decision tree
but the GLM doesn't have meters gained (the most important variable)
according to the decision trees?

Oh and meters gained isn't available throughout the dataset.

"GLMs were also created to predict Score Margin in each timeframe and
the prediction accuracies are similar for each time frame but are higher
than the decision tree prediction accuracies."

And whats with figure 2?

![](media/image15.png){width="6.270833333333333in" height="4.375in"}

The end buckets are supposed to be meters gained difference, but we have
a ?

Example 2:
[[https://twitter.com/championdata/status/472174368059957248]{.ul}](https://twitter.com/championdata/status/472174368059957248)

![](media/image5.png){width="3.7031255468066493in"
height="4.275152012248469in"}

![](media/image3.png){width="6.270833333333333in"
height="2.8333333333333335in"}

So if meters gained was first recorded in 2007, what does this mean for
the time frames being compared?

![](media/image11.png){width="6.270833333333333in"
height="2.2222222222222223in"}

Like OK that is 2009-2016, but what about the 2001-2016 was meters
gained only used for 2007-2016 so what was inputted 2001-2007 and if we
compared it to 2009-2016 where meters gained was FULLY available we have
a worse win/loss pred and a in the dataset with the MOST IMPORTANT
VARIABLE AVAILABLE ALL THE WAY THROUGH IT I am really confused.

**Modelling Match Outcome in Australian Football: Improved accuracy with
large databases **

(similar authors as above---Tran, Young etc)

![](media/image10.png){width="3.1875240594925636in"
height="0.7656889763779527in"}

Problem: The most important variable, in 2019 published without
mentioning its availability, nor are results presented in a way that
gives the reader an idea of what was done? Seriously your trying to
predict an outcome and literally the variable doesn\'t EXIST what was
entered in?

Data sharing/code sharing would have made this available to reviewers
and perhaps changed the tone of the papers.

Example 3:\
[[https://shapeamerica.tandfonline.com/doi/abs/10.1080/02640414.2015.1066026\#.XpkfaFMzZhE]{.ul}](https://protect-au.mimecast.com/s/FqW3CZYMqwSk7AkMsKELku?domain=shapeamerica.tandfonline.com)

These results on data available to public (basic stuff like counting of
kicks isn\'t reproducible)---i.e., the table here:

![](media/image9.png){width="3.557292213473316in"
height="2.951379046369204in"}

Means should be available?

Example 4:
[[https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full]{.ul}](https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full)

Problem:
[[https://twitter.com/RankingSw/status/961502281874092032]{.ul}](https://twitter.com/RankingSw/status/961502281874092032)

![A screenshot of a cell phone Description automatically
generated](media/image7.png){width="1.4457130358705161in"
height="2.085090769903762in"}

![](media/image14.png){width="1.4609733158355205in"
height="1.6510684601924759in"}

So why didn\'t the authors of the paper know this? Was it because the
underlying model (AFLPR) wasn\'t available at the time? etc But
certainly as a reviewer this is a thing someone should know before
reviewing?

Example 5:
[[https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511]{.ul}](https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511)

Figure below seems odd:

![](media/image4.png){width="2.2067705599300087in"
height="1.4527034120734907in"}

-   Progression of science, self-correction and the need for
    > 'self-correction to not be strike against your name'

-   Speak about the content in the papers wanting to push the knowledge,
    > not just get rushed for the purpose of a publication.

Recommendations -- Ask the question: what am I being provided here to
review?

-   Barriers

    -   Data privacy agreements (e.g., NDA), the reviewer would have to
        > sign one, maybe this is something that's not can occur, so
        > what's the minimum.

    -   Betting privacy, players union---you couldn't ever share who the
        > slowest player in the NRL same with the AFL

```{=html}
<!-- -->
```
-   Framed from the reviewer's point of view, needs to be a minimum
    > understanding of what is being done in the paper.

-   Exploratory and missingness (common problem in sports, maybe not
    > able to get access to the player)

```{=html}
<!-- -->
```
-   Minimum requirements: please give the summary stats, i.e., the
    > variable name, description, the dates collect, number of [missing
    > values]{.ul}, etc.

    -   Describe the data in some type of meaningful sense

    -   If there's definitions, over what period does the definition
        > hold for.

-   Idea scenario: Start data, end date.

Example from Twitter message

Dave sent a twitter message about this paper here.

[[https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490]{.ul}](https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490)

From the Paper

This study used commercially available data from the 2015 AFL season
supplied by Champion Data, the official statistics provider for the AFL.
An exemption from ethics review was granted by the institutional ethics
committee. The original data set included all 149,449 possession events
from all matches played during the 2015 AFL season (18 teams, 665
players, 206 regular season and finals series games). A data subset was
created to focus the analysis on regular season matches in order to
explore wholeof-league patterns (i.e., finals matches were omitted as
they are only played by teams in the top 8 end-of-season ladder
positions). This subset was filtered further to include only those
sequences of possession events that i) started with a kick-in following
a behind, and ii) ended with the offensive team losing possession
through a score, stoppage, or turnover. These sequences are referred to
herein as "kick-in possession chains", where a sequence begins with the
player who first receives the ball from a kick-in and excludes the
initial disposal out of the goal square, unless the player starting with
possession "plays on" to himself. Possession chains that involved only
one player or were incomplete due to end-of-quarter time restrictions
were excluded. Chains from 2 drawn matches were also excluded. The final
data set used for analysis comprised a total of 1,720 kick-in possession
chains from 194 games.

Now thankfully we can use fitzRoy to check out how many behinds there
were in 2015 as per the subset.

> fitzRoy::get_afltables_stats(start_date=\"2015-01-01\",
> end_date=\"2016-01-01\")%\>% \#Season chosen
>
> select(Season,Home.score, Away.score, Round, Home.team, Away.team,
> Round, AQ4B, HQ4B)%\>%\# variables needed
>
> filter(!Round %in% c(\"EF\", \"SF\",\"QF\",\"PF\", \"GF\"))%\>%
> \#excluding finals
>
> filter(Home.score != Away.score)%\>% \#excluding draws
>
> distinct()%\>%
>
> mutate(totalbehinds=AQ4B+HQ4B)%\>% \#away behinds + home behinds
>
> summarise(totalbehindsin2015=sum(totalbehinds))
>
> totalbehindsin2015
>
> \<int\>
>
> 1 4275
>
> \>
>
> So if each quarter ended with a invalid observation (didn't result in
> a score, turnover etc) this would be pretty high I'd imagine but help
> illustrate the point
>
> 4\*194
>
> So we have 4275 kick ins from behinds, of which at MOST 194\*4 (776)
> would be excluded say if every quarter had a kick after the siren that
> resulted in a behind.
>
> So that leaves us with 4275-776 total kick ins which is 3499 kick in
> ins total
>
> **So this data suggests that only 1720/3499 or 0.491569 of kick ins
> resulted in a chain? in the study.**
>
> This seems not good.
>
> To put another way,
>
> out of the total behinds or ''opportunities for chains to start'' if
> we take the exclusion rule. **" Possession chains that involved only
> one player or were incomplete due to end-of-quarter time restrictions
> were excluded."**
>
> **Then**
>
> **(4275-1720 )/ 4275 or 0.5976608 of chains are EXCLUDED that seems
> massive!**

![](media/image12.png){width="6.270833333333333in"
height="1.8194444444444444in"}

New Paper: Recommending how to share data in sports science

-   Simplest approach

    -   If you report the summary statistics for each variable

    -   Then recreate the data from that

        -   The min-max and sample evenly

        -   the distribution

### 2020-08-21 R/read-gs.R

@@ -1,0 +1,7 @@
+library(googlesheets4)
+# insert your own email address
+gs4_auth("nicholas.tierney@monash.edu") 
+sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
+gs_football <- read_sheet(sheet_url, skip = 1)
+
+gs_football

### 2020-08-21 R/read-markdown.R

@@ -1,0 +1,2 @@
+library(markdrive)
+gdoc_checkout("Missing data in sports and exercise science research")

### 2020-08-21 paper/figures/create-table.R

@@ -1,0 +1,45 @@
+library(magick)
+library(tibble)
+library(gt)
+library(tidyr)
+
+player_table_implicit <- tibble::tribble(
+  ~Player,    ~Quarter,   ~Goals,
+  "Coleman",   1,         0,
+  "Coleman",   2,         1,
+  "Coleman",   3,         2,
+  "Coleman",   4,         0,
+  "Koenen",    1,         1,
+  "Koenen",    3,         2
+)
+
+player_table_explicit <- player_table_implicit %>% 
+  pivot_wider(id_cols = Player,
+              names_from = Quarter,
+              values_from = Goals)
+
+gt(player_table_implicit) %>% 
+  gtsave("paper/figures/implicit.png")
+
+gt(player_table_explicit) %>% 
+  tab_style(
+    style = cell_fill(color = "lightblue"),
+    locations = cells_body(
+      columns = vars(`2`,`4`),
+      rows = is.na(`2`) | is.na(`4`)
+    )
+  ) %>% 
+  gtsave("paper/figures/explicit.png")
+
+img_implicit <- image_read("paper/figures/implicit.png")
+img_explicit <- image_read("paper/figures/explicit.png")
+img_arrow <- image_read("paper/figures/arrow.png")
+
+img_tables <- 
+image_append(c(img_implicit, 
+               {image_scale(img_arrow, "10%") %>% 
+                   image_border("white", geometry = "50x220x50x50")
+                 },
+               image_border(img_explicit, "white", geometry = "0x142+0+0")))
+
+image_write(img_tables, "paper/figures/tables.png")

### 2020-08-24 paper/paper.Rmd

---
title: "Missing data in sports science: Current practice in football research and recommendations for reporting missingness"
authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: references.bib
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE)
```

```{r libraries, message = FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)
```

Affiliations:

1: Griffith University, Menzies Health Institute Queensland, The Hopkins
Centre, Brisbane, Australia.

2: University of New South Wales, Department of Statistics, School of
Mathematics and Statistics, Sydney, Australia

3: Monash University, Department of Econometrics and Business Statistics,
Melbourne, Australia

4: Australian Centre of Excellence for Mathematical and Statistical
Frontiers (ACEMS), Melbourne, Australia

**Acknowledgement/conflict of interest**

NJT naniar package?

**Corresponding author**

Nicholas J Tierney

Monash University, Department of Econometrics and Business Statistics,
Melbourne, Australia

**Data availability**

# Abstract {-}

Missing data are often unavoidable. The reason values go missing, along
with decisions made in how missing data are handled (deleted or
imputed), can have a profound effect on the validity, usability and
accuracy of a study's results. In this article, we aimed to: estimate
the proportion of studies in football research that included a missing
data statement, highlight several improvable practices in relation to
missing data, and provide recommendations for exploring, visualising and
reporting missingness. Football related articles, published in 2019 were
studied. A survey of 137 articles, sampled at random, was conducted to
determine whether a missing data was included. As expected, the
proportion of studies in football research including a missing data
statement was low, at only X%; which suggests that this aspect of data
analysis is seldom considered by researchers. We suspect that this
result is consistent with the wider sports science literature. We
recommend researchers describe the number and percentage of missing
values, including when there are no missing values. Exploratory analysis
should be conducted to explore missing values, and visualisations
describing missingness overall should be provided in the paper or at
least supplementary materials. Missing values should almost always be
imputed, and imputation methods should be explored to ensure they are
appropriately representative.

**Keywords:** Imputation, missingness, missings, naniar, sport, football

# Introduction {#intro}

Missing data can undermine the validity of research results
[@Sainani2015; @Nakagawa2008; Sterne2009]. Data can go
missing many ways. For example, accidentally skipping a survey question,
equipment failure, or intentionally not recording values. Before
analysis, missing data must be handled, with values typically deleted or
imputed. There is no universal approach to handling missing data with
contextual factors -- such as, the study design and objective, and
pattern of missingness -- determining how to handle missings on a
case-by-case basis [@Sainani2015].
The effect of missing data on the validity and accuracy of a study's
results is largely determined by the reasons for missingness, and how
missing data are handled.

Data does not go missing the same way every time. There are three broad
categorisations describing why data can be missing: Missing Completely
at Random (MCAR), Missing at Random (MAR), and Missing Not at Random
(MNAR). These categorisations help describe whether the missingness
occurs by chance (MCAR), due to some observed variable (MAR) or an
unobserved variable (MNAR), with MCAR being benign and the latter two
biased. Football relevant examples are provided in Supplement X, and
more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
[@Sainani2015; @Nakagawa2008; Sterne2009].

The three categorisations of missingness types provide a useful
framework to describe bias arising from missing data. If you strongly
suspect data is MCAR, you can be more certain your data isn't biased,
otherwise there may be bias in your sample. This does not mean data MAR
or MNAR make data invalid, instead it identifies potential biases, which
can help improve future research design. However, these categorisations
do not necessarily guide the user in their subsequent actions in
analysis, or describe specifically how values go missing. To identify
possible mechanisms for missingness, the data must be explored. It can
be challenging to identify the missingness mechanism in data, but it is
critical, since missing values could change the outcome of a study. This
could mean the work cannot be reproduced, and worse, may lead scientists
and practitioners to the wrong outcome.

It is imperative that missing data (or lack thereof -- Absence of
evidence is not evidence of absence) are reported. However, outside of
sports science, researchers seldom consider missing data [@Schafer2002].
Moreover, while there has been significant interest in modelling missing
data [@refs], the exploration and reporting of missing data have received
less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
of articles that report missing data in football research; (2) highlight
several bad missing data practices; and (3) provide recommendations for
exploring and reporting missingness.

# Methods and materials

To estimate the proportion of articles that report missing data in
football research, we conducted a systematic search [@Schafer2002]. Football related
articles published in 2019 were studied. Three major sports medicine
databases (SPORTDiscus, Embase and Cinahl) were searched, using key
terms and search limits (Supplement 1). Included articles were: (a)
published in 2019; (b) written in English; (c) had an accessible
full-text; and (d) included quantitative data. The search process is
summarised in Figure 1.

```{r prisma-example, eval = FALSE}
require(PRISMAstatement)

prisma()

[[https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html]{.ul}](https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html)

library(PRISMAstatement)

e.g.,

prisma(found = x,

found_other = x,

no_dupes = x,

screened = x,

screen_exclusions = x,

full_text = x,

full_text_exclusions = x,

qualitative = x,

quantitative = x,

width = 800, height = 800)
```

+---------------------------------+-----------------------------------+
| Imported                        | 3,863                             |
+=================================+===================================+
| After duplicates removed        | 1,954                             |
+---------------------------------+-----------------------------------+
| After title and abstract screen | 1,373                             |
+---------------------------------+-----------------------------------+
| Random sample full-text         | 199 (needed for 10% sample)       |
+---------------------------------+-----------------------------------+
| Full-text included              | 137                               |
+---------------------------------+-----------------------------------+
| Full-text excluded              | Full-text papers excluded (n=62)  |
|                                 | with reasons:                     |
|                                 |                                   |
|                                 | Qualitative study (n=)            |
|                                 |                                   |
|                                 | Not in english (n=)               |
|                                 |                                   |
|                                 | Unable to access full-text (n=)   |
+---------------------------------+-----------------------------------+

**Figure 1.** PRISMA flow chart.

Our intention was to survey, at random, 10% of the total articles found
(after title and abstract screening). We were primarily interested in
whether missing data were reported or acknowledged; and if not, whether
a dataset was shared. Where applicable, information about how missing
values were handled was also extracted (see Supplement 1). Two authors
independently extracted the data (X% agreement). Findings are reported
as the proportion and 95% confidence interval. Supplement 1 provides
further details of the search.

# Results

Of the 199 articles screened, 136 were deemed eligible, representing
\~10% of the articles found (Figure X). XX articles (of the 199) were
excluded due to not meeting the inclusion criteria. The proportion of
articles that reported missing data was 5.9% (95% confidence interval
X% to X%) or 8/136. Of the articles that did not mention missing data
(128/136), 7.1% (3.3% to 13.1%) or 9/128 shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration -- especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).

# Discussion

Our survey of 136 articles published on football related topics found
that only about one in 20 papers reported missing data. It is reasonable
to assume that this result is representative of current practices in the
wider sports science literature. This finding may be largely explained
by a lack of awareness and education on missing data practices (Kristin
BJSM paper), similar to other areas of sciences (refs-education and
ecology examples)(2). Below we discuss several poor practices in
relation to missing data, and provide recommendations for exploring,
visualising and reporting missingness.

## Missing data across the analysis pathway

Problem: *Unreported missing values*. While not always the case, even if
not reported, missing data can be obvious. A recent paper examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016. Problematically, at least
one of the 91 team performance indicators included in the analysis was
not available over the entire 2001-2016 period [@Young2019]. The variable
'meters gained' was not collected over the specified duration [@ref].
This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Unfortunately, the absence of a
missing data statement when data are missing is not unique to the study
above [@Young2019]. Authors should include a missing data statement,
irrespective of whether data is missing, or not.

Problem: *Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. One imputation method is imputing the mean value
(e.g., participant or group average) [@Young2019]. While imputing the mean
preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002].
Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data -- such as linear regression, or K nearest neighbours -- should
be considered. This is discussed in the recommendations section.

Problem: *Not evaluating the effect, or choice, of imputation.* The
imputation method used to impute missing data has the potential to
affect the outcome of a study. Therefore, it is important that authors
understand, and document, how imputation affected the goal of the
analysis (i.e., inference, prediction, or both). For example,
understanding the implications in using mean imputation, compared to
K-nearest neighbour imputation -- the effect on parameter estimates
('significant' versus 'not significant'), and the (un)certainty of the
coefficients (i.e., less/more). When using imputation, authors need to
perform a sensitivity analysis (ref circulation article-Nick). This is
discussed below.

## Common causes of missing values

Broadly, there are two types of missing values in data: implicit and
explicit missings. Explicit values are missing, but recorded; whereas
with implicit missinges, their presence is implied based on other
information in the data. For example, in  Table \@ref(tab:implicit-missings), player Koenen has
missing values for quarters two and four. Sometimes values like these
can be logically imputed, as it might be known that these values are
recorded as 0 in this format, rather than NA. Other places missing data
can arise include: bad joins when merging data without corresponding
values, surveys, an inability to collect a biological sample (e.g.,
venous blood), equipment malfunction or failure, or or not being worn.

```{r implicit-missings, fig.cap = "Two tables demonstrating explicit and implicit missing. The first table shows the number of goals scored for a player in a given quarter of an AFL match with the first column showing the player name, the second the quarter they played, and the third the goals they scored. Note that Player, 'Koenen' has no entries for Quarter 2 and 4. The second table shows the same information from the first table pivoted, with each row being a player and the number of goals they scored in each quarter, with each quarter being a column. We notice that in the second form of the data, we can clearly see that Koenen has missing values. These types of missing values have a name, implicit missing values. The first table has implicit missing values, meaning they are implied, and the second table has those implicit missing values explicitly expressed."}

knitr::include_graphics(here::here("paper/figures/tables.png"))
```

## Visualising missing values

We recommend that researchers use missing data overview graphics (see
Graphics section in [@Tierney2018], For example, overview plots [@Tierney2017].

(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data.

```{r missing-overview, fig.cap = "Overviews of missing values in airquality data. Panel A shows... Panel B shows ... Panel C shows ... (simulate data from football study, provide write up in the supplementary materials)", fig.height = 10}

aq_vis_miss <- vis_miss(airquality)

aq_upset <- gg_miss_upset(airquality)

aq_fct <- gg_miss_fct(airquality, Month)

aq_miss_var <- gg_miss_var(airquality) + labs(title = "A")
aq_miss_case <- gg_miss_case(airquality) + labs(title = "B")
aq_vis_miss <- vis_miss(airquality) + labs(title = "C")

(aq_miss_var | aq_miss_case) / (aq_vis_miss) / (aq_fct)

# can

```

To learn more about exploring missing values, we recommend the vignettes
in naniar [@naniar], and the methods in (@Tierney2018)

## Understanding imputation

Imputing data might feel wrong, as we are 'making up data'. The truth
is, we (generally) can never know what the missing values were. The goal
of imputing data is to make the best possible inference from the data.
We recommend data is imputed, with a few caveats. Consider removing
variables with a high proportion of missing data (e.g., variables with
the majority or more missing than not). With respect to a 'rule of
thumb' for when data should be imputed, we deliberately do not suggest
any specific threshold (e.g., impute values when >5% of data are
missing otherwise delete rows with <5% missing) and caution against
such practice. Rather, we recommend values should generally be imputed..
We recommend avoiding imputation methods that impute the same value,
such as mean or median imputation, and instead suggest using imputation
methods such as: linear regression, k nearest neighbours, or expectation
maximisation.

Much in the way of there is no single 'best' statistical method, there
is no perfect, one-size-fits-all approach for imputing data. The goal is
to generate similar values that might have been otherwise recorded -
sometimes this means using a neighborhood approach of finding similar
values, other times this means predicting responses using linear or tree
models, and other times the most likely value might have been 0 or the
last or first value carried forward. For detailed descriptions and a
summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].

The imputation methods discussed so far impute a single value for each
missing value, and have the eponymous name, "single imputation".
Multiple imputation is a method where multiple values are imputed for
each missing value, creating "m" datasets, which are then specially
combined during analysis. Multiple imputation is generally seen as the
best method to get the most reasonable inference from the data, as it
reflects the uncertainty in the missing values. For more information on
using multiple imputation, we recommend Stefan van Buuren's book, @Van_Buuren2012.

No matter the imputation method used, it is essential to compare results
of different missing data handling to understand how they may bias the
results. For example, comparing analysis results from applying listwise
deletion, compared to mean imputation, compared to linear regression
imputation. This can reveal bias occurring in imputation methods. An
example of this approach is described in the Case Study in
@Tierney2018.

# Recommendations

The current state of reporting missing data in the sports literature is
poor. In this section we discuss recommended practices for exploring and
reporting missing data. Our intention is not to provide a 'cookbook'
style approach to missing data, but rather broad recommendations to help
researchers when writing the methods and results sections of a study,
and assist researchers when evaluating a study during the peer-review
process.

In the methods section of a study, we recommend the following points are
addressed:

1.  Describe screening procedures (define or example -- think about screened out?).

2.  State if any observations were dropped (if any).

3.  Provide the number and percentage of observations dropped.

4.  State if any potential bias was incurred as a result of the screening.

5.  If screening is complex, consider providing a workflow diagram explaining how observations were kept or dropped.

In the results section, we recommend addressing the following points:

1.  State the number and percentage of observations missing.

> Examples: "*There were no missing values*" or "*25 of 280 values were missing (8.92%)*"

2.  State if missing data was explored. If there is missingness, provide a summary graphic (see figure XX)

> Examples: "*Missing data was explored, revealing a relationship between missingness and age*", "*Missing data was not explored, as the reasons and impacts were known*"

3.  State any reasons known or unknown for missing values and if bias have occurred as a result. For example, a study on injuries, removing players who get injured during the study seems potentially problematic.

> Examples: "*Values were missing due to faults in GPS tracking*", "*Values were missing in a survey as participants did not complete the section*"

4.  State actions taken in handling missing data.

> Were they removed? Imputed? Another approach?

5.  Describe and justify any imputation process used.

6.  Explore how imputation or lack of imputation impacts results.

> Missing values of ZZ were imputed with a linear regression using terms XX.
> Missing values were imputed with the mean values

An example writeup of missing data for a results section is provided
below:

> *25 of the 280 values in the dataset (~8.92%) were missing. These were due to faults in GPS tracking as participants ran through sections of track covered by forest, a known issue with some GPS tracking. These speed values were interpolated using a nearest neighbours approach, taking inputs of speed, and altitude. The track was imputed using the known track in the area for those sections.*

# Conclusion

More often than not, researchers will be faced with missing data. The
reasons for the missingness, along with the decision of how missing data
are treated, can have a profound effect on the validity and accuracy of
a study's results. A survey of 2019 articles on football related topics
showed that current practice is poor. Only about one in 20 studies
reported missing data. To help address this issue, we have discussed
recommendations for exploring and reporting missing data, to assist
researchers during this important process in the analytical pipeline.
Research should consider these recommendations, and pay greater
attention to missing data and its influence on research results.

# References {-}

### 2020-08-28 paper/paper.Rmd
remove a space
add supplement headings

@@
     toc_float: true
+    keep_md: true
 bibliography: references.bib
@@
 **Data availability**
+
+**Word count**
+
+```{r count-words}
+wordcountaddin::text_stats()
+```
 
@@
 
-It is imperative that missing data (or lack thereof -- Absence of
+It is imperative that missing data (or lack thereof --Absence of
 evidence is not evidence of absence) are reported. However, outside of
@@
 # References {-}
+
+# Supplement 1
+
+# Supplement 2
+
+# Supplement 3

### 2020-08-28 R/paper-statement-analysis.R
tidies up R folder

@@ -1,28 +1,24 @@
-library(tidyverse)
-#library(ids)
-
-setwd("~/Dropbox/Research projects/Editorial - Missing data in sport and exercise science research")
-d = read.csv("missing-data-endnote-library.csv")
-
-# repeatable analysis
-set.seed(2020-05-21)
-articles <- d %>% sample_frac(size = 0.1)
-#write.csv(articles, file = "random-sample-137.csv")
-
-articles$sampled <- 1
-data <- merge(d, articles, all.x = T)
-data$sampled <- data$sampled %>% replace_na(0)
-d_second_sample <- filter(data, sampled < 1)
-articles_second_sample <- d_second_sample %>% sample_frac(size = 0.05)
-#write.csv(articles_second_sample, file = "random-sample-62.csv")
-
-
 library(binom)
-library(dplyr)
+library(here)
+library(tidyverse)
+library(conflicted)
+library(googlesheets4)
+conflicted::conflict_prefer("select", "dplyr")
+conflicted::conflict_prefer("filter", "dplyr")
+
+# insert your own email address
+gs4_auth("nicholas.tierney@monash.edu") 
+sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
+gs_football <- read_sheet(sheet_url, skip = 1)
+
+gs_football
+
 
 #### Calculate proportions
-d = read.csv("missing-data-in-football-preliminary.csv")
-d = filter(d, Study_eligibility == "1")
+d <- read_csv(here("data",
+                  "missing-data-in-football-preliminary.csv"))
+
+d <- filter(d, Study_eligibility == "1")
 
 # Reporting missing
 table(d$Any_missing_data_statement, useNA = 'ifany')

### 2020-08-28 R/read-markdown.R

@@ -1,2 +1,3 @@
 library(markdrive)
+
 gdoc_checkout("Missing data in sports and exercise science research")

### 2020-08-28 R/sampling-papers.R

@@ -1,0 +1,25 @@
+# Code to sample the papers for review
+# possible papers to review after title abstract screening are in file
+# "missing-data-endnote-library.csv"
+# We then sampled possible papers to get to 10% of the number of screened papers
+# Two files were written, "random-sample-137", and "random-sample-62". These two
+# were created to get close to 10% of the number of papers in total 199 papers
+# were reviewed.
+
+library(tidyverse)
+library(here)
+d <- read_csv(here("data",
+                   "missing-data-endnote-library.csv"))
+
+# repeatable analysis
+set.seed(2020-05-21)
+articles <- d %>% sample_frac(size = 0.1)
+#write.csv(articles, file = "random-sample-137.csv")
+
+articles$sampled <- 1
+data <- merge(d, articles, all.x = T)
+data$sampled <- data$sampled %>% replace_na(0)
+d_second_sample <- filter(data, sampled < 1)
+articles_second_sample <- d_second_sample %>% sample_frac(size = 0.05)
+#write.csv(articles_second_sample, file = "random-sample-62.csv")
+

### 2020-08-28 paper/paper.Rmd
updates references and changes "problem" to "case study"

@@
 Missing data can undermine the validity of research results
-[@Sainani2015; @Nakagawa2008; Sterne2009]. Data can go
+[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
 missing many ways. For example, accidentally skipping a survey question,
@@
 
-Problem: *Unreported missing values*. While not always the case, even if
+Case study: *Unreported missing values*. While not always the case, even if
 not reported, missing data can be obvious. A recent paper examined the
@@
 
-Problem: *Mean imputation.* The default of most statistical software for
+Case study: *Mean imputation.* The default of most statistical software for
 handling missing data in modelling is listwise deletion. Listwise
@@
 
-Problem: *Not evaluating the effect, or choice, of imputation.* The
+Case study: *Not evaluating the effect, or choice, of imputation.* The
 imputation method used to impute missing data has the potential to
@@
 # References {-}
-
-# Supplement 1
-
-# Supplement 2
-
-# Supplement 3

### 2020-08-28 paper/supplement-1.Rmd
first pass at supplementary materials for #7

--- 2020-08-21 paper/supplementaries.Rmd
@@
-**Supplement 1.** \[fix ordering\]
+---
+title: Supplement 1
+output: html_document
+---
+
+# missing data examples relevant to sports (MAR, MCAR, MNAR)
 
@@
 any other variable that was collected (hair length was not recorded).
-
-**Supplement 2. Systematic search**
-
-The systematic search was performed in accordance with the reporting
-requirements of the PRISMA statement (ref). The search strategy is
-described below. The search protocol was not registered, and no funding
-was provided to support the undertaking of the search.
-
-**Objective**
-
-This review aimed to estimate the proportion of articles published on
-the topic area of football in 2019 that reported on missing data.
-
-**Method**
-
-Football related articles, published in 2019 were studied. Three major
-sports medicine databases were searched using the key terms and limits,
-on May 20, 2020. The databases were: SPORTDiscus, Embase and Cinahl. A
-list of search terms and complete search statements are provided in
-Table X.
-
-Included articles were: (a) published in 2019; (b) had an accessible
-English abstract; (c) had an accessible full text, in English; and (d)
-included quantitative data. Studies containing original data, including
-case studies (with time-series data) and n=1 trials, were considered.
-The search results were imported into EndNote (version 9.3.2).
-Duplicates were removed, and titles and abstracts were screened against
-the inclusion criteria. Of the articles identified for full text
-screening, 10% were sampled, at random, for survey. Because \~30% of the
-papers from the initial random sample were not eligible for inclusion
-(e.g., full-texts not in English, conference abstracts, qualitative
-studies), an additional 5% random sample was taken\--with the
-expectation that 30% of the articles from the additional sample would
-also be ineligible\--so to preserve a 10% screen of the total papers
-published on football related topics in 2019. Figure X summarises the
-search process.
-
-Articles were surveyed to determine: (a) whether a missing data
-statement was included, or if missing data were acknowledged in a
-figure/table; (b) where data were missing, whether data were considered
-MCAR, MAR or MNAR; (c) whether imputation was used, and the imputation
-method; and (d) whether articles that did not include a missing data
-statement shared their data. Two authors independently extracted the
-data. Agreement was X%; discrepancies were resolved by NJT. Findings are
-reported as the proportion and 95% confidence interval (CI). Confidence
-intervals were calculated using the Clopper-Pearson method for the
-binomial distribution (ref),via the 'binom' package (ref) in R (version
-3.5.1).
-
-**Results**
-
-The search returned a total of 3,863 articles (Figure X). Of the 199
-articles screened, 136 were deemed eligible, representing \~10% of total
-articles published on football related studies in 2019 (Figure X). The
-proportion of articles that included a missing data statement (or
-acknowledged missing data in a table/figure) was 5.9% \[95% confidence
-interval; X% to X%\] or 8/136. A breakdown of the extracted information
-is provided in Table X.
-
-**Summary**
-
-A survey of 136 articles published in 2019, on football related topics
-showed that only one in 20 reported missing data. This result suggests
-that missing data are rarely considered by sport science researchers.
-
-**References**
-
-\[Dave use plos one reference\] Moher D. Preferred Reporting Items for
-Systematic Reviews and Meta-Analyses: The PRISMA Statement. Ann Intern
-Med. 2009;151(4):264--269.
-[[https://doi.org/10.7326/0003-4819-151-4-200908180-00135]{.ul}](https://doi.org/10.7326/0003-4819-151-4-200908180-00135).
-
-Clopper CJ, Pearson ES. The use of confidence or fiducial limits
-illustrated in the case of the binomial. Biometrika.
-1934;26(4):404--413.
-[[https://doi.org/10.1093/biomet/26.4.404]{.ul}](https://doi.org/10.1093/biomet/26.4.404)
-
-Dorai-Raj S. binom: Binomial confidence intervals for several
-parameterizations. R Package 'binom' version 1.1-1. 2014.
-<https://cran.r-project.org/web/packages/binom/binom.pdf>
-
-Search details
-
-  Sportdiscus                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          Embase                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       Cinahl
-  -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-  exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       (exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR \'performance analysis\' OR ((\'performance\'/exp OR performance) AND (\'analysis\'/exp OR analysis)) OR \'performance\'/exp OR performance OR \'sport psychology\'/exp OR \'sport psychology\' OR ((\'sport\'/exp OR sport) AND (\'psychology\'/exp OR psychology)) OR \'coaching\'/exp OR coaching) AND \[2019-2019\]/py                                                                                       exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching
-  football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      (football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR \'australian rules footbal\' OR ((\'australian\'/exp OR australian) AND rules AND footbal) OR afl OR \'american football\'/exp OR \'american football\' OR ((\'american\'/exp OR american) AND (\'football\'/exp OR football)) OR \'gridiron football\' OR (gridiron AND (\'football\'/exp OR football)) OR \'touch football\' OR ((\'touch\'/exp OR touch) AND (\'football\'/exp OR football))) AND \[2019-2019\]/py   football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football
-  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         \#1 AND \#2 AND \[english\]/lim AND 2019:py                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed
-  Articles: 1,372                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      Articles: 1,756                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              Articles: 735
-  https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=sph&bquery=(((football)+OR+(soccer)+OR+(indoor+soccer)+OR+(futsal)+OR+(rugby+league)+OR+(rugby+union)+OR+(gaelic+football)+OR+(australian+rules+football)+OR+(AFL)+OR+(american+football)+OR+(gridiron+football)+OR+(touch+football)))+AND+(((exercise+science)+OR+(sports+science)+OR+(physiology)+OR+(biomechanics)+OR+(nutrition)+OR+(training)+OR+(testing)+OR+(sports+medicine)+OR+(performance+analysis)+OR+(performance)+OR+(sport+psychology)+OR+(coaching)))&cli0=DT1&clv0=201901-201912&cli1=AA1&clv1=Y&cli2=RV&clv2=Y&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=c8h&bquery=(+exercise+science+OR+sports+science+OR+physiology+OR+biomechanics+OR+nutrition+OR+training+OR+testing+OR+sports+medicine+OR+performance+analysis+OR+performance+OR+sport+psychology+OR+coaching+)+AND+(+football+OR+soccer+OR+indoor+soccer+OR+futsal+OR+rugby+league+OR+rugby+union+OR+gaelic+football+OR+australian+rules+football+OR+AFL+OR+american+football+OR+gridiron+football+OR+touch+football+)&cli0=DT1&clv0=201901-201912&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site
-
-Search results
-
-+---------------------------------+-----------------------------------+
-| Imported                        | 3,863                             |
-+=================================+===================================+
-| After duplicates removed        | 1,954                             |
-+---------------------------------+-----------------------------------+
-| After title and abstract screen | 1,373                             |
-+---------------------------------+-----------------------------------+
-| Random sample full-text         | 199 (needed for 10% sample)       |
-+---------------------------------+-----------------------------------+
-| Full-text included              | 137                               |
-+---------------------------------+-----------------------------------+
-| Full-text excluded              | Full-text papers excluded (n=62)  |
-|                                 | with reasons:                     |
-|                                 |                                   |
-|                                 | Qualitative study (n=)            |
-|                                 |                                   |
-|                                 | Not in english (n=)               |
-|                                 |                                   |
-|                                 | Unable to access full-text (n=)   |
-+---------------------------------+-----------------------------------+
-
-**Figure 1.** PRISMA flow chart.
-
-Rob's idea \[case study section\]:
-
--   Including a spectrum of worst practice - best practice
-
-    -   Worst
-
-        -   Missing data can be observed in how statistical methods are
-            > reported
-
-            -   e.g., missing meters gained
-
-    -   less worse
-
-        -   Missing data reported, but no handling stated
-
-            -   e.g., 17% of data is missing, nothing mentioned about
-                > deletion/imputation\
-                > [[https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ]{.ul}](https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ)
-
-            -   e.g., N = 18, and statistical analysis reports N = 14
-                > (Robin, tweet, Dave to request paper)
-
-    -   neutral?
-
-        -   There is missing data, and this is reported?
-
-    -   good
-
-        -   Missing data reported, handling is reported, might be bad
-            > imputation like mean imputation / data not shared.
-
-    -   best
-
-        -   Missing data reported, handling is reported, appropriate
-            > imputation, data shared.
-
-Appendix of techniques for dealing with missing values
-
--   scripts for exploring missingness
-
--   How to unpack the missingness in the paper text
-
-```{=html}
-<!-- -->
-```
--   Ways to visualise missing data for reviewers, and readers of paper
-    > to get a feel of how much to take away.
-
-    -   So the meters gained example, I think a vis could be something
-        > like the figure below. Also a way to visualise missing values
-        > in athlete studies over time, if for whatever reason on
-        > X1,x7,x35 day couldn't do a measurement on athlete 1.
-        > \^\^instead of google sheet something ggplot\^\^
-
-        -   vis_miss from visdat/naniar:
-
-![](media/image2.png){width="6.267716535433071in"
-height="0.8888888888888888in"}
-
-\^\^Background in sports and exercise science\^\^
-
--   Missing data (or the absence of missing data) is often not reported
-
--   In general, imputation is not used
-
--   Missing data is never really discussed
-
--   Data are usually fit with ANOVA's, therefore values are subbed in,
-    > e.g., mean. (really common practice)
-
-**Notes/examples**
-
-Example 1:
-[[https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141]{.ul}](https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141)
-
-Problem---The most important predictor, meters gained relative isn\'t
-available in the vast majority of the dataset (2001-2008 for example)
-(which is where the \'\'prediction accuracy\'\' was the worst. This
-isn\'t mentioned anywhere in the paper, that their most important
-predictor isn\'t available. So as a reviewer, unless data is provided,
-or code shared. You are relying on domain knowledge to know this.
-
-The supplementary materials which provides coefficients for the GLMS
-fitted
-
-![](media/image13.png){width="6.270833333333333in"
-height="5.888888888888889in"}
-
-No meters gained? This is weird right, talk about a circle, we have
-meters gained most important, but GLM perform similar to decision tree
-but the GLM doesn't have meters gained (the most important variable)
-according to the decision trees?
-
-Oh and meters gained isn't available throughout the dataset.
-
-"GLMs were also created to predict Score Margin in each timeframe and
-the prediction accuracies are similar for each time frame but are higher
-than the decision tree prediction accuracies."
-
-And whats with figure 2?
-
-![](media/image15.png){width="6.270833333333333in" height="4.375in"}
-
-The end buckets are supposed to be meters gained difference, but we have
-a ?
-
-Example 2:
-[[https://twitter.com/championdata/status/472174368059957248]{.ul}](https://twitter.com/championdata/status/472174368059957248)
-
-![](media/image5.png){width="3.7031255468066493in"
-height="4.275152012248469in"}
-
-![](media/image3.png){width="6.270833333333333in"
-height="2.8333333333333335in"}
-
-So if meters gained was first recorded in 2007, what does this mean for
-the time frames being compared?
-
-![](media/image11.png){width="6.270833333333333in"
-height="2.2222222222222223in"}
-
-Like OK that is 2009-2016, but what about the 2001-2016 was meters
-gained only used for 2007-2016 so what was inputted 2001-2007 and if we
-compared it to 2009-2016 where meters gained was FULLY available we have
-a worse win/loss pred and a in the dataset with the MOST IMPORTANT
-VARIABLE AVAILABLE ALL THE WAY THROUGH IT I am really confused.
-
-**Modelling Match Outcome in Australian Football: Improved accuracy with
-large databases **
-
-(similar authors as above---Tran, Young etc)
-
-![](media/image10.png){width="3.1875240594925636in"
-height="0.7656889763779527in"}
-
-Problem: The most important variable, in 2019 published without
-mentioning its availability, nor are results presented in a way that
-gives the reader an idea of what was done? Seriously your trying to
-predict an outcome and literally the variable doesn\'t EXIST what was
-entered in?
-
-Data sharing/code sharing would have made this available to reviewers
-and perhaps changed the tone of the papers.
-
-Example 3:\
-[[https://shapeamerica.tandfonline.com/doi/abs/10.1080/02640414.2015.1066026\#.XpkfaFMzZhE]{.ul}](https://protect-au.mimecast.com/s/FqW3CZYMqwSk7AkMsKELku?domain=shapeamerica.tandfonline.com)
-
-These results on data available to public (basic stuff like counting of
-kicks isn\'t reproducible)---i.e., the table here:
-
-![](media/image9.png){width="3.557292213473316in"
-height="2.951379046369204in"}
-
-Means should be available?
-
-Example 4:
-[[https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full]{.ul}](https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full)
-
-Problem:
-[[https://twitter.com/RankingSw/status/961502281874092032]{.ul}](https://twitter.com/RankingSw/status/961502281874092032)
-
-![A screenshot of a cell phone Description automatically
-generated](media/image7.png){width="1.4457130358705161in"
-height="2.085090769903762in"}
-
-![](media/image14.png){width="1.4609733158355205in"
-height="1.6510684601924759in"}
-
-So why didn\'t the authors of the paper know this? Was it because the
-underlying model (AFLPR) wasn\'t available at the time? etc But
-certainly as a reviewer this is a thing someone should know before
-reviewing?
-
-Example 5:
-[[https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511]{.ul}](https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511)
-
-Figure below seems odd:
-
-![](media/image4.png){width="2.2067705599300087in"
-height="1.4527034120734907in"}
-
--   Progression of science, self-correction and the need for
-    > 'self-correction to not be strike against your name'
-
--   Speak about the content in the papers wanting to push the knowledge,
-    > not just get rushed for the purpose of a publication.
-
-Recommendations -- Ask the question: what am I being provided here to
-review?
-
--   Barriers
-
-    -   Data privacy agreements (e.g., NDA), the reviewer would have to
-        > sign one, maybe this is something that's not can occur, so
-        > what's the minimum.
-
-    -   Betting privacy, players union---you couldn't ever share who the
-        > slowest player in the NRL same with the AFL
-
-```{=html}
-<!-- -->
-```
--   Framed from the reviewer's point of view, needs to be a minimum
-    > understanding of what is being done in the paper.
-
--   Exploratory and missingness (common problem in sports, maybe not
-    > able to get access to the player)
-
-```{=html}
-<!-- -->
-```
--   Minimum requirements: please give the summary stats, i.e., the
-    > variable name, description, the dates collect, number of [missing
-    > values]{.ul}, etc.
-
-    -   Describe the data in some type of meaningful sense
-
-    -   If there's definitions, over what period does the definition
-        > hold for.
-
--   Idea scenario: Start data, end date.
-
-Example from Twitter message
-
-Dave sent a twitter message about this paper here.
-
-[[https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490]{.ul}](https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490)
-
-From the Paper
-
-This study used commercially available data from the 2015 AFL season
-supplied by Champion Data, the official statistics provider for the AFL.
-An exemption from ethics review was granted by the institutional ethics
-committee. The original data set included all 149,449 possession events
-from all matches played during the 2015 AFL season (18 teams, 665
-players, 206 regular season and finals series games). A data subset was
-created to focus the analysis on regular season matches in order to
-explore wholeof-league patterns (i.e., finals matches were omitted as
-they are only played by teams in the top 8 end-of-season ladder
-positions). This subset was filtered further to include only those
-sequences of possession events that i) started with a kick-in following
-a behind, and ii) ended with the offensive team losing possession
-through a score, stoppage, or turnover. These sequences are referred to
-herein as "kick-in possession chains", where a sequence begins with the
-player who first receives the ball from a kick-in and excludes the
-initial disposal out of the goal square, unless the player starting with
-possession "plays on" to himself. Possession chains that involved only
-one player or were incomplete due to end-of-quarter time restrictions
-were excluded. Chains from 2 drawn matches were also excluded. The final
-data set used for analysis comprised a total of 1,720 kick-in possession
-chains from 194 games.
-
-Now thankfully we can use fitzRoy to check out how many behinds there
-were in 2015 as per the subset.
-
-> fitzRoy::get_afltables_stats(start_date=\"2015-01-01\",
-> end_date=\"2016-01-01\")%\>% \#Season chosen
->
-> select(Season,Home.score, Away.score, Round, Home.team, Away.team,
-> Round, AQ4B, HQ4B)%\>%\# variables needed
->
-> filter(!Round %in% c(\"EF\", \"SF\",\"QF\",\"PF\", \"GF\"))%\>%
-> \#excluding finals
->
-> filter(Home.score != Away.score)%\>% \#excluding draws
->
-> distinct()%\>%
->
-> mutate(totalbehinds=AQ4B+HQ4B)%\>% \#away behinds + home behinds
->
-> summarise(totalbehindsin2015=sum(totalbehinds))
->
-> totalbehindsin2015
->
-> \<int\>
->
-> 1 4275
->
-> \>
->
-> So if each quarter ended with a invalid observation (didn't result in
-> a score, turnover etc) this would be pretty high I'd imagine but help
-> illustrate the point
->
-> 4\*194
->
-> So we have 4275 kick ins from behinds, of which at MOST 194\*4 (776)
-> would be excluded say if every quarter had a kick after the siren that
-> resulted in a behind.
->
-> So that leaves us with 4275-776 total kick ins which is 3499 kick in
-> ins total
->
-> **So this data suggests that only 1720/3499 or 0.491569 of kick ins
-> resulted in a chain? in the study.**
->
-> This seems not good.
->
-> To put another way,
->
-> out of the total behinds or ''opportunities for chains to start'' if
-> we take the exclusion rule. **" Possession chains that involved only
-> one player or were incomplete due to end-of-quarter time restrictions
-> were excluded."**
->
-> **Then**
->
-> **(4275-1720 )/ 4275 or 0.5976608 of chains are EXCLUDED that seems
-> massive!**
-
-![](media/image12.png){width="6.270833333333333in"
-height="1.8194444444444444in"}
-
-New Paper: Recommending how to share data in sports science
-
--   Simplest approach
-
-    -   If you report the summary statistics for each variable
-
-    -   Then recreate the data from that
-
-        -   The min-max and sample evenly
-
-        -   the distribution

### 2020-08-28 paper/supplement-2.Rmd
first pass at supplementary materials for #7

---
title: "Supplement 2"
output: 
  bookdown::html_document2
bibliography: "supp-2.bib"
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE)
```

```{r libraries, message = FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(conflicted)
library(gt)
library(PRISMAstatement)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

# Systematic search details

The systematic search was performed in accordance with the reporting
requirements of the PRISMA statement [@Moher2009]. The search strategy is
described below. The search protocol was not registered, and no funding
was provided to support the undertaking of the search.

# Objective

This review aimed to estimate the proportion of articles published on
the topic area of football in 2019 that reported on missing data.

# Method

Football related articles, published in 2019 were studied. Three major
sports medicine databases were searched using the key terms and limits,
on May 20, 2020. The databases were: SPORTDiscus, Embase and Cinahl. A
list of search terms and complete search statements are provided in
Table \@ref(tab:search-table).

```{r search-table}
search_table <- tibble(
  Sportdiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football",
    "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed",
    "Articles: 1,372",
    "https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=sph&bquery=(((football)+OR+(soccer)+OR+(indoor+soccer)+OR+(futsal)+OR+(rugby+league)+OR+(rugby+union)+OR+(gaelic+football)+OR+(australian+rules+football)+OR+(AFL)+OR+(american+football)+OR+(gridiron+football)+OR+(touch+football)))+AND+(((exercise+science)+OR+(sports+science)+OR+(physiology)+OR+(biomechanics)+OR+(nutrition)+OR+(training)+OR+(testing)+OR+(sports+medicine)+OR+(performance+analysis)+OR+(performance)+OR+(sport+psychology)+OR+(coaching)))&cli0=DT1&clv0=201901-201912&cli1=AA1&clv1=Y&cli2=RV&clv2=Y&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site"),
  Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py",
             "#1 AND #2 AND [english]/lim AND 2019:py",
             "Articles: 1,756",
             ""),
  Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football",
              "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed",
              "Articles: 735",
              "https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=c8h&bquery=(+exercise+science+OR+sports+science+OR+physiology+OR+biomechanics+OR+nutrition+OR+training+OR+testing+OR+sports+medicine+OR+performance+analysis+OR+performance+OR+sport+psychology+OR+coaching+)+AND+(+football+OR+soccer+OR+indoor+soccer+OR+futsal+OR+rugby+league+OR+rugby+union+OR+gaelic+football+OR+australian+rules+football+OR+AFL+OR+american+football+OR+gridiron+football+OR+touch+football+)&cli0=DT1&clv0=201901-201912&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site")

)

search_table %>% 
  gt() %>% 
  cols_width(
    vars(Sportdiscus, Cinahl) ~ px(240),
    vars(Embase) ~ px(300)
    )
```

Included articles were: (a) published in 2019; (b) had an accessible
English abstract; (c) had an accessible full text, in English; and (d)
included quantitative data. Studies containing original data, including
case studies (with time-series data) and n=1 trials, were considered.
The search results were imported into EndNote (version 9.3.2).
Duplicates were removed, and titles and abstracts were screened against
the inclusion criteria. Of the articles identified for full text
screening, 10% were sampled, at random, for survey. Because \~30% of the
papers from the initial random sample were not eligible for inclusion
(e.g., full-texts not in English, conference abstracts, qualitative
studies), an additional 5% random sample was taken\--with the
expectation that 30% of the articles from the additional sample would
also be ineligible\--so to preserve a 10% screen of the total papers
published on football related topics in 2019. Figure \@ref(fig:prisma) summarises the search process.

```{r prisma}
  # | Imported                        | 3,863                             |
  # | After duplicates removed        | 1,954                             |
  # | After title and abstract screen | 1,373                             |
  # | Random sample full-text         | 199 (needed for 10% sample)       |
  # | Full-text included              | 137                               |
  # | Full-text excluded              | Full-text papers excluded (n=62)  |
  # |                                 | with reasons:                     |
  # |                                 | Qualitative study (n=)            |
  # |                                 | Not in english (n=)               |
  # |                                 | Unable to access full-text (n=)   |

prisma(found = 3863,
       found_other = 0,
       no_dupes = 1954, 
       screened = 1373, # (1954 - 1373) 
       screen_exclusions = 581, 
       full_text = 199,
       full_text_exclusions = 62, 
       qualitative = 0, 
       quantitative = 0,
       width = 800, 
       height = 800)
```

Articles were surveyed to determine: (a) whether a missing data
statement was included, or if missing data were acknowledged in a
figure/table; (b) where data were missing, whether data were considered
MCAR, MAR or MNAR; (c) whether imputation was used, and the imputation
method; and (d) whether articles that did not include a missing data
statement shared their data. Two authors independently extracted the
data. Agreement was X%; discrepancies were resolved by NJT. Findings are
reported as the proportion and 95% confidence interval (CI). Confidence
intervals were calculated using the Clopper-Pearson method for the
binomial distribution [@Clopper1934],via the 'binom' package [@binom] in R (version 4.0.2) [@rcore].

# Results

The search returned a total of 3,863 articles (Figure X). Of the 199
articles screened, 136 were deemed eligible, representing \~10% of total
articles published on football related studies in 2019 (Figure X). The
proportion of articles that included a missing data statement (or
acknowledged missing data in a table/figure) was 5.9% \[95% confidence
interval; X% to X%\] or 8/136. A breakdown of the extracted information
is provided in Table XX.

```{r extracted-info}

extractions <- tibble(
  Imported = 3863,
  `After duplicates removed` = 1954,
  `After title and abstract screen` = 1373,
  `Random sample full-text` = "199 for 10% sample",
  `Full-text included` = 137,
  `Full-text excluded` = 62
)

```

# Summary

A survey of 136 articles published in 2019, on football related topics
showed that only one in 20 reported missing data. This result suggests
that missing data are rarely considered by sport science researchers.

# References {-}

### 2020-08-28 paper/supplement-3.Rmd
first pass at supplementary materials for #7

--- 2020-08-21 paper/supplementaries.Rmd
@@
-**Supplement 1.** \[fix ordering\]
+---
+title: "Supplement 3"
+output: html_document
+---
 
-For example, if a players' player ratings\--a statistic derived from
-many variables in the dataset\--are unavailable for one, or several,
-Australian Football matches, despite the individual participating in the
-match and the other variables being recorded.
-
-MCAR refers to missingness that is unrelated to any observed or
-unobserved data. Missing Australian Football player ratings could be
-considered MCAR, if the missing ratings are not systematically related
-to other variables. MAR is where missingness is related to some observed
-data. Data MAR could arise if\... \^\^Rob's out on the full
-example\--players who are injured are less likely to fill out wellness
-surveys than those who are fully fit.\^\^. MNAR is when missingness is
-related to something unobservable. For example, data MNAR could arise
-when, if studying doping, a player was to remove their head and body
-hair, to avoid providing a hair sample for testing
-[\[4\]](https://www.zotero.org/google-docs/?4Ildjf). In this case, data
-is missing (the hair sample), and this information is not captured by
-any other variable that was collected (hair length was not recorded).
-
-**Supplement 2. Systematic search**
-
-The systematic search was performed in accordance with the reporting
-requirements of the PRISMA statement (ref). The search strategy is
-described below. The search protocol was not registered, and no funding
-was provided to support the undertaking of the search.
-
-**Objective**
-
-This review aimed to estimate the proportion of articles published on
-the topic area of football in 2019 that reported on missing data.
-
-**Method**
-
-Football related articles, published in 2019 were studied. Three major
-sports medicine databases were searched using the key terms and limits,
-on May 20, 2020. The databases were: SPORTDiscus, Embase and Cinahl. A
-list of search terms and complete search statements are provided in
-Table X.
-
-Included articles were: (a) published in 2019; (b) had an accessible
-English abstract; (c) had an accessible full text, in English; and (d)
-included quantitative data. Studies containing original data, including
-case studies (with time-series data) and n=1 trials, were considered.
-The search results were imported into EndNote (version 9.3.2).
-Duplicates were removed, and titles and abstracts were screened against
-the inclusion criteria. Of the articles identified for full text
-screening, 10% were sampled, at random, for survey. Because \~30% of the
-papers from the initial random sample were not eligible for inclusion
-(e.g., full-texts not in English, conference abstracts, qualitative
-studies), an additional 5% random sample was taken\--with the
-expectation that 30% of the articles from the additional sample would
-also be ineligible\--so to preserve a 10% screen of the total papers
-published on football related topics in 2019. Figure X summarises the
-search process.
-
-Articles were surveyed to determine: (a) whether a missing data
-statement was included, or if missing data were acknowledged in a
-figure/table; (b) where data were missing, whether data were considered
-MCAR, MAR or MNAR; (c) whether imputation was used, and the imputation
-method; and (d) whether articles that did not include a missing data
-statement shared their data. Two authors independently extracted the
-data. Agreement was X%; discrepancies were resolved by NJT. Findings are
-reported as the proportion and 95% confidence interval (CI). Confidence
-intervals were calculated using the Clopper-Pearson method for the
-binomial distribution (ref),via the 'binom' package (ref) in R (version
-3.5.1).
-
-**Results**
-
-The search returned a total of 3,863 articles (Figure X). Of the 199
-articles screened, 136 were deemed eligible, representing \~10% of total
-articles published on football related studies in 2019 (Figure X). The
-proportion of articles that included a missing data statement (or
-acknowledged missing data in a table/figure) was 5.9% \[95% confidence
-interval; X% to X%\] or 8/136. A breakdown of the extracted information
-is provided in Table X.
-
-**Summary**
-
-A survey of 136 articles published in 2019, on football related topics
-showed that only one in 20 reported missing data. This result suggests
-that missing data are rarely considered by sport science researchers.
-
-**References**
-
-\[Dave use plos one reference\] Moher D. Preferred Reporting Items for
-Systematic Reviews and Meta-Analyses: The PRISMA Statement. Ann Intern
-Med. 2009;151(4):264--269.
-[[https://doi.org/10.7326/0003-4819-151-4-200908180-00135]{.ul}](https://doi.org/10.7326/0003-4819-151-4-200908180-00135).
-
-Clopper CJ, Pearson ES. The use of confidence or fiducial limits
-illustrated in the case of the binomial. Biometrika.
-1934;26(4):404--413.
-[[https://doi.org/10.1093/biomet/26.4.404]{.ul}](https://doi.org/10.1093/biomet/26.4.404)
-
-Dorai-Raj S. binom: Binomial confidence intervals for several
-parameterizations. R Package 'binom' version 1.1-1. 2014.
-<https://cran.r-project.org/web/packages/binom/binom.pdf>
-
-Search details
-
-  Sportdiscus                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          Embase                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       Cinahl
-  -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ -------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-  exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       (exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR \'performance analysis\' OR ((\'performance\'/exp OR performance) AND (\'analysis\'/exp OR analysis)) OR \'performance\'/exp OR performance OR \'sport psychology\'/exp OR \'sport psychology\' OR ((\'sport\'/exp OR sport) AND (\'psychology\'/exp OR psychology)) OR \'coaching\'/exp OR coaching) AND \[2019-2019\]/py                                                                                       exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching
-  football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      (football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR \'australian rules footbal\' OR ((\'australian\'/exp OR australian) AND rules AND footbal) OR afl OR \'american football\'/exp OR \'american football\' OR ((\'american\'/exp OR american) AND (\'football\'/exp OR football)) OR \'gridiron football\' OR (gridiron AND (\'football\'/exp OR football)) OR \'touch football\' OR ((\'touch\'/exp OR touch) AND (\'football\'/exp OR football))) AND \[2019-2019\]/py   football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football
-  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         \#1 AND \#2 AND \[english\]/lim AND 2019:py                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed
-  Articles: 1,372                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      Articles: 1,756                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              Articles: 735
-  https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=sph&bquery=(((football)+OR+(soccer)+OR+(indoor+soccer)+OR+(futsal)+OR+(rugby+league)+OR+(rugby+union)+OR+(gaelic+football)+OR+(australian+rules+football)+OR+(AFL)+OR+(american+football)+OR+(gridiron+football)+OR+(touch+football)))+AND+(((exercise+science)+OR+(sports+science)+OR+(physiology)+OR+(biomechanics)+OR+(nutrition)+OR+(training)+OR+(testing)+OR+(sports+medicine)+OR+(performance+analysis)+OR+(performance)+OR+(sport+psychology)+OR+(coaching)))&cli0=DT1&clv0=201901-201912&cli1=AA1&clv1=Y&cli2=RV&clv2=Y&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=c8h&bquery=(+exercise+science+OR+sports+science+OR+physiology+OR+biomechanics+OR+nutrition+OR+training+OR+testing+OR+sports+medicine+OR+performance+analysis+OR+performance+OR+sport+psychology+OR+coaching+)+AND+(+football+OR+soccer+OR+indoor+soccer+OR+futsal+OR+rugby+league+OR+rugby+union+OR+gaelic+football+OR+australian+rules+football+OR+AFL+OR+american+football+OR+gridiron+football+OR+touch+football+)&cli0=DT1&clv0=201901-201912&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site
-
-Search results
-
-+---------------------------------+-----------------------------------+
-| Imported                        | 3,863                             |
-+=================================+===================================+
-| After duplicates removed        | 1,954                             |
-+---------------------------------+-----------------------------------+
-| After title and abstract screen | 1,373                             |
-+---------------------------------+-----------------------------------+
-| Random sample full-text         | 199 (needed for 10% sample)       |
-+---------------------------------+-----------------------------------+
-| Full-text included              | 137                               |
-+---------------------------------+-----------------------------------+
-| Full-text excluded              | Full-text papers excluded (n=62)  |
-|                                 | with reasons:                     |
-|                                 |                                   |
-|                                 | Qualitative study (n=)            |
-|                                 |                                   |
-|                                 | Not in english (n=)               |
-|                                 |                                   |
-|                                 | Unable to access full-text (n=)   |
-+---------------------------------+-----------------------------------+
-
-**Figure 1.** PRISMA flow chart.
+(Content pulled from the google doc)
 

### 2020-09-01 paper/paper.Rmd
updated abstract

@@
 studied. A survey of 137 articles, sampled at random, was conducted to
-determine whether a missing data was included. As expected, the
+determine whether a missing data statement was included. As expected, the
 proportion of studies in football research including a missing data
-statement was low, at only X%; which suggests that this aspect of data
-analysis is seldom considered by researchers. We suspect that this
-result is consistent with the wider sports science literature. We
-recommend researchers describe the number and percentage of missing
+statement was low, at only X% [95% confidence interval; X% to X%];
+suggesting that this aspect of data analysis is seldom considered by researchers.
+We suspect that this result is consistent with the wider sports science literature.
+We recommend researchers describe the number and percentage of missing
 values, including when there are no missing values. Exploratory analysis

### 2020-09-01 paper/paper.Rmd
title change

@@
 ---
-title: "Missing data in sports science: Current practice in football research and recommendations for reporting missingness"
+title: "Missing data: Current practice in football research and recommendations for reporting missingness"
 authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)

### 2020-09-02 paper/paper.Rmd
change abstract wording

@@
 Missing data are often unavoidable. The reason values go missing, along
-with decisions made in how missing data are handled (deleted or
+with decisions made of how missing data are handled (deleted or
 imputed), can have a profound effect on the validity, usability and
@@
 determine whether a missing data statement was included. As expected, the
-proportion of studies in football research including a missing data
-statement was low, at only X% [95% confidence interval; X% to X%];
-suggesting that this aspect of data analysis is seldom considered by researchers.
-We suspect that this result is consistent with the wider sports science literature.
-We recommend researchers describe the number and percentage of missing
+proportion of studies in football research that included a missing data
+statement was low, at only X% [95% confidence interval; X% to X%]; suggesting that 
+missingness is seldom considered by researchers. We suspect that this result 
+is consistent with the wider sports science literature. We recommend 
+researchers describe the number and percentage of missing
 values, including when there are no missing values. Exploratory analysis

### 2020-09-02 paper/paper.Rmd
edit wording first paragraph of the introduction

@@
 should be conducted to explore missing values, and visualisations
-describing missingness overall should be provided in the paper or at
+describing missingness overall should be provided in the paper, or at
 least supplementary materials. Missing values should almost always be
@@
 
-Missing data can undermine the validity of research results
+Missing data can undermine the validity and accuracy of research results
 [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
@@
 analysis, missing data must be handled, with values typically deleted or
-imputed. There is no universal approach to handling missing data with
-contextual factors -- such as, the study design and objective, and
-pattern of missingness -- determining how to handle missings on a
+imputed. There is no universal approach to handling missing data. 
+Contextual factors--such as, the study design and objective, and
+pattern of missingness--largely determine how to handle missings on a
 case-by-case basis [@Sainani2015].

### 2020-09-02 paper/paper.Rmd
edit wording in intro

@@
 Contextual factors--such as, the study design and objective, and
-pattern of missingness--largely determine how to handle missings on a
+pattern of missingness--determine how to handle missings on a
 case-by-case basis [@Sainani2015].

### 2020-09-02 paper/paper.Rmd
first pass at addressing #3

@@
 
-Missing data can undermine the validity and accuracy of research results
+Missing data are values that should have been observed, but were not. Since the
+values are unobserved, this can mean the population of interest might not be 
+properly sampled, inducing bias into study results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results
 [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
@@
 missing data are handled.
+
+For a value to be missing, it should have been observed in the data, but was not.
+For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event.
+This "goal" is now a missing value.
+Compare this to a game being rained out and no game taking place.
+There is no data to be observed, so there is no missing data.
 

### 2020-09-02 paper/paper.Rmd
some more expansion on MCAR and friends

@@
 
-For a value to be missing, it should have been observed in the data, but was not.
+A missing value should have been observed or measured in the data, but was not.
 For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event.
@@
 Compare this to a game being rained out and no game taking place.
-There is no data to be observed, so there is no missing data.
+There is no missing data because there is no data to be observed.
+A final example - a study screens out individuals who do not meet criteria for the study.
+These are not missing values, because there was no intent for these to be observed or measured.
 
@@
 occurs by chance (MCAR), due to some observed variable (MAR) or an
-unobserved variable (MNAR), with MCAR being benign and the latter two
-biased. Football relevant examples are provided in Supplement X, and
+unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR
+biased [@Newman2014]. Football relevant examples are provided in Supplement X, and
 more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
@@
 framework to describe bias arising from missing data. If you strongly
-suspect data is MCAR, you can be more certain your data isn't biased,
-otherwise there may be bias in your sample. This does not mean data MAR
-or MNAR make data invalid, instead it identifies potential biases, which
-can help improve future research design. However, these categorisations
+suspect data is MCAR, your results are less likely to be biased. If you suspect
+MAR or MNAR, there is bias in your results. This does not mean data MAR
+or MNAR make results invalid, instead it identifies potential bias, which can be
+used to help improve future research design. However, these categorisations
 do not necessarily guide the user in their subsequent actions in
@@
 evidence is not evidence of absence) are reported. However, outside of
-sports science, researchers seldom consider missing data [@Schafer2002].
+sports science, researchers seldom consider missing data [@Schafer2002] (<< NOTE - I'm not sure we mean that?).
 Moreover, while there has been significant interest in modelling missing

### 2020-09-02 paper/paper.Rmd
change intro wording

--- 2020-09-02 paper/paper.Rmd
@@
 
-**Keywords:** Imputation, missingness, missings, naniar, sport, football
+**Keywords:** Exercise, imputation, missingness, missings, naniar, sport
 
@@
 
-Missing data can undermine the validity and accuracy of research results
+Missing data can undermine the validity of research results
 [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
@@
 Contextual factors--such as, the study design and objective, and
-pattern of missingness--determine how to handle missings on a
+pattern of missingness--determine how missings should be handled, on a
 case-by-case basis [@Sainani2015].

### 2020-09-02 paper/paper.Rmd
update intro wording

@@
 unobserved variable (MNAR), with MCAR being benign and the latter two
-biased. Football relevant examples are provided in Supplement X, and
+biased. Football relevant examples are provided in Supplement 1, and
 more detailed explanations of MCAR, MAR, MNAR can be found elsewhere

### 2020-09-02 paper/paper.Rmd
change wording in the intro

@@
 The three categorisations of missingness types provide a useful
-framework to describe bias arising from missing data. If you strongly
-suspect data is MCAR, you can be more certain your data isn't biased,
-otherwise there may be bias in your sample. This does not mean data MAR
+framework to describe bias arising from missing data. If one strongly
+suspects data is MCAR, one can be more certain your data isn't biased,
+otherwise there may be bias in your sample. This does not mean that data MAR
 or MNAR make data invalid, instead it identifies potential biases, which
-can help improve future research design. However, these categorisations
-do not necessarily guide the user in their subsequent actions in
+can help improve future research design **(Nick, can we look at this sentence please)**. However, these categorisations
+do not necessarily guide a researcher in their subsequent actions in
 analysis, or describe specifically how values go missing. To identify

### 2020-09-02 paper/paper.Rmd
change wording in the intro

@@
 
-It is imperative that missing data (or lack thereof --Absence of
-evidence is not evidence of absence) are reported. However, outside of
-sports science, researchers seldom consider missing data [@Schafer2002].
-Moreover, while there has been significant interest in modelling missing
-data [@refs], the exploration and reporting of missing data have received
+It is imperative that missing data, including lack thereof, are reported. However, outside of 
+sports science, there is evidence that researchers seldom consider missing data [@Schafer2002]. 
+While there has been significant interest in modelling missing
+data [@refs], the exploration and reporting of missings have received
 less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
 of articles that report missing data in football research; (2) highlight
-several bad missing data practices; and (3) provide recommendations for
+several improvable missing data practices; and (3) provide recommendations for
 exploring and reporting missingness.

### 2020-09-03 paper/paper.Rmd
change wording in the intro

@@
 
-It is imperative that missing data, including lack thereof, are reported. However, outside of 
-sports science, there is evidence that researchers seldom consider missing data [@Schafer2002]. 
+It is imperative that missing data, or lack thereof, are reported. However, there is 
+evidence that outside of sports science, researchers seldom consider missing data [@Schafer2002]. 
 While there has been significant interest in modelling missing

### 2020-09-03 paper/paper.Rmd
change description of the search, change PRISMA figure details, remove the table for the breakdown of the PRISMA

@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@Schafer2002]. Football related
+football research, we conducted a systematic search [@Schafer2002] **(should this reference be here?)**. Football related
 articles published in 2019 were studied. Three major sports medicine
 databases (SPORTDiscus, Embase and Cinahl) were searched, using key
-terms and search limits (Supplement 1). Included articles were: (a)
-published in 2019; (b) written in English; (c) had an accessible
-full-text; and (d) included quantitative data. The search process is
+terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
+full-text; and (c) included quantitative data. The search process is
 summarised in Figure 1.
@@
 
-prisma(found = x,
+prisma(found = 3863,
 
-found_other = x,
+found_other = NA,
 
-no_dupes = x,
+no_dupes = 1909,
 
-screened = x,
+screened = 1954,
 
-screen_exclusions = x,
+random_sample = 199,
 
-full_text = x,
+full_text = 199,
 
-full_text_exclusions = x,
-
-qualitative = x,
-
-quantitative = x,
+full_text_exclusions = 62,
 
@@
 ```
-
-+---------------------------------+-----------------------------------+
-| Imported                        | 3,863                             |
-+=================================+===================================+
-| After duplicates removed        | 1,954                             |
-+---------------------------------+-----------------------------------+
-| After title and abstract screen | 1,373                             |
-+---------------------------------+-----------------------------------+
-| Random sample full-text         | 199 (needed for 10% sample)       |
-+---------------------------------+-----------------------------------+
-| Full-text included              | 137                               |
-+---------------------------------+-----------------------------------+
-| Full-text excluded              | Full-text papers excluded (n=62)  |
-|                                 | with reasons:                     |
-|                                 |                                   |
-|                                 | Qualitative study (n=)            |
-|                                 |                                   |
-|                                 | Not in english (n=)               |
-|                                 |                                   |
-|                                 | Unable to access full-text (n=)   |
-+---------------------------------+-----------------------------------+
 

### 2020-09-03 paper/paper.Rmd
update details of the search

@@
 full-text; and (c) included quantitative data. The search process is
-summarised in Figure 1.
-
-```{r prisma-example, eval = FALSE}
-require(PRISMAstatement)
-
-prisma()
-
-[[https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html]{.ul}](https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html)
-
-library(PRISMAstatement)
-
-e.g.,
-
-prisma(found = 3863,
-
-found_other = NA,
-
-no_dupes = 1909,
-
-screened = 1954,
-
-random_sample = 199,
-
-full_text = 199,
-
-full_text_exclusions = 62,
-
-width = 800, height = 800)
-```
-
-**Figure 1.** PRISMA flow chart.
+summarised in the PRISMA figure provided in Supplement 2.
 
@@
 a dataset was shared. Where applicable, information about how missing
-values were handled was also extracted (see Supplement 1). Two authors
+values were handled was also extracted (see Supplement 2). Two authors
 independently extracted the data (X% agreement). Findings are reported
-as the proportion and 95% confidence interval. Supplement 1 provides
+as the proportion and 95% confidence interval. Supplement 2 provides
 further details of the search.

### 2020-09-03 paper/paper.Rmd
change wording

@@
 proportion of studies in football research that included a missing data
-statement was low, at only X% [95% confidence interval; X% to X%]; suggesting that 
+statement was low, at only X% (95% confidence interval; X% to X%); suggesting that 
 missingness is seldom considered by researchers. We suspect that this result 

### 2020-09-03 paper/paper.Rmd

@@
 
+A missing value should have been observed or measured in the data, but was not.
+For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event.
+This "goal" is now a missing value.
+Compare this to a game being rained out and no game taking place.
+There is no missing data because there is no data to be observed.
+A final example - a study screens out individuals who do not meet criteria for the study.
+These are not missing values, because there was no intent for these to be observed or measured.
+
 Data does not go missing the same way every time. There are three broad
@@
 occurs by chance (MCAR), due to some observed variable (MAR) or an
-unobserved variable (MNAR), with MCAR being benign and the latter two
-biased. Football relevant examples are provided in Supplement 1, and
+unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR
+biased [@Newman2014]. Football relevant examples are provided in Supplement X, and
 more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
@@
 The three categorisations of missingness types provide a useful
-framework to describe bias arising from missing data. If one strongly
-suspects data is MCAR, one can be more certain your data isn't biased,
-otherwise there may be bias in your sample. This does not mean that data MAR
-or MNAR make data invalid, instead it identifies potential biases, which
-can help improve future research design **(Nick, can we look at this sentence please)**. However, these categorisations
-do not necessarily guide a researcher in their subsequent actions in
+framework to describe bias arising from missing data. If you strongly
+suspect data is MCAR, your results are less likely to be biased. If you suspect
+MAR or MNAR, there is bias in your results. This does not mean data MAR
+or MNAR make results invalid, instead it identifies potential bias, which can be
+used to help improve future research design. However, these categorisations
+do not necessarily guide the user in their subsequent actions in
 analysis, or describe specifically how values go missing. To identify
@@
 
-It is imperative that missing data, or lack thereof, are reported. However, there is 
-evidence that outside of sports science, researchers seldom consider missing data [@Schafer2002]. 
+It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. 
 While there has been significant interest in modelling missing
@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@Schafer2002] **(should this reference be here?)**. Football related
+football research, we conducted a systematic search [@PRISMA]. Football related
 articles published in 2019 were studied. Three major sports medicine

### 2020-09-03 paper/paper.Rmd
add changes from meeting

@@
 proportion of studies in football research that included a missing data
-statement was low, at only X% (95% confidence interval; X% to X%); suggesting that 
+statement was low, at only X% [95% confidence interval; X% to X%]; suggesting that 
 missingness is seldom considered by researchers. We suspect that this result 
@@
 
-**Keywords:** Exercise, imputation, missingness, missings, naniar, sport
+**Keywords:** Imputation, missingness, missings, naniar, sport, football
 
@@
 Contextual factors--such as, the study design and objective, and
-pattern of missingness--determine how missings should be handled, on a
+pattern of missingness--determine how to handle missings on a
 case-by-case basis [@Sainani2015].
-The effect of missing data on the validity and accuracy of a study's
-results is largely determined by the reasons for missingness, and how
-missing data are handled.
 
-A missing value should have been observed or measured in the data, but was not.
-For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event.
-This "goal" is now a missing value.
-Compare this to a game being rained out and no game taking place.
+For clarity, we describe some examples of missingness.
+An Australian football data collector was taking lunch during the third quarter, so no data was recorded for this time period.
+Compare this to an AFL game being cancelled due to COVID19.
 There is no missing data because there is no data to be observed.
-A final example - a study screens out individuals who do not meet criteria for the study.
-These are not missing values, because there was no intent for these to be observed or measured.
+If a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
 
@@
 
-It is imperative that missing data, or lack thereof, are reported [@Schafer2002]. 
-While there has been significant interest in modelling missing
-data [@refs], the exploration and reporting of missings have received
+It is imperative that missing data (or lack thereof --Absence of
+evidence is not evidence of absence) are reported. However, outside of
+sports science, researchers seldom consider missing data [@Schafer2002] (<< NOTE - I'm not sure we mean that?).
+Moreover, while there has been significant interest in modelling missing
+data [@refs], the exploration and reporting of missing data have received
 less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
 of articles that report missing data in football research; (2) highlight
-several improvable missing data practices; and (3) provide recommendations for
+several bad missing data practices; and (3) provide recommendations for
 exploring and reporting missingness.
@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@PRISMA]. Football related
+football research, we conducted a systematic search [@Schafer2002]. Football related
 articles published in 2019 were studied. Three major sports medicine
 databases (SPORTDiscus, Embase and Cinahl) were searched, using key
-terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
-full-text; and (c) included quantitative data. The search process is
-summarised in the PRISMA figure provided in Supplement 2.
+terms and search limits (Supplement 1). Included articles were: (a)
+published in 2019; (b) written in English; (c) had an accessible
+full-text; and (d) included quantitative data. The search process is
+summarised in Figure 1.
+
+```{r prisma-example, eval = FALSE}
+require(PRISMAstatement)
+
+prisma()
+
+[[https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html]{.ul}](https://cran.r-project.org/web/packages/PRISMAstatement/vignettes/PRISMA.html)
+
+library(PRISMAstatement)
+
+e.g.,
+
+prisma(found = x,
+
+found_other = x,
+
+no_dupes = x,
+
+screened = x,
+
+screen_exclusions = x,
+
+full_text = x,
+
+full_text_exclusions = x,
+
+qualitative = x,
+
+quantitative = x,
+
+width = 800, height = 800)
+```
+
++---------------------------------+-----------------------------------+
+| Imported                        | 3,863                             |
++=================================+===================================+
+| After duplicates removed        | 1,954                             |
++---------------------------------+-----------------------------------+
+| After title and abstract screen | 1,373                             |
++---------------------------------+-----------------------------------+
+| Random sample full-text         | 199 (needed for 10% sample)       |
++---------------------------------+-----------------------------------+
+| Full-text included              | 137                               |
++---------------------------------+-----------------------------------+
+| Full-text excluded              | Full-text papers excluded (n=62)  |
+|                                 | with reasons:                     |
+|                                 |                                   |
+|                                 | Qualitative study (n=)            |
+|                                 |                                   |
+|                                 | Not in english (n=)               |
+|                                 |                                   |
+|                                 | Unable to access full-text (n=)   |
++---------------------------------+-----------------------------------+
+
+**Figure 1.** PRISMA flow chart.
 
@@
 a dataset was shared. Where applicable, information about how missing
-values were handled was also extracted (see Supplement 2). Two authors
+values were handled was also extracted (see Supplement 1). Two authors
 independently extracted the data (X% agreement). Findings are reported
-as the proportion and 95% confidence interval. Supplement 2 provides
+as the proportion and 95% confidence interval. Supplement 1 provides
 further details of the search.

### 2020-09-03 paper/paper.Rmd

--- 2020-09-03 paper/paper.Rmd
@@
 case-by-case basis [@Sainani2015].
-The effect of missing data on the validity and accuracy of a study's
-results is largely determined by the reasons for missingness, and how
-missing data are handled.
 
-A missing value should have been observed or measured in the data, but was not.
-For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event.
-This "goal" is now a missing value.
-Compare this to a game being rained out and no game taking place.
+For clarity, we describe some examples of missingness.
+An Australian football data collector was taking lunch during the third quarter, so no data was recorded for this time period.
+Compare this to an AFL game being cancelled due to COVID19.
 There is no missing data because there is no data to be observed.
-A final example - a study screens out individuals who do not meet criteria for the study.
-These are not missing values, because there was no intent for these to be observed or measured.
+If a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
 

### 2020-09-03 paper/paper.Rmd
add note on missing in Xs and Ys

@@
 missing otherwise delete rows with <5% missing) and caution against
-such practice. Rather, we recommend values should generally be imputed..
+such practice. Rather, we recommend values should generally be imputed.
+
+Missing values can occur in the both the predictors (the "independent variables") and response variable (the "dependent variables").
+Care should be taken when imputing data, as typically only predictors should be imputed.
+For discussion on imputing the response value, see ...
+[@refs].
+(note: take a look again)
+
 We recommend avoiding imputation methods that impute the same value,
@@
 
-The current state of reporting missing data in the sports literature is
-poor. In this section we discuss recommended practices for exploring and
+The current state of reporting missing data in the sports literature has room for improvement. In this section we discuss recommended practices for exploring and
 reporting missing data. Our intention is not to provide a 'cookbook'

### 2020-09-03 paper/paper.Rmd
adjust title and conflict of interest: closes #9

@@
 ---
-title: "Missing data: Current practice in football research and recommendations for reporting missingness"
+title: "Missing data: Current practice in football research and recommendations for improvement"
 authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
@@
 **Acknowledgement/conflict of interest**
-
-NJT naniar package?
 

### 2020-09-03 paper/supplement-2.Rmd
add figure caption

@@
   bookdown::html_document2
+  # bookdown::pdf_document2
 bibliography: "supp-2.bib"
@@
 
-```{r prisma}
+```{r prisma, fig.cap = "Write the figure caption here"}
   # | Imported                        | 3,863                             |

### 2020-09-03 R/read-gs.R
work on finalising the spreadsheet

@@ -1,7 +1,41 @@
+library(tidyverse)
 library(googlesheets4)
+library(janitor)
 # insert your own email address
-gs4_auth("nicholas.tierney@monash.edu") 
+gs4_auth(email = "nicholas.tierney@monash.edu") 
 sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
-gs_football <- read_sheet(sheet_url, skip = 1)
+gs_football_r1 <- read_sheet(sheet_url, range = "A2:O201")
+gs_football_r2_cols <- read_sheet(sheet_url, range = "A2:H201")
+gs_football_r2_results <- read_sheet(sheet_url, range = "P2:V201")
 
-gs_football
+gs_football_r2 <- bind_cols(gs_football_r2_cols,
+                            gs_football_r2_results)
+
+gs_football <- bind_rows(reviewer_1 = gs_football_r1,
+                         reviewer_2 = gs_football_r2,
+                         .id = "reviewer") %>% 
+  clean_names() 
+
+unlist_as_factor <- function(x){
+  x %>%
+    na_if("NA") %>%
+    unlist() %>% 
+    as.logical()
+}
+
+gs_football %>% 
+  select(where(is.list)) %>% 
+  pull(1) %>% 
+  unlist()
+  
+gs_football %>% 
+  mutate(any_missing_data_statement = unlist_as_logical(any_missing_data_statement))
+
+gs_football %>% 
+  mutate_if(.predicate = is.list,
+            unlist_as_logical)
+  
+
+library(visdat)
+vis_dat(gs_football, sort_type = FALSE)  
+

### 2020-09-03 R/read-gs.R
fix issue with incorrect data type on read in, related to #4

@@ -4,9 +4,15 @@ library(janitor)
 # insert your own email address
 gs4_auth(email = "nicholas.tierney@monash.edu") 
 sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
-gs_football_r1 <- read_sheet(sheet_url, range = "A2:O201")
-gs_football_r2_cols <- read_sheet(sheet_url, range = "A2:H201")
-gs_football_r2_results <- read_sheet(sheet_url, range = "P2:V201")
+gs_football_r1 <- read_sheet(sheet_url, 
+                             range = "A2:O201",
+                             col_types = "ciccccdcccccccc")
+gs_football_r2_cols <- read_sheet(sheet_url, 
+                                  range = "A2:H201",
+                                  col_types = "ciccccdc")
+gs_football_r2_results <- read_sheet(sheet_url, 
+                                     range = "P2:V201",
+                                     col_types = "ccccccc")
 
 gs_football_r2 <- bind_cols(gs_football_r2_cols,
                             gs_football_r2_results)
@@ -14,28 +20,19 @@ gs_football_r2 <- bind_cols(gs_football_r2_cols,
 gs_football <- bind_rows(reviewer_1 = gs_football_r1,
                          reviewer_2 = gs_football_r2,
                          .id = "reviewer") %>% 
-  clean_names() 
-
-unlist_as_factor <- function(x){
-  x %>%
-    na_if("NA") %>%
-    unlist() %>% 
-    as.logical()
-}
-
-gs_football %>% 
-  select(where(is.list)) %>% 
-  pull(1) %>% 
-  unlist()
-  
-gs_football %>% 
-  mutate(any_missing_data_statement = unlist_as_logical(any_missing_data_statement))
-
-gs_football %>% 
-  mutate_if(.predicate = is.list,
-            unlist_as_logical)
-  
+  clean_names()  %>% 
+  mutate(study_eligibility = as.logical(study_eligibility))
 
 library(visdat)
 vis_dat(gs_football, sort_type = FALSE)  
 
+# preliminary study
+gs_football %>% 
+  filter(study_eligibility) %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement) 
+
+gs_football %>% 
+  filter(study_eligibility) %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement)

### 2020-09-04 paper/supplement-3.Rmd
update simulated dataset for missing data example write up

@@
 
-(Content pulled from the google doc)
+```{r}
+library(tidyverse)
+library(fitzRoy)
+library(mice)
+library(naniar)
 
-Rob's idea \[case study section\]:
+# Aim of this section:
+# Simulate data from a real data set, insert missing
+# Mini study: predicting super coach scores
 
--   Including a spectrum of worst practice - best practice
+d = fitzRoy::get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
 
-    -   Worst
+data <- d %>% select(ID,  Round, First.name, Surname, Kicks, Handballs, Marks, Hit.Outs, Contested.Possessions, Uncontested.Possessions)
+#View(data)
 
-        -   Missing data can be observed in how statistical methods are
-            > reported
+df <- fitzRoy::get_footywire_stats(10286:10290) # Get the fantasy scores
+View(df)
+df #%>% View("footywire dataset")
 
-            -   e.g., missing meters gained
+# Simulate data from summarises of the above dataset
+rnorm(n = 100, # N is the amount of data points being generated
+      mean = 5,
+      sd = 2)
+sample.int(100, # numbers between 1 and 100
+           10) # observations
+rpois(n = 100,
+      lambda = 5)
 
-    -   less worse
+# Generate the y-variable
+fit <- lm(AF ~ K + HB, data = df)
+summary(fit)
+data %>% mutate(fanasty_score = 4.106*Kicks + 2.175*Handballs)
 
-        -   Missing data reported, but no handling stated
+# Completed dataset
+# Drop a team
+No_Richmond <- df %>% filter(Team!="Richmond")
+sort(unique(No_Richmond$Team))
+sort(unique(df$Team))
 
-            -   e.g., 17% of data is missing, nothing mentioned about
-                > deletion/imputation\
-                > [[https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ]{.ul}](https://www.sciencedirect.com/science/article/pii/S0167945719301939?casa_token=QfaCiQJaBQMAAAAA:tL3pKZG7s8-umTgU-cVZ53ZuPLchja5OExOEnkohWhjLH4NPGtqeFkSWlZq0EiPG9ip-ZnQ)
+# Impute missing values
+df_first_try <- mice::ampute(data = df[,10:43])
+df_first_try <- df_first_try$amp
 
-            -   e.g., N = 18, and statistical analysis reports N = 14
-                > (Robin, tweet, Dave to request paper)
+# Examples of presenting of missing values
+naniar::vis_miss(df_first_try)
+naniar::vis_miss(df_first_try)
+naniar::gg_miss_fct(x = riskfactors, fct = marital)
 
-    -   neutral?
+# Example write up of missing values
 
-        -   There is missing data, and this is reported?
+# Things to do
+# Get Nick's thoughts on what we include; and how to write it up.
+# What do we drop to introduce bias, a team, a position, etc.
 
-    -   good
-
-        -   Missing data reported, handling is reported, might be bad
-            > imputation like mean imputation / data not shared.
-
-    -   best
-
-        -   Missing data reported, handling is reported, appropriate
-            > imputation, data shared.
+```
 
@@
 ```
--   Ways to visualise missing data for reviewers, and readers of paper
-    > to get a feel of how much to take away.
-
-    -   So the meters gained example, I think a vis could be something
-        > like the figure below. Also a way to visualise missing values
-        > in athlete studies over time, if for whatever reason on
-        > X1,x7,x35 day couldn't do a measurement on athlete 1.
-        > \^\^instead of google sheet something ggplot\^\^
-
-        -   vis_miss from visdat/naniar:
-
-![](media/image2.png){width="6.267716535433071in"
-height="0.8888888888888888in"}
-
-\^\^Background in sports and exercise science\^\^
-
--   Missing data (or the absence of missing data) is often not reported
-
--   In general, imputation is not used
-
--   Missing data is never really discussed
-
--   Data are usually fit with ANOVA's, therefore values are subbed in,
-    > e.g., mean. (really common practice)
-
-**Notes/examples**
-
-Example 1:
-[[https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141]{.ul}](https://www.sciencedirect.com/science/article/abs/pii/S1440244018303141)
-
-Problem---The most important predictor, meters gained relative isn\'t
-available in the vast majority of the dataset (2001-2008 for example)
-(which is where the \'\'prediction accuracy\'\' was the worst. This
-isn\'t mentioned anywhere in the paper, that their most important
-predictor isn\'t available. So as a reviewer, unless data is provided,
-or code shared. You are relying on domain knowledge to know this.
-
-The supplementary materials which provides coefficients for the GLMS
-fitted
-
-![](media/image13.png){width="6.270833333333333in"
-height="5.888888888888889in"}
-
-No meters gained? This is weird right, talk about a circle, we have
-meters gained most important, but GLM perform similar to decision tree
-but the GLM doesn't have meters gained (the most important variable)
-according to the decision trees?
-
-Oh and meters gained isn't available throughout the dataset.
-
-"GLMs were also created to predict Score Margin in each timeframe and
-the prediction accuracies are similar for each time frame but are higher
-than the decision tree prediction accuracies."
-
-And whats with figure 2?
-
-![](media/image15.png){width="6.270833333333333in" height="4.375in"}
-
-The end buckets are supposed to be meters gained difference, but we have
-a ?
-
-Example 2:
-[[https://twitter.com/championdata/status/472174368059957248]{.ul}](https://twitter.com/championdata/status/472174368059957248)
-
-![](media/image5.png){width="3.7031255468066493in"
-height="4.275152012248469in"}
-
-![](media/image3.png){width="6.270833333333333in"
-height="2.8333333333333335in"}
-
-So if meters gained was first recorded in 2007, what does this mean for
-the time frames being compared?
-
-![](media/image11.png){width="6.270833333333333in"
-height="2.2222222222222223in"}
-
-Like OK that is 2009-2016, but what about the 2001-2016 was meters
-gained only used for 2007-2016 so what was inputted 2001-2007 and if we
-compared it to 2009-2016 where meters gained was FULLY available we have
-a worse win/loss pred and a in the dataset with the MOST IMPORTANT
-VARIABLE AVAILABLE ALL THE WAY THROUGH IT I am really confused.
-
-**Modelling Match Outcome in Australian Football: Improved accuracy with
-large databases **
-
-(similar authors as above---Tran, Young etc)
-
-![](media/image10.png){width="3.1875240594925636in"
-height="0.7656889763779527in"}
-
-Problem: The most important variable, in 2019 published without
-mentioning its availability, nor are results presented in a way that
-gives the reader an idea of what was done? Seriously your trying to
-predict an outcome and literally the variable doesn\'t EXIST what was
-entered in?
-
-Data sharing/code sharing would have made this available to reviewers
-and perhaps changed the tone of the papers.
-
-Example 3:\
-[[https://shapeamerica.tandfonline.com/doi/abs/10.1080/02640414.2015.1066026\#.XpkfaFMzZhE]{.ul}](https://protect-au.mimecast.com/s/FqW3CZYMqwSk7AkMsKELku?domain=shapeamerica.tandfonline.com)
-
-These results on data available to public (basic stuff like counting of
-kicks isn\'t reproducible)---i.e., the table here:
-
-![](media/image9.png){width="3.557292213473316in"
-height="2.951379046369204in"}
-
-Means should be available?
-
-Example 4:
-[[https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full]{.ul}](https://journals.sagepub.com/eprint/Pv3bEcp6EQMnPAa24cuP/full)
-
-Problem:
-[[https://twitter.com/RankingSw/status/961502281874092032]{.ul}](https://twitter.com/RankingSw/status/961502281874092032)
-
-![A screenshot of a cell phone Description automatically
-generated](media/image7.png){width="1.4457130358705161in"
-height="2.085090769903762in"}
-
-![](media/image14.png){width="1.4609733158355205in"
-height="1.6510684601924759in"}
-
-So why didn\'t the authors of the paper know this? Was it because the
-underlying model (AFLPR) wasn\'t available at the time? etc But
-certainly as a reviewer this is a thing someone should know before
-reviewing?
-
-Example 5:
-[[https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511]{.ul}](https://www.tandfonline.com/doi/abs/10.1080/02640414.2015.1066511)
-
-Figure below seems odd:
-
-![](media/image4.png){width="2.2067705599300087in"
-height="1.4527034120734907in"}
-
--   Progression of science, self-correction and the need for
-    > 'self-correction to not be strike against your name'
-
--   Speak about the content in the papers wanting to push the knowledge,
-    > not just get rushed for the purpose of a publication.
-
-Recommendations -- Ask the question: what am I being provided here to
-review?
-
--   Barriers
-
-    -   Data privacy agreements (e.g., NDA), the reviewer would have to
-        > sign one, maybe this is something that's not can occur, so
-        > what's the minimum.
-
-    -   Betting privacy, players union---you couldn't ever share who the
-        > slowest player in the NRL same with the AFL
-
-```{=html}
-<!-- -->
-```
--   Framed from the reviewer's point of view, needs to be a minimum
-    > understanding of what is being done in the paper.
-
--   Exploratory and missingness (common problem in sports, maybe not
-    > able to get access to the player)
-
-```{=html}
-<!-- -->
-```
--   Minimum requirements: please give the summary stats, i.e., the
-    > variable name, description, the dates collect, number of [missing
-    > values]{.ul}, etc.
-
-    -   Describe the data in some type of meaningful sense
-
-    -   If there's definitions, over what period does the definition
-        > hold for.
-
--   Idea scenario: Start data, end date.
-
-Example from Twitter message
-
-Dave sent a twitter message about this paper here.
-
-[[https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490]{.ul}](https://sci-hub.tw/https://doi.org/10.1080/02640414.2020.1740490)
-
-From the Paper
-
-This study used commercially available data from the 2015 AFL season
-supplied by Champion Data, the official statistics provider for the AFL.
-An exemption from ethics review was granted by the institutional ethics
-committee. The original data set included all 149,449 possession events
-from all matches played during the 2015 AFL season (18 teams, 665
-players, 206 regular season and finals series games). A data subset was
-created to focus the analysis on regular season matches in order to
-explore wholeof-league patterns (i.e., finals matches were omitted as
-they are only played by teams in the top 8 end-of-season ladder
-positions). This subset was filtered further to include only those
-sequences of possession events that i) started with a kick-in following
-a behind, and ii) ended with the offensive team losing possession
-through a score, stoppage, or turnover. These sequences are referred to
-herein as "kick-in possession chains", where a sequence begins with the
-player who first receives the ball from a kick-in and excludes the
-initial disposal out of the goal square, unless the player starting with
-possession "plays on" to himself. Possession chains that involved only
-one player or were incomplete due to end-of-quarter time restrictions
-were excluded. Chains from 2 drawn matches were also excluded. The final
-data set used for analysis comprised a total of 1,720 kick-in possession
-chains from 194 games.
-
-Now thankfully we can use fitzRoy to check out how many behinds there
-were in 2015 as per the subset.
-
-> fitzRoy::get_afltables_stats(start_date=\"2015-01-01\",
-> end_date=\"2016-01-01\")%\>% \#Season chosen
->
-> select(Season,Home.score, Away.score, Round, Home.team, Away.team,
-> Round, AQ4B, HQ4B)%\>%\# variables needed
->
-> filter(!Round %in% c(\"EF\", \"SF\",\"QF\",\"PF\", \"GF\"))%\>%
-> \#excluding finals
->
-> filter(Home.score != Away.score)%\>% \#excluding draws
->
-> distinct()%\>%
->
-> mutate(totalbehinds=AQ4B+HQ4B)%\>% \#away behinds + home behinds
->
-> summarise(totalbehindsin2015=sum(totalbehinds))
->
-> totalbehindsin2015
->
-> \<int\>
->
-> 1 4275
->
-> \>
->
-> So if each quarter ended with a invalid observation (didn't result in
-> a score, turnover etc) this would be pretty high I'd imagine but help
-> illustrate the point
->
-> 4\*194
->
-> So we have 4275 kick ins from behinds, of which at MOST 194\*4 (776)
-> would be excluded say if every quarter had a kick after the siren that
-> resulted in a behind.
->
-> So that leaves us with 4275-776 total kick ins which is 3499 kick in
-> ins total
->
-> **So this data suggests that only 1720/3499 or 0.491569 of kick ins
-> resulted in a chain? in the study.**
->
-> This seems not good.
->
-> To put another way,
->
-> out of the total behinds or ''opportunities for chains to start'' if
-> we take the exclusion rule. **" Possession chains that involved only
-> one player or were incomplete due to end-of-quarter time restrictions
-> were excluded."**
->
-> **Then**
->
-> **(4275-1720 )/ 4275 or 0.5976608 of chains are EXCLUDED that seems
-> massive!**
-
-![](media/image12.png){width="6.270833333333333in"
-height="1.8194444444444444in"}
-
-New Paper: Recommending how to share data in sports science
-
--   Simplest approach
-
-    -   If you report the summary statistics for each variable
-
-    -   Then recreate the data from that
-
-        -   The min-max and sample evenly
-
-        -   the distribution

### 2020-09-04 paper/paper.Rmd
change wording in the intro.

@@
 2: University of New South Wales, Department of Statistics, School of
-Mathematics and Statistics, Sydney, Australia
+Mathematics and Statistics, Sydney, Australia.
 
 3: Monash University, Department of Econometrics and Business Statistics,
-Melbourne, Australia
+Melbourne, Australia.
 
 4: Australian Centre of Excellence for Mathematical and Statistical
-Frontiers (ACEMS), Melbourne, Australia
+Frontiers (ACEMS), Melbourne, Australia.
 
 **Acknowledgement/conflict of interest**
+
+The authors declare no conflicts of interest.
 
@@
 Nicholas J Tierney
-
 Monash University, Department of Econometrics and Business Statistics,
-Melbourne, Australia
+Melbourne, Australia.
+Address:
+Email:
 
@@
 with decisions made of how missing data are handled (deleted or
-imputed), can have a profound effect on the validity, usability and
-accuracy of a study's results. In this article, we aimed to: estimate
+imputed), can have a profound effect on the validity, accuracy and usability 
+of a study's results. In this article, we aimed to: estimate
 the proportion of studies in football research that included a missing
-data statement, highlight several improvable practices in relation to
+data statement, highlight several practices to avoid in relation to
 missing data, and provide recommendations for exploring, visualising and
@@
 statement was low, at only X% (95% confidence interval; X% to X%); suggesting that 
-missingness is seldom considered by researchers. We suspect that this result 
-is consistent with the wider sports science literature. We recommend 
+missingness is seldom considered by researchers. It is suspected that this result 
+is consistent with practice in the wider sports science literature. We recommend 
 researchers describe the number and percentage of missing
@@
 values are unobserved, this can mean the population of interest might not be 
-properly sampled, inducing bias into study results. Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results
+properly sampled, inducing bias into a study's results. Left undetected or poorly handled, 
+missing data can undermine the validity and accuracy of research results
 [@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
@@
 
-For clarity, we describe some examples of missingness.
-An Australian football data collector was taking lunch during the third quarter, so no data was recorded for this time period.
-Compare this to an AFL game being cancelled due to COVID19.
-There is no missing data because there is no data to be observed.
-If a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
+For clarity, we describe an example of missingness. An Australian Rules Football data collector 
+was taking lunch during the third quarter, so no data was recorded for this time period. This
+would be considered missing data. Compare this to the game being cancelled due to a pandemic. 
+There is no missing data because there is no data to be observed. Similiarly, if a study screens out individuals who 
+do not meet inclusion criteria, this is not a missing value, because there is no intent for these 
+values to be observed or measured.
 

### 2020-09-04 R/paper-statement-analysis.R
update wording in the intro.

@@ -1,41 +1,0 @@
-library(binom)
-library(here)
-library(tidyverse)
-library(conflicted)
-library(googlesheets4)
-conflicted::conflict_prefer("select", "dplyr")
-conflicted::conflict_prefer("filter", "dplyr")
-
-# insert your own email address
-gs4_auth("nicholas.tierney@monash.edu") 
-sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
-gs_football <- read_sheet(sheet_url, skip = 1)
-
-gs_football
-
-
-#### Calculate proportions
-d <- read_csv(here("data",
-                  "missing-data-in-football-preliminary.csv"))
-
-d <- filter(d, Study_eligibility == "1")
-
-# Reporting missing
-table(d$Any_missing_data_statement, useNA = 'ifany')
-successes <- sum(d$Any_missing_data_statement == 1, na.rm = T)
-trials <- sum(!is.na(d$Any_missing_data_statement))
-binom.confint(x = successes, n = trials, conf.level = 0.95, methods = "exact")
-
-# Non-reported missing who shared data
-d_shared_data = filter(d, Any_missing_data_statement == "0")
-table(d_shared_data$Was_data_shared, useNA = 'ifany')
-successes <- sum(d_shared_data$Was_data_shared == 1, na.rm = T)
-trials <- sum(!is.na(d_shared_data$Was_data_shared))
-binom.confint(x = successes, n = trials, conf.level = 0.95, methods = "exact")
-
-
-
-# Shared code
-table(d$Code_shared, useNA = 'ifany')
-# !!comment to explain NAs
-# Zero shared code

### 2020-09-04 R/read-gs.R

@@ -1,38 +1,0 @@
-library(tidyverse)
-library(googlesheets4)
-library(janitor)
-# insert your own email address
-gs4_auth(email = "nicholas.tierney@monash.edu") 
-sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
-gs_football_r1 <- read_sheet(sheet_url, 
-                             range = "A2:O201",
-                             col_types = "ciccccdcccccccc")
-gs_football_r2_cols <- read_sheet(sheet_url, 
-                                  range = "A2:H201",
-                                  col_types = "ciccccdc")
-gs_football_r2_results <- read_sheet(sheet_url, 
-                                     range = "P2:V201",
-                                     col_types = "ccccccc")
-
-gs_football_r2 <- bind_cols(gs_football_r2_cols,
-                            gs_football_r2_results)
-
-gs_football <- bind_rows(reviewer_1 = gs_football_r1,
-                         reviewer_2 = gs_football_r2,
-                         .id = "reviewer") %>% 
-  clean_names()  %>% 
-  mutate(study_eligibility = as.logical(study_eligibility))
-
-library(visdat)
-vis_dat(gs_football, sort_type = FALSE)  
-
-# preliminary study
-gs_football %>% 
-  filter(study_eligibility) %>% 
-  group_by(reviewer) %>% 
-  count(any_missing_data_statement) 
-
-gs_football %>% 
-  filter(study_eligibility) %>% 
-  group_by(reviewer) %>% 
-  count(any_missing_data_statement)

### 2020-09-04 R/sampling-papers.R

@@ -1,25 +1,0 @@
-# Code to sample the papers for review
-# possible papers to review after title abstract screening are in file
-# "missing-data-endnote-library.csv"
-# We then sampled possible papers to get to 10% of the number of screened papers
-# Two files were written, "random-sample-137", and "random-sample-62". These two
-# were created to get close to 10% of the number of papers in total 199 papers
-# were reviewed.
-
-library(tidyverse)
-library(here)
-d <- read_csv(here("data",
-                   "missing-data-endnote-library.csv"))
-
-# repeatable analysis
-set.seed(2020-05-21)
-articles <- d %>% sample_frac(size = 0.1)
-#write.csv(articles, file = "random-sample-137.csv")
-
-articles$sampled <- 1
-data <- merge(d, articles, all.x = T)
-data$sampled <- data$sampled %>% replace_na(0)
-d_second_sample <- filter(data, sampled < 1)
-articles_second_sample <- d_second_sample %>% sample_frac(size = 0.05)
-#write.csv(articles_second_sample, file = "random-sample-62.csv")
-

### 2020-09-04 paper/figures/create-table.R

@@ -1,45 +1,0 @@
-library(magick)
-library(tibble)
-library(gt)
-library(tidyr)
-
-player_table_implicit <- tibble::tribble(
-  ~Player,    ~Quarter,   ~Goals,
-  "Coleman",   1,         0,
-  "Coleman",   2,         1,
-  "Coleman",   3,         2,
-  "Coleman",   4,         0,
-  "Koenen",    1,         1,
-  "Koenen",    3,         2
-)
-
-player_table_explicit <- player_table_implicit %>% 
-  pivot_wider(id_cols = Player,
-              names_from = Quarter,
-              values_from = Goals)
-
-gt(player_table_implicit) %>% 
-  gtsave("paper/figures/implicit.png")
-
-gt(player_table_explicit) %>% 
-  tab_style(
-    style = cell_fill(color = "lightblue"),
-    locations = cells_body(
-      columns = vars(`2`,`4`),
-      rows = is.na(`2`) | is.na(`4`)
-    )
-  ) %>% 
-  gtsave("paper/figures/explicit.png")
-
-img_implicit <- image_read("paper/figures/implicit.png")
-img_explicit <- image_read("paper/figures/explicit.png")
-img_arrow <- image_read("paper/figures/arrow.png")
-
-img_tables <- 
-image_append(c(img_implicit, 
-               {image_scale(img_arrow, "10%") %>% 
-                   image_border("white", geometry = "50x220x50x50")
-                 },
-               image_border(img_explicit, "white", geometry = "0x142+0+0")))
-
-image_write(img_tables, "paper/figures/tables.png")

### 2020-09-04 R/paper-statement-analysis.R
change wording in the search.

@@ -1,0 +1,41 @@
+library(binom)
+library(here)
+library(tidyverse)
+library(conflicted)
+library(googlesheets4)
+conflicted::conflict_prefer("select", "dplyr")
+conflicted::conflict_prefer("filter", "dplyr")
+
+# insert your own email address
+gs4_auth("nicholas.tierney@monash.edu") 
+sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
+gs_football <- read_sheet(sheet_url, skip = 1)
+
+gs_football
+
+
+#### Calculate proportions
+d <- read_csv(here("data",
+                  "missing-data-in-football-preliminary.csv"))
+
+d <- filter(d, Study_eligibility == "1")
+
+# Reporting missing
+table(d$Any_missing_data_statement, useNA = 'ifany')
+successes <- sum(d$Any_missing_data_statement == 1, na.rm = T)
+trials <- sum(!is.na(d$Any_missing_data_statement))
+binom.confint(x = successes, n = trials, conf.level = 0.95, methods = "exact")
+
+# Non-reported missing who shared data
+d_shared_data = filter(d, Any_missing_data_statement == "0")
+table(d_shared_data$Was_data_shared, useNA = 'ifany')
+successes <- sum(d_shared_data$Was_data_shared == 1, na.rm = T)
+trials <- sum(!is.na(d_shared_data$Was_data_shared))
+binom.confint(x = successes, n = trials, conf.level = 0.95, methods = "exact")
+
+
+
+# Shared code
+table(d$Code_shared, useNA = 'ifany')
+# !!comment to explain NAs
+# Zero shared code

### 2020-09-04 R/read-gs.R

@@ -1,0 +1,38 @@
+library(tidyverse)
+library(googlesheets4)
+library(janitor)
+# insert your own email address
+gs4_auth(email = "nicholas.tierney@monash.edu") 
+sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
+gs_football_r1 <- read_sheet(sheet_url, 
+                             range = "A2:O201",
+                             col_types = "ciccccdcccccccc")
+gs_football_r2_cols <- read_sheet(sheet_url, 
+                                  range = "A2:H201",
+                                  col_types = "ciccccdc")
+gs_football_r2_results <- read_sheet(sheet_url, 
+                                     range = "P2:V201",
+                                     col_types = "ccccccc")
+
+gs_football_r2 <- bind_cols(gs_football_r2_cols,
+                            gs_football_r2_results)
+
+gs_football <- bind_rows(reviewer_1 = gs_football_r1,
+                         reviewer_2 = gs_football_r2,
+                         .id = "reviewer") %>% 
+  clean_names()  %>% 
+  mutate(study_eligibility = as.logical(study_eligibility))
+
+library(visdat)
+vis_dat(gs_football, sort_type = FALSE)  
+
+# preliminary study
+gs_football %>% 
+  filter(study_eligibility) %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement) 
+
+gs_football %>% 
+  filter(study_eligibility) %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement)

### 2020-09-04 R/read-markdown.R

@@ -1,0 +1,3 @@
+library(markdrive)
+
+gdoc_checkout("Missing data in sports and exercise science research")

### 2020-09-04 paper/paper.Rmd
change wording in the search.

@@
 
-For clarity, we describe an example of missingness. An Australian Rules Football data collector 
+For clarity, we describe an example of missingness. An Australian Football data collector 
 was taking lunch during the third quarter, so no data was recorded for this time period. This
 would be considered missing data. Compare this to the game being cancelled due to a pandemic. 
-There is no missing data because there is no data to be observed. Similiarly, if a study screens out individuals who 
-do not meet inclusion criteria, this is not a missing value, because there is no intent for these 
-values to be observed or measured.
+There is no missing data because there is no data to be observed.
+Similiarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
 
@@
 unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR
-biased [@Newman2014]. Football relevant examples are provided in Supplement X, and
+biased [@Newman2014]. Football relevant examples are provided in Supplement 1, and
 more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
@@
 of articles that report missing data in football research; (2) highlight
-several improvable missing data practices; and (3) provide recommendations for
+several practices that should be avoided; and (3) provide recommendations for
 exploring and reporting missingness.
@@
 
-Our intention was to survey, at random, 10% of the total articles found
-(after title and abstract screening). We were primarily interested in
-whether missing data were reported or acknowledged; and if not, whether
-a dataset was shared. Where applicable, information about how missing
-values were handled was also extracted (see Supplement 2). Two authors
+Our intention was to survey, at random, 10% of the total articles found, after 
+title and abstract screening. We were primarily interested in
+whether missing data was reported or acknowledged; and if not, whether
+a dataset was shared. Where applicable, information relating to how missing
+values were treated was also extracted (Supplement 2). Two authors
 independently extracted the data (X% agreement). Findings are reported
-as the proportion and 95% confidence interval. Supplement 2 provides
-further details of the search.
+as the proportion and 95% confidence interval. 
+Further details of the search can be found in Supplement 2.
 

### 2020-09-04 paper/paper.Rmd
alter search wording.

@@
 excluded due to not meeting the inclusion criteria. The proportion of
-articles that reported missing data was 5.9% (95% confidence interval
+articles that reported missing data was 5.9% (95% confidence interval;
 X% to X%) or 8/136. Of the articles that did not mention missing data
@@
 data are not being reported, it does suggest that this aspect of the
-analysis process receives little consideration -- especially in contrast
+analysis process receives little consideration--especially in contrast
 to many of the papers that included (potentially meaningless) tests of

### 2020-09-04 paper/paper.Rmd
changes in discussion.

@@
 wider sports science literature. This finding may be largely explained
-by a lack of awareness and education on missing data practices (Kristin
-BJSM paper), similar to other areas of sciences (refs-education and
-ecology examples)(2). Below we discuss several poor practices in
+by a lack of awareness and education on missing data practices (https://bjsm.bmj.com/content/early/2020/08/19/bjsports-2020-102607.abstract), 
+similar to other areas of sciences (refs-education and
+ecology examples)(2). Below we discuss several practices to avoid in
 relation to missing data, and provide recommendations for exploring,
@@
 
-Case study: *Unreported missing values*. While not always the case, even if
+*Unreported missing values*. While not always the case, even if
 not reported, missing data can be obvious. A recent paper examined the
 relationship between match performance indicators and outcome, in
-Australian Football between the 2001 and 2016. Problematically, at least
-one of the 91 team performance indicators included in the analysis was
-not available over the entire 2001-2016 period [@Young2019]. The variable
-'meters gained' was not collected over the specified duration [@ref].
+Australian Football between the 2001 and 2016. Of the 91 team performance indicators included in the analysis,
+at least one variable (i.e., 'meters gained') was not available over the entire 2001-2016 period [@Young2019]**Check this reference**.
 This was not noted anywhere in the paper, or in the supplementary
@@
 analysis. It is also difficult for a reader to evaluate how the
-missingness could bias the results. Unfortunately, the absence of a
+missingness could bias the results. The absence of a
 missing data statement when data are missing is not unique to the study
-above [@Young2019]. Authors should include a missing data statement,
-irrespective of whether data is missing, or not.
+above [@Young2019]**Check this reference**. Authors should include a missing data statement,
+irrespective of whether there is data missing, or not.
 
-Case study: *Mean imputation.* The default of most statistical software for
+*Mean imputation.* The default of most statistical software for
 handling missing data in modelling is listwise deletion. Listwise
@@
 method is documented. One imputation method is imputing the mean value
-(e.g., participant or group average) [@Young2019]. While imputing the mean
+(e.g., participant or group average) [@Young2019]**Check this reference**. While imputing the mean
 preserves the study sample size (and point estimates), it also reduces
@@
 
-Case study: *Not evaluating the effect, or choice, of imputation.* The
+*Not evaluating the effect, or choice, of imputation.* The
 imputation method used to impute missing data has the potential to

### 2020-09-04 paper/paper.Rmd
wording
changes in search method.

@@
 terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
-full-text; and (c) included quantitative data. The search process is
-summarised in the PRISMA figure provided in Supplement 2.
+full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process.
 
@@
 a dataset was shared. Where applicable, information relating to how missing
-values were treated was also extracted (Supplement 2). Two authors
+values were treated was also extracted (see Supplement 2). Two authors
 independently extracted the data (X% agreement). Findings are reported

### 2020-09-04 paper/paper.Rmd
wording in the search method.
search method wording.

@@
 terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
-full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process.
+full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process,
+and further details of the search can also be found in Supplement 2.
 
-Our intention was to survey, at random, 10% of the total articles found, after 
-title and abstract screening. We were primarily interested in
+Our intention was to survey, at random, 10% of the total articles found (after 
+title and abstract screening). We were primarily interested in
 whether missing data was reported or acknowledged; and if not, whether
@@
 as the proportion and 95% confidence interval.
-Further details of the search can be found in Supplement 2.
 

### 2020-09-05 paper/paper.Rmd
edit results section.
edit first paragraph of the discussion.

@@
 
-Of the 199 articles screened, 136 were deemed eligible, representing
-\~10% of the articles found (Figure X). XX articles (of the 199) were
-excluded due to not meeting the inclusion criteria. The proportion of
+Of the 199 articles screened, 136 met the inclcusion criteria, representing
+\~10% of the articles found (see Supplement 2, Figure 1). The proportion of
 articles that reported missing data was 5.9% (95% confidence interval;
@@
 that only about one in 20 papers reported missing data. It is reasonable
-to assume that this result is representative of current practices in the
-wider sports science literature. This finding may be largely explained
-by a lack of awareness and education on missing data practices (https://bjsm.bmj.com/content/early/2020/08/19/bjsports-2020-102607.abstract), 
+to assume that this result is representative of current practice in the wider sports science literature. 
+The low proportion of papers reporting missingness could be explained by a lack of awareness and education on missing data (https://bjsm.bmj.com/content/early/2020/08/19/bjsports-2020-102607.abstract),
 similar to other areas of sciences (refs-education and
-ecology examples)(2). Below we discuss several practices to avoid in
-relation to missing data, and provide recommendations for exploring,
-visualising and reporting missingness.
+ecology examples)(2)**references needed**. Below we discuss several practices to avoid in
+relation to missingness, and provide recommendations for exploring,
+visualising and reporting missing data.
 

### 2020-09-05 paper/paper.Rmd
wording changes.

@@
 title: "Missing data: Current practice in football research and recommendations for improvement"
-authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
+authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
 output: 
@@
 
-2: University of New South Wales, Department of Statistics, School of
+2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
+
+3: University of New South Wales, Department of Statistics, School of
 Mathematics and Statistics, Sydney, Australia.
 
-3: Monash University, Department of Econometrics and Business Statistics,
+4: Monash University, Department of Econometrics and Business Statistics,
 Melbourne, Australia.
 
-4: Australian Centre of Excellence for Mathematical and Statistical
+5: Australian Centre of Excellence for Mathematical and Statistical
 Frontiers (ACEMS), Melbourne, Australia.
@@
 ## Missing data across the analysis pathway
+
+Missing data requires careful consideration across the entire analysis pathway. 
+Here we present several practices to avoid.
 

### 2020-09-05 paper/paper.Rmd
change wording in the discussion.

@@
 
-Missing data requires careful consideration across the entire analysis pathway. 
-Here we present several practices to avoid.
-
-*Unreported missing values*. While not always the case, even if
+Practice to avoid: *Unreported missing values.* While not always the case, even if
 not reported, missing data can be obvious. A recent paper examined the
@@
 
-*Mean imputation.* The default of most statistical software for
+Practice to avoid: *Mean imputation.* The default of most statistical software for
 handling missing data in modelling is listwise deletion. Listwise
@@
 
-*Not evaluating the effect, or choice, of imputation.* The
+Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The
 imputation method used to impute missing data has the potential to

### 2020-09-05 paper/paper.Rmd
change wording in the discussion.
edit wording in the discussion.

@@
 at least one variable (i.e., 'meters gained') was not available over the entire 2001-2016 period [@Young2019]**Check this reference**.
+**We might need to add a sentence about 'question mark' in the figures?**
 This was not noted anywhere in the paper, or in the supplementary
@@
 Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
-imputation method used to impute missing data has the potential to
-affect the outcome of a study. Therefore, it is important that authors
+method used to impute missing data has the potential to
+affect the outcome of a study. It is important that authors
 understand, and document, how imputation affected the goal of the
@@
 understanding the implications in using mean imputation, compared to
-K-nearest neighbour imputation -- the effect on parameter estimates
+K-nearest neighbour imputation--in terms of the effect on parameter estimates
 ('significant' versus 'not significant'), and the (un)certainty of the
 coefficients (i.e., less/more). When using imputation, authors need to
-perform a sensitivity analysis (ref circulation article-Nick). This is
+perform a sensitivity analysis (**ref circulation article-Nick**). This is
 discussed below.
@@
 recorded as 0 in this format, rather than NA. Other places missing data
-can arise include: bad joins when merging data without corresponding
+can arise include: through joins when merging data without corresponding
 values, surveys, an inability to collect a biological sample (e.g.,
-venous blood), equipment malfunction or failure, or or not being worn.
+venous blood), equipment malfunction, failure, or not being worn.
 
@@
 We recommend that researchers use missing data overview graphics (see
-Graphics section in [@Tierney2018], For example, overview plots [@Tierney2017].
+Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
 
@@
 To learn more about exploring missing values, we recommend the vignettes
-in naniar [@naniar], and the methods in (@Tierney2018)
+in naniar [@naniar], and the methods in (@Tierney2018).
 
@@
 such practice. Rather, we recommend values should generally be imputed. 
-
-Missing values can occur in the both the predictors (the "independent variables") and response variable (the "dependent variables").
-Care should be taken when imputing data, as typically only predictors should be imputed.
-For discussion on imputing the response value, see ...
-[@refs].
-(note: take a look again)
-
 We recommend avoiding imputation methods that impute the same value,
-such as mean or median imputation, and instead suggest using imputation
-methods such as: linear regression, k nearest neighbours, or expectation
-maximisation.
+such as the mean or median, and instead suggest using imputation
+methods such as: linear regression, k-nearest neighbours, or expectation
+maximisation. Missing values can occur in the both the predictors (the "independent variables") 
+and response variable (the "dependent variables"). Care should be taken 
+when imputing data, as typically only predictors should be imputed. 
+For discussion on imputing the response value, see ... [@refs]. **(note: take a look again)**
 
@@
 is no perfect, one-size-fits-all approach for imputing data. The goal is
-to generate similar values that might have been otherwise recorded -
-sometimes this means using a neighborhood approach of finding similar
-values, other times this means predicting responses using linear or tree
-models, and other times the most likely value might have been 0 or the
-last or first value carried forward. For detailed descriptions and a
+to generate similar values that might have been otherwise recorded.
+Sometimes this means using a neighborhood approach of finding similar
+values. Or it could mean predicting responses using linear or tree
+models. Other times the most likely value might have been 0 or the
+last or first value carried forward. For detailed descriptions, and a
 summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].
@@
 
-No matter the imputation method used, it is essential to compare results
+Irrespective of the imputation method used, it is essential to compare results
 of different missing data handling to understand how they may bias the
@@
 reporting missing data. Our intention is not to provide a 'cookbook'
-style approach to missing data, but rather broad recommendations to help
+style approach to missing data, but rather, broad recommendations to help
 researchers when writing the methods and results sections of a study,
@@
 
-More often than not, researchers will be faced with missing data. The
-reasons for the missingness, along with the decision of how missing data
-are treated, can have a profound effect on the validity and accuracy of
-a study's results. A survey of 2019 articles on football related topics
-showed that current practice is poor. Only about one in 20 studies
-reported missing data. To help address this issue, we have discussed
-recommendations for exploring and reporting missing data, to assist
-researchers during this important process in the analytical pipeline.
+Our survey of 2019 articles on football related topics
+showed that current practice of reporting missing data is poor, 
+with only about one in 20 studies reported missing data. To address this issue,
+we have provided recommendations for reporting and exploring missing data, to assist researchers 
+during this important process in the analytical pipeline.
 Research should consider these recommendations, and pay greater

### 2020-09-05 paper/paper.Rmd
change wording in the conclusion.

@@
 showed that current practice of reporting missing data is poor, 
-with only about one in 20 studies reported missing data. To address this issue,
-we have provided recommendations for reporting and exploring missing data, to assist researchers 
-during this important process in the analytical pipeline.
+with only about one in 20 studies reported missing data. 
+This could suggest that this aspect of data anlaysis receives little attention. 
+To address this issue and assist researchers,
+we have provided recommendations for reporting and exploring missing data.
 Research should consider these recommendations, and pay greater

### 2020-09-08 paper/supplement-2.Rmd
edit wording in supplement 2.

@@
 
-This review aimed to estimate the proportion of articles published on
-the topic area of football in 2019 that reported on missing data.
+This review aimed to estimate the proportion of articles that reported 
+missing data in football research.
 
@@
 
-Football related articles, published in 2019 were studied. Three major
-sports medicine databases were searched using the key terms and limits,
-on May 20, 2020. The databases were: SPORTDiscus, Embase and Cinahl. A
-list of search terms and complete search statements are provided in
+Football related articles published in 2019 were studied. Three major
+sports medicine databases were searched on May 20, 2020, using key terms and limits. 
+The databases were: SPORTDiscus, Embase and Cinahl. A
+list of the search terms and complete search statements are provided in
 Table \@ref(tab:search-table).

### 2020-09-08 paper/supplement-2.Rmd
edit supplement 2, including figure.

@@
 
-Included articles were: (a) published in 2019; (b) had an accessible
-English abstract; (c) had an accessible full text, in English; and (d)
-included quantitative data. Studies containing original data, including
-case studies (with time-series data) and n=1 trials, were considered.
+Included articles were: (a) published in 2019; (b) in English; 
+(c) had an accessible full-text; and (d) included quantitative data. 
+Studies containing original data, including case studies (with time-series data) 
+and n=1 trials, were considered.
 The search results were imported into EndNote (version 9.3.2).
 Duplicates were removed, and titles and abstracts were screened against
-the inclusion criteria. Of the articles identified for full text
+the inclusion criteria. Of the articles identified for full-text
 screening, 10% were sampled, at random, for survey. Because \~30% of the
@@
 (e.g., full-texts not in English, conference abstracts, qualitative
-studies), an additional 5% random sample was taken\--with the
-expectation that 30% of the articles from the additional sample would
-also be ineligible\--so to preserve a 10% screen of the total papers
-published on football related topics in 2019. Figure \@ref(fig:prisma) summarises the search process.
+studies), an additional 5% random sample was taken. It was expected that 
+30% of the additional 5% random sample would also be ineligible. With this loss,
+it was expected that 10% of the articles that remained after title and abstract 
+screening would undergo survey. 
+Figure \@ref(fig:prisma) summarises the search process.
 
@@
   # | After title and abstract screen     | 1,373                             |
-  # | Random sample full-text         | 199 (needed for 10% sample)       |
+  # | Random sample for full-text screen  | 199 (needed for 10% inclusion)    |
   # | Full-text included                  | 137                               |
   # | Full-text excluded                  | Full-text papers excluded (n=62)  |
-  # |                                 | with reasons:                     |
-  # |                                 | Qualitative study (n=)            |
-  # |                                 | Not in english (n=)               |
-  # |                                 | Unable to access full-text (n=)   |
 
 prisma(found = 3863,
-       found_other = 0,
+       found_other = F,
        no_dupes = 1954, 
@@
        full_text_exclusions = 62, 
-       qualitative = 0, 
+       qualitative = 137, 
        quantitative = 0,
-       width = 800, 
-       height = 800)
+       width = 800, height = 800,
+       labels = list(
+full_text_exclusions = "Full text articles excluded (n = 62),
+with reasons:
+Commentary/letter (n = 4),
+Conference abstract (n = 21),
+Full-text not available (n = 9),
+Infographic (n = 2),
+Not in English (n = 17),
+Qualitative study (n = 9)
+"))
+
 ```
@@
 
-The search returned a total of 3,863 articles (Figure X). Of the 199
+The search returned a total of 3,863 articles (Figure 1). Of the 199
 articles screened, 136 were deemed eligible, representing \~10% of total

### 2020-09-08 R/read-gs.R

@@ -3,6 +3,7 @@ library(googlesheets4)
 library(janitor)
 # insert your own email address
 gs4_auth(email = "nicholas.tierney@monash.edu") 
+#gs4_auth(email = "davidnicholas.borg@gmail.com") 
 sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
 gs_football_r1 <- read_sheet(sheet_url, 
                              range = "A2:O201",

### 2020-09-08 paper/supplement-2.Rmd
update data extraction table. update search results and conclusion.

@@
 statement was included, or if missing data were acknowledged in a
-figure/table; (b) where data were missing, whether data were considered
-MCAR, MAR or MNAR; (c) whether imputation was used, and the imputation
-method; and (d) whether articles that did not include a missing data
+figure/table; where data were missing, whether (b) data were considered
+MCAR, MAR or MNAR; (c) imputation was used; (d) the method used for imputation; 
+and (e) whether articles that did not include a missing data
 statement shared their data. Two authors independently extracted the
@@
 The search returned a total of 3,863 articles (Figure 1). Of the 199
-articles screened, 136 were deemed eligible, representing \~10% of total
-articles published on football related studies in 2019 (Figure X). The
-proportion of articles that included a missing data statement (or
-acknowledged missing data in a table/figure) was 5.9% \[95% confidence
+articles screened, 137 were deemed eligible, representing \~10% of total
+articles published on football related studies in 2019 (Figure 1). The
+proportion of articles that reported missing data was 5.9% \[95% confidence
 interval; X% to X%\] or 8/136. A breakdown of the extracted information
-is provided in Table XX.
+is provided in Table 2.
 
@@
 
-extractions <- tibble(
-  Imported = 3863,
-  `After duplicates removed` = 1954,
-  `After title and abstract screen` = 1373,
-  `Random sample full-text` = "199 for 10% sample",
-  `Full-text included` = 137,
-  `Full-text excluded` = 62
+extractions <- tibble( # This is Table 2
+  `Missing data reported` = "XX",
+  `Type of missingness reported (i.e., MCAR, MAR or MNAR)` = "XX",
+  `Missing data explored` = "XX",
+  `Missing observations imputed` = "XX",
+  `Imputation method` = "XX",
+  `Data shared` = "XX"
 )
@@
 
-A survey of 136 articles published in 2019, on football related topics
-showed that only one in 20 reported missing data. This result suggests
+A survey of 137 articles published in 2019 on football related topics
+showed that only one in 20 reported missing data, suggesting
 that missing data are rarely considered by sport science researchers.

### 2020-09-08 paper/supplement-1.Rmd
changes to example 1. stole Nick's idea from the main text of the paper.

@@
 
-For example, if a players' player ratings\--a statistic derived from
-many variables in the dataset\--are unavailable for one, or several,
-Australian Football matches, despite the individual participating in the
-match and the other variables being recorded.
+There are three broad categorisations describing why data can be 
+missing: Missing Completely at Random (MCAR), Missing at Random (MAR), 
+and Missing Not at Random (MNAR). These categorisations help describe 
+whether the missingness occurs by chance (MCAR), due to some observed 
+variable (MAR) or an unobserved variable (MNAR), with MCAR being 
+the least harmful, and MAR and MNAR biased. Here, we provide an example 
+of data MAR, MCAR and MNAR.
 
-MCAR refers to missingness that is unrelated to any observed or
-unobserved data. Missing Australian Football player ratings could be
-considered MCAR, if the missing ratings are not systematically related
-to other variables. MAR is where missingness is related to some observed
+Data MCAR could occur if an Australian Football data collector was taking 
+lunch during the third quarter of a match, resulting in data not being 
+recorded for this time period. Data is MCAR, because the missings are 
+not systematically related to other variables.
+
+Data may be considered MAR is where missingness is related to some observed
 data. Data MAR could arise if\... \^\^Rob's out on the full
 example\--players who are injured are less likely to fill out wellness
-surveys than those who are fully fit.\^\^. MNAR is when missingness is
+surveys than those who are fully fit.\^\^.
+
+MNAR is when missingness is
 related to something unobservable. For example, data MNAR could arise
@@
 any other variable that was collected (hair length was not recorded).
+
+# Older ideas
+
+For example, if a players' player ratings\--a statistic derived from
+many variables in the dataset\--are unavailable for one, or several,
+Australian Football matches, despite the individual participating in the
+match and the other variables being recorded.

### 2020-09-08 paper/supplement-1.Rmd
update missing data examples.

@@
 Data MCAR could occur if an Australian Football data collector was taking 
-lunch during the third quarter of a match, resulting in data not being 
+a lunch break during the third quarter of a match, resulting in data not being 
 recorded for this time period. Data is MCAR, because the missings are 
@@
 
-Data may be considered MAR is where missingness is related to some observed
-data. Data MAR could arise if\... \^\^Rob's out on the full
-example\--players who are injured are less likely to fill out wellness
-surveys than those who are fully fit.\^\^.
+Data MAR are more likely to occur than data MCAR. Data MAR could arise if
+a weighing scale used to measure players' pre- and post-match body mass 
+produced more missing values when placed on soft carpet, rather than hard tiles. 
+The scale may produce more missing values because of the soft surface. If 
+surface type was recorded, data could be considered MAR, because the 
+missingness is related to some observed data (i.e., surface), but not related 
+to the value of the variable with the missing values itself.
 

### 2020-09-08 paper/supplement-1.Rmd
change examples of missing data in supplement one.

@@
 
-Data MAR are more likely to occur than data MCAR. Data MAR could arise if
-a weighing scale used to measure players' pre- and post-match body mass 
-produced more missing values when placed on soft carpet, rather than hard tiles. 
-The scale may produce more missing values because of the soft surface. If 
-surface type was recorded, data could be considered MAR, because the 
-missingness is related to some observed data (i.e., surface), but not related 
-to the value of the variable with the missing values itself.
+Data MAR could arise if a weighing scale used to measure players' pre- and 
+post-match body mass produced more missing values when placed on soft carpet, 
+rather than hard tiles. If the scales produced more missing values because of 
+the softer surface, and surface type was a recorded variable, data would be MAR, 
+because the missingness is related to some observed data (i.e., surface), but 
+not related to the value of the variable with the missing values itself (i.e., body mass). 
+(ref: https://stefvanbuuren.name/fimd/sec-MCAR.html).
 
-MNAR is when missingness is
-related to something unobservable. For example, data MNAR could arise
-when, if studying doping, a player was to remove their head and body
-hair, to avoid providing a hair sample for testing
+Data MNAR is when the missings are related to something unobservable. 
+For example, data MNAR could arise when, if studying doping, a player 
+was to remove their head and body hair, to avoid providing a hair sample for testing
 [\[4\]](https://www.zotero.org/google-docs/?4Ildjf). In this case, data
@@
 any other variable that was collected (hair length was not recorded).
-
-# Older ideas
-
-For example, if a players' player ratings\--a statistic derived from
-many variables in the dataset\--are unavailable for one, or several,
-Australian Football matches, despite the individual participating in the
-match and the other variables being recorded.

### 2020-09-08 paper/supplement-1.Rmd
editing missing day examples, supplement one.

@@
 
-# missing data examples relevant to sports (MAR, MCAR, MNAR)
+# Missing data examples relevant to football
 
@@
 and Missing Not at Random (MNAR). These categorisations help describe 
-whether the missingness occurs by chance (MCAR), due to some observed 
-variable (MAR) or an unobserved variable (MNAR), with MCAR being 
+whether the missingness occurs by chance (i.e., MCAR), due to some observed 
+variable (i.e., MAR) or an unobserved variable (i.e., MNAR), with MCAR being 
 the least harmful, and MAR and MNAR biased. Here, we provide an example 
-of data MAR, MCAR and MNAR.
+of data MAR, MCAR and MNAR, relevant to football.
 
-Data MCAR could occur if an Australian Football data collector was taking 
+Data could be considered MCAR if an Australian Football data collector was taking 
 a lunch break during the third quarter of a match, resulting in data not being 
@@
 
-Data MAR could arise if a weighing scale used to measure players' pre- and 
-post-match body mass produced more missing values when placed on soft carpet, 
-rather than hard tiles. If the scales produced more missing values because of 
-the softer surface, and surface type was a recorded variable, data would be MAR, 
-because the missingness is related to some observed data (i.e., surface), but 
-not related to the value of the variable with the missing values itself (i.e., body mass). 
+Data MAR could arise if players' body mass measurements were taken with
+a weighing scale that produced more missing values when placed on soft carpet, rather than 
+hard tiles, in the team change room. If surface type was recorded, data would be MAR, 
+because the missingness is related the surface, but not related to the value of body mass itself. 
 (ref: https://stefvanbuuren.name/fimd/sec-MCAR.html).
 
-Data MNAR is when the missings are related to something unobservable. 
+**this example still needs work** Data MNAR is when the missings are related to something unobservable. 
 For example, data MNAR could arise when, if studying doping, a player 
 was to remove their head and body hair, to avoid providing a hair sample for testing 
-[\[4\]](https://www.zotero.org/google-docs/?4Ildjf). In this case, data
-is missing (the hair sample), and this information is not captured by
-any other variable that was collected (hair length was not recorded).
+(https://www.zotero.org/google-docs/?4Ildjf). In this case, data is missing (the hair sample), 
+and this information is not captured by any other variable that was collected 
+(hair length was not recorded).

### 2020-09-09 paper/supplement-1.Rmd
This should address #6

@@
 
-There are three broad categorisations describing why data can be 
-missing: Missing Completely at Random (MCAR), Missing at Random (MAR), 
-and Missing Not at Random (MNAR). These categorisations help describe 
-whether the missingness occurs by chance (i.e., MCAR), due to some observed 
-variable (i.e., MAR) or an unobserved variable (i.e., MNAR), with MCAR being 
-the least harmful, and MAR and MNAR biased. Here, we provide an example 
-of data MAR, MCAR and MNAR, relevant to football.
+One way to describe patterns of missing values is through Rubin's [@Rubin1976]
+three missing data mechanisms: Missing Completely at Random (MCAR), Missing at
+Random (MAR), and Missing Not at Random (MNAR). These mechanisms help describe
+whether the missingness occurs by chance (MCAR), due to some observed variable
+(MAR) or an unobserved variable (MNAR), with MCAR being the least harmful, and
+MAR and MNAR biased. Here, we provide an example of data MAR, MCAR and MNAR,
+relevant to football.
 
-Data could be considered MCAR if an Australian Football data collector was taking 
-a lunch break during the third quarter of a match, resulting in data not being 
-recorded for this time period. Data is MCAR, because the missings are 
-not systematically related to other variables.
+Data could be considered MCAR if an Australian Football data collector happened
+to take a lunch break during the third quarter of a match, resulting in data not
+being recorded for this time period. Data is MCAR, because the missings are not
+systematically related to other measured variables. It would also require that this lunch break related issue didn't systematically repeat itself. That is, it is completely random.
 
-Data MAR could arise if players' body mass measurements were taken with
-a weighing scale that produced more missing values when placed on soft carpet, rather than 
-hard tiles, in the team change room. If surface type was recorded, data would be MAR, 
-because the missingness is related the surface, but not related to the value of body mass itself. 
-(ref: https://stefvanbuuren.name/fimd/sec-MCAR.html).
+Data MAR could arise if players' body mass measurements were taken with a
+weighing scale that produced more missing values when placed on soft carpet,
+rather than hard tiles, in the team change room. If surface type was recorded,
+data would be MAR, because the missingness is related the surface, but not
+related to the value of body mass itself. (ref:
+https://stefvanbuuren.name/fimd/sec-MCAR.html).
 
-**this example still needs work** Data MNAR is when the missings are related to something unobservable. 
+Data MNAR is when the missings are related to something unobservable or something unobserved.
+The previous example of data MAR could be data MNAR if the surface type was not recorded in the dataset - it would then be unobserved.
+Data MNAR also arises when missing values are related to the values missing.
+This could take place if, for example, the scales used only go up to 110Kg, and then produce missing values.
+This is a systematic issue - it happens each time, and means that heavier players are not recorded.
+
+**Let me know what you think of ^^, I decided to avoid the doping example**
 For example, data MNAR could arise when, if studying doping, a player 

### 2020-09-09 paper/supplement-1.Rmd
add bibliography for supp1

@@
 output: html_document
+bibliography: supp1.bib
 ---
@@
 data would be MAR, because the missingness is related the surface, but not
-related to the value of body mass itself. (ref:
-https://stefvanbuuren.name/fimd/sec-MCAR.html).
+related to the value of body mass itself [@VanBuuren2012].
 
@@
 (hair length was not recorded).
+
+# References

### 2020-09-09 paper/supplement-2.Rmd
deleted a space.
add references for #12

@@
   # bookdown::pdf_document2
-bibliography: "supp-2.bib"
+bibliography: "supp2.bib"
 ---
@@
 list of the search terms and complete search statements are provided in
-Table \@ref(tab:search-table).
+Table 1 below
 
@@
              "Articles: 1,756",
-             ""),
+             "blank"),
   Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
@@
     )
+
 ```
+
+_Table 1. Caption here._
 

### 2020-09-10 paper/paper.Rmd
minor clean ups to text

@@
 Imputing data might feel wrong, as we are 'making up data'. The truth
-is, we (generally) can never know what the missing values were. The goal
+is, we can (generally) never know what the missing values were. The goal
 of imputing data is to make the best possible inference from the data.
@@
 variables with a high proportion of missing data (e.g., variables with
-the majority or more missing than not). With respect to a 'rule of
-thumb' for when data should be imputed, we deliberately do not suggest
-any specific threshold (e.g., impute values when >5% of data are
-missing otherwise delete rows with <5% missing) and caution against
-such practice. Rather, we recommend values should generally be imputed. 
-We recommend avoiding imputation methods that impute the same value,
-such as the mean or median, and instead suggest using imputation
-methods such as: linear regression, k-nearest neighbours, or expectation
-maximisation. Missing values can occur in the both the predictors (the "independent variables") 
+the majority or more missing than not). We deliberately do not suggest a 'rule of
+thumb' for the amount of missingness that should be imputed, and caution against
+looking for a specific threshold.
+Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (such as the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
+maximisation.
+
+Missing values can occur in the both the predictors (the "independent variables") 
 and response variable (the "dependent variables"). Care should be taken 

### 2020-09-10 paper/supplement-2.Rmd
minor clean ups to text

--- 2020-09-08 paper/supplement-2.Rmd
@@
   cols_width(
-    vars(Sportdiscus, Cinahl) ~ px(240),
-    vars(Embase) ~ px(300)
+    vars(Sportdiscus) ~ px(320),
+    vars(Cinahl) ~ px(280),
+    vars(Embase) ~ px(280)
     )
+
 ```
+
+_Table 1. Caption here._
 

### 2020-09-10 paper/supplement-3.Rmd
minor clean ups to text

@@
 
-```{r}
+```{r library}
 library(tidyverse)
@@
 library(naniar)
+```
 
+```{r}
 # Aim of this section:
@@
 
-d = fitzRoy::get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
+d <- get_afltables_stats(start_date = "2020-01-01", 
+                         end_date = "2020-09-04")
 
@@
 
-df <- fitzRoy::get_footywire_stats(10286:10290) # Get the fantasy scores
-View(df)
+df <- get_footywire_stats(10286:10290) # Get the fantasy scores
+
+# View(df)
 df #%>% View("footywire dataset")
+```
 
+```{r}
 # Simulate data from summarises of the above dataset
@@
       lambda = 5)
+```
 
+```{r}
 # Generate the y-variable
@@
 data %>% mutate(fanasty_score = 4.106*Kicks + 2.175*Handballs)
+```
 
+```{r}
 # Completed dataset
@@
 sort(unique(df$Team))
+```
 
+```{r}
 # Impute missing values
@@
 df_first_try <- df_first_try$amp
+```
 
+```{r}
 # Examples of presenting of missing values

### 2020-09-10 paper/supplement-1.Rmd
finalise supp 1

@@
 
-**Let me know what you think of ^^, I decided to avoid the doping example**
-For example, data MNAR could arise when, if studying doping, a player 
-was to remove their head and body hair, to avoid providing a hair sample for testing 
-(https://www.zotero.org/google-docs/?4Ildjf). In this case, data is missing (the hair sample), 
-and this information is not captured by any other variable that was collected 
-(hair length was not recorded).
-
 # References

### 2020-09-10 paper/supplement-2.Rmd
update data munging from work meeting

--- 2020-09-08 paper/supplement-2.Rmd
@@
 output: 
-  bookdown::html_document2
-  # bookdown::pdf_document2
-bibliography: "supp-2.bib"
+  bookdown::html_document2:
+    keep_md: true
+  bookdown::pdf_document2:
+    toc: false
+bibliography: "supp2.bib"
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, 
-                      message = FALSE)
+                      message = FALSE,
+                      fig.align = "center")
 ```
@@
 library(PRISMAstatement)
+library(DiagrammeRsvg)
 conflicted::conflict_prefer("select", "dplyr")
@@
 list of the search terms and complete search statements are provided in
-Table \@ref(tab:search-table).
+Table \@ref(tab:search-table-print) below.
 
 ```{r search-table}
-search_table <- tibble(
-  Sportdiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
-    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football",
-    "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed",
-    "Articles: 1,372",
-    "https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=sph&bquery=(((football)+OR+(soccer)+OR+(indoor+soccer)+OR+(futsal)+OR+(rugby+league)+OR+(rugby+union)+OR+(gaelic+football)+OR+(australian+rules+football)+OR+(AFL)+OR+(american+football)+OR+(gridiron+football)+OR+(touch+football)))+AND+(((exercise+science)+OR+(sports+science)+OR+(physiology)+OR+(biomechanics)+OR+(nutrition)+OR+(training)+OR+(testing)+OR+(sports+medicine)+OR+(performance+analysis)+OR+(performance)+OR+(sport+psychology)+OR+(coaching)))&cli0=DT1&clv0=201901-201912&cli1=AA1&clv1=Y&cli2=RV&clv2=Y&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site"),
+table_terms <- tibble(
+  SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
+    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
   Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
-              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py",
-             "#1 AND #2 AND [english]/lim AND 2019:py",
-             "Articles: 1,756",
-             ""),
+              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
   Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
-              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football",
-              "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed",
-              "Articles: 735",
-              "https://gateway.library.qut.edu.au/login?url=https://search.ebscohost.com/login.aspx?direct=true&db=c8h&bquery=(+exercise+science+OR+sports+science+OR+physiology+OR+biomechanics+OR+nutrition+OR+training+OR+testing+OR+sports+medicine+OR+performance+analysis+OR+performance+OR+sport+psychology+OR+coaching+)+AND+(+football+OR+soccer+OR+indoor+soccer+OR+futsal+OR+rugby+league+OR+rugby+union+OR+gaelic+football+OR+australian+rules+football+OR+AFL+OR+american+football+OR+gridiron+football+OR+touch+football+)&cli0=DT1&clv0=201901-201912&authtype=sso&custid=qut&type=1&searchMode=Standard&site=ehost-live&scope=site")
+              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")
 
 )
+```
 
-search_table %>% 
-  gt() %>% 
-  cols_width(
-    vars(Sportdiscus, Cinahl) ~ px(240),
-    vars(Embase) ~ px(300)
-    )
+```{r notes-to-add-for-index, echo = FALSE, eval = FALSE}
+# from sport discuss
+# "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed"
+# "Articles: 1,372"
+
+# from Embase
+# "#1 AND #2 AND [english]/lim AND 2019:py"
+# "Articles: 1,756"
+
+# from Cinahl
+# "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed"
+# "Articles: 735"
+```
+
+```{r search-table-print}
+table_terms %>% 
+  kable(caption = "Key search terms and complete search statements and limits for the databases, SPORTDiscuss, Embase, and Cinahl") %>% 
+  kable_styling("striped",
+                full_width = TRUE) 
 ```
@@
 
-```{r prisma, fig.cap = "Write the figure caption here"}
-  # | Imported                            | 3,863                             |
-  # | After duplicates removed            | 1,954                             |
-  # | After title and abstract screen     | 1,373                             |
-  # | Random sample for full-text screen  | 199 (needed for 10% inclusion)    |
-  # | Full-text included                  | 137                               |
-  # | Full-text excluded                  | Full-text papers excluded (n=62)  |
-
+```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 8}
 prisma(found = 3863,
@@
        quantitative = 0,
-       width = 800, height = 800,
+       width = 800, 
+       height = 800,
+       font_size = 16,
        labels = list(
-full_text_exclusions = "Full text articles excluded (n = 62),
+full_text_exclusions = "Full-text articles excluded (n = 62),
 with reasons:
@@
 
-The search returned a total of 3,863 articles (Figure 1). Of the 199
+The search returned a total of 3,863 articles (Figure \@ref(fig:prisma)). Of the 199
 articles screened, 137 were deemed eligible, representing \~10% of total
@@
 is provided in Table 2.
+
+```{r read-gs-data}
+
+```
 

### 2020-09-10 R/read-gs.R

@@ -7,10 +7,10 @@ gs4_auth(email = "nicholas.tierney@monash.edu")
 sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
 gs_football_r1 <- read_sheet(sheet_url, 
                              range = "A2:O201",
-                             col_types = "ciccccdcccccccc")
+                             col_types = "ciccccccccccccc")
 gs_football_r2_cols <- read_sheet(sheet_url, 
                                   range = "A2:H201",
-                                  col_types = "ciccccdc")
+                                  col_types = "cicccccc")
 gs_football_r2_results <- read_sheet(sheet_url, 
                                      range = "P2:V201",
                                      col_types = "ccccccc")
@@ -21,11 +21,23 @@ gs_football_r2 <- bind_cols(gs_football_r2_cols,
 gs_football <- bind_rows(reviewer_1 = gs_football_r1,
                          reviewer_2 = gs_football_r2,
                          .id = "reviewer") %>% 
-  clean_names()  %>% 
-  mutate(study_eligibility = as.logical(study_eligibility))
+  clean_names() %>% 
+  mutate(study_eligibility = as.logical(as.numeric(study_eligibility)),
+         exclusion_reason = replace_na(exclusion_reason, "Included"))
 
 library(visdat)
-vis_dat(gs_football, sort_type = FALSE)  
+vis_dat(gs_football_r1, sort_type = FALSE)
+vis_dat(gs_football_r2, sort_type = FALSE)
+vis_dat(gs_football, sort_type = FALSE)
+
+gs_football %>% 
+  filter(study_eligibility) %>% 
+  vis_dat(sort_type = FALSE)  
+
+gs_football_r1 %>% 
+  filter(study_eligibility == "1") %>% 
+  vis_dat(sort_type = FALSE)  
+
 
 # preliminary study
 gs_football %>% 

### 2020-09-10 paper/paper.Rmd
move supplement 3 into main text

@@
 
+# Aim of this section:
+# Simulate data from a real data set, insert missing
+# Mini study: predicting super coach scores
+
+d <- get_afltables_stats(start_date = "2020-01-01", 
+                         end_date = "2020-09-04")
+
+data <-
+  d %>% select(
+    ID,
+    Round,
+    First.name,
+    Surname,
+    Kicks,
+    Handballs,
+    Marks,
+    Hit.Outs,
+    Contested.Possessions,
+    Uncontested.Possessions
+  )
+#View(data)
+
+df <- get_footywire_stats(10286:10290) # Get the fantasy scores
+
+# View(df)
+df #%>% View("footywire dataset")
+
+# Generate the y-variable
+fit <- lm(AF ~ K + HB, data = df)
+summary(fit)
+data %>% mutate(fanasty_score = 4.106*Kicks + 2.175*Handballs)
+
+# Completed dataset
+# Drop a team
+No_Richmond <- df %>% filter(Team!="Richmond")
+sort(unique(No_Richmond$Team))
+sort(unique(df$Team))
+
+# Impute missing values
+df_first_try <- mice::ampute(data = df[,10:43])
+df_first_try <- df_first_try$amp
+
+# Examples of presenting of missing values
+naniar::vis_miss(df_first_try)
+naniar::vis_miss(df_first_try)
+naniar::gg_miss_fct(x = riskfactors, fct = marital)
+
+# Example write up of missing values
+
+```
+
+```{r}
 aq_vis_miss <- vis_miss(airquality)

### 2020-09-10 paper/paper.Rmd
merge supplement 3.

--- 2020-09-10 paper/paper.Rmd
@@
 title: "Missing data: Current practice in football research and recommendations for improvement"
-authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
+authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
 output: 
@@
 library(patchwork)
+library(fitzRoy)
+library(mice)
 ```
@@
 
-2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
-
-3: University of New South Wales, Department of Statistics, School of
+2: University of New South Wales, Department of Statistics, School of
 Mathematics and Statistics, Sydney, Australia.
 
-4: Monash University, Department of Econometrics and Business Statistics,
+3: Monash University, Department of Econometrics and Business Statistics,
 Melbourne, Australia.
 
-5: Australian Centre of Excellence for Mathematical and Statistical
+4: Australian Centre of Excellence for Mathematical and Statistical
 Frontiers (ACEMS), Melbourne, Australia.
@@
 (Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data. 
+The simualted data set contains ... **Add short description (1-2 sentences) here of the simulated data set**
+
+```{r}
+
+# Aim of this section:
+# Simulate data from a real data set, insert missing
+# 'Study aim': predicting super coach scores
+
+# Generate the x-variables
+d <- get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
+
+data <-
+  d %>% select(
+    ID,
+    Round,
+    First.name,
+    Surname,
+    Kicks,
+    Handballs,
+    Marks,
+    Hit.Outs,
+    Contested.Possessions,
+    Uncontested.Possessions
+  )
+
+```
+
+```{r}
+
+# Generate the y-variable
+df <- get_footywire_stats(10286:10290) # Get the fantasy scores
+fit <- lm(AF ~ K + HB, data = df)
+summary(fit)
+data %>% mutate(fanasty_score = 4.106*Kicks + 2.175*Handballs) # Need to work on this line
+
+```
+
+```{r}
+
+# Put the dataset together
+# Impute missing values
+data <- ampute(data = df[,10:43])
+data <- df_first_try$amp
+
+```
 
@@
 (aq_miss_var | aq_miss_case) / (aq_vis_miss) / (aq_fct)
-
-# can
 

### 2020-09-10 paper/paper.Rmd
edit data for missing data figure in the paper.

@@
 # Simulate data from a real data set, insert missing
-# 'Study aim': predicting super coach scores
+# 'Study aim': predicting hamstring strain injury, using number of sprints and sprint distance as x-variables
+# Add a sentence in the paper talking about this
+# I'm stuck at the momement how to simulate the occurance of an injury (0,1), but with only low incidence of 1's in the dataset
+# I'm also stuck how to simulate the number of sprints, but between say 50 and 200 sprints
 
@@
     First.name,
-    Surname,
-    Kicks,
-    Handballs,
-    Marks,
-    Hit.Outs,
-    Contested.Possessions,
-    Uncontested.Possessions
+    Surname
     )
@@
 
-# Generate the y-variable
-df <- get_footywire_stats(10286:10290) # Get the fantasy scores
-fit <- lm(AF ~ K + HB, data = df)
-summary(fit)
-data %>% mutate(fanasty_score = 4.106*Kicks + 2.175*Handballs) # Need to work on this line
+# Simulate number of sprints per game
+Observations = nrow(data)
+sprints <- sample.int(100, # numbers of sprints between 1 and 100
+           nrow(data),
+           replace = T) # observations
 
-```
+# Simulate incidence of hamstring strain injuries
+# sample.int(100, # numbers of sprints between 1 and 100
+#            nrow(data),
+#            replace = T) # observations
 
-```{r}
+data <- merge(data, sprints)
+
+# Impute missing values
+data <- ampute(data = df[,10:43])
+data <- df_first_try$amp
 
@@
 data <- ampute(data = df[,10:43])
-data <- df_first_try$amp
+data <- data$amp
 

### 2020-09-10 paper/paper.Rmd
change wording in simulation dataset section.

@@
 # 'Study aim': predicting hamstring strain injury, using number of sprints and sprint distance as x-variables
-# Add a sentence in the paper talking about this
+# We need to add a sentence in the paper talking about this
 # I'm stuck at the momement how to simulate the occurance of an injury (0,1), but with only low incidence of 1's in the dataset

### 2020-09-11 paper/paper.Rmd
edit discussion text.
edit simulation code.

@@
 ---
-title: "Missing data: Current practice in football research and recommendations for improvement"
+title: 'Missing data: Current practice in football research and recommendations for
+  improvement'
+bibliography: references.bib
+output:
+  word_document:
+    toc: yes
+  pdf_document:
+    toc: yes
+  html_document:
+    keep_md: yes
+    toc: yes
+    toc_float: yes
 authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
-output: 
-  html_document:
-    toc: true
-    toc_float: true
-    keep_md: true
-bibliography: references.bib
 ---
@@
 more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
-[@Sainani2015; @Nakagawa2008; Sterne2009].
+[@Sainani2015; @Nakagawa2008; @Sterne2009].
 
@@
 
-## Missing data across the analysis pathway
+## Practices to avoid across the analysis pathway
 
@@
 # Generate the x-variables
-d <- get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
-
-data <-
-  d %>% select(
-    ID,
-    Round,
-    First.name,
-    Surname
-    )
+# d <- get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
+# 
+# data <-
+#   d %>% select(
+#     ID,
+#     Round,
+#     First.name,
+#     Surname
+#     )
 
@@
 
-# Simulate number of sprints per game
-Observations = nrow(data)
-sprints <- sample.int(100, # numbers of sprints between 1 and 100
-           nrow(data),
-           replace = T) # observations
-
-# Simulate incidence of hamstring strain injuries
-# sample.int(100, # numbers of sprints between 1 and 100
+# # Simulate number of sprints per game
+# Observations = nrow(data)
+# sprints <- sample.int(100, # numbers of sprints between 1 and 100
 #            nrow(data),
 #            replace = T) # observations
-
-data <- merge(data, sprints)
-
-# Impute missing values
-data <- ampute(data = df[,10:43])
-data <- df_first_try$amp
-
-# Put the dataset together
-# Impute missing values
-data <- ampute(data = df[,10:43])
-data <- data$amp
+# 
+# # Simulate incidence of hamstring strain injuries
+# # sample.int(100, # numbers of sprints between 1 and 100
+# #            nrow(data),
+# #            replace = T) # observations
+# 
+# data <- merge(data, sprints)
+# 
+# # Impute missing values
+# data <- ampute(data = df[,10:43])
+# data <- df_first_try$amp
+# 
+# 
+# 
+# # Put the dataset together
+# # Impute missing values
+# data <- ampute(data = df[,10:43])
+# data <- data$amp
 

### 2020-09-11 paper/paper.Rmd
make changes to the simulated data set and missing data plots (a,b,c).

@@
 output:
-  word_document:
-    toc: yes
-  pdf_document:
-    toc: yes
   html_document:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
+  word_document:
+    toc: yes
 authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
@@
 library(mice)
+library(truncnorm)
+library(janitor)
 ```
@@
 
-(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data. 
-The simualted data set contains ... **Add short description (1-2 sentences) here of the simulated data set**
+(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data, in an simulated injury dataset, from Australian Rules football matches, played in August 2020.
+The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
+Researchers might be interested in investigating the relationship between number of sprints and total sprint distance with incidence of injury.
+**should we include this last sentence?**
 
@@
 # 'Study aim': predicting hamstring strain injury, using number of sprints and sprint distance as x-variables
-# We need to add a sentence in the paper talking about this
-# I'm stuck at the momement how to simulate the occurance of an injury (0,1), but with only low incidence of 1's in the dataset
-# I'm also stuck how to simulate the number of sprints, but between say 50 and 200 sprints
 
-# Generate the x-variables
-# d <- get_afltables_stats(start_date = "2020-01-01", end_date = "2020-09-04")
-# 
-# data <-
-#   d %>% select(
-#     ID,
-#     Round,
-#     First.name,
-#     Surname
-#     )
+# Simulate dataset
+d <- get_afltables_stats(start_date = "2020-08-01", end_date = "2020-09-01")
+
+data_afl <- d %>% select(ID, Round, Playing.for, Venue)
+
+# Simulate injury and sprint variables
+injury <- rbinom(n = nrow(d), 1, prob = 0.1)
+sprints <- rpois(n = nrow(d), lambda = 100)
+sprint_distance <- truncnorm::rtruncnorm(n = nrow(d), a = 0, b = Inf, mean = 2000, sd = 500) # Check if this is reasonable (mean 2 km per game)
+
+data <- cbind(injury, sprints, sprint_distance)
+
+# Impute missing values
+data_impute <- ampute(data = data[,1:3])
+data_with_missing <- data_impute$amp
+
+# Put the dataset together
+data_plot <- cbind(data_afl, data_with_missing)
+
+# Clean variable names
+data_plot <- data_plot %>% clean_names() %>% rename(team = playing_for)
 
@@
 
-```{r}
-
-# # Simulate number of sprints per game
-# Observations = nrow(data)
-# sprints <- sample.int(100, # numbers of sprints between 1 and 100
-#            nrow(data),
-#            replace = T) # observations
-# 
-# # Simulate incidence of hamstring strain injuries
-# # sample.int(100, # numbers of sprints between 1 and 100
-# #            nrow(data),
-# #            replace = T) # observations
-# 
-# data <- merge(data, sprints)
-# 
-# # Impute missing values
-# data <- ampute(data = df[,10:43])
-# data <- df_first_try$amp
-# 
-# 
-# 
-# # Put the dataset together
-# # Impute missing values
-# data <- ampute(data = df[,10:43])
-# data <- data$amp
-
-```
-
-```{r missing-overview, fig.cap = "Overviews of missing values in airquality data.
+```{r missing-overview, fig.cap = "Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
 Panel A shows...
 Panel B shows ...
-Panel C shows ...
-(simulate data from football study, provide write up in the supplementary materials)", fig.height = 10}
+Panel C shows ...", fig.height = 10}
 
-aq_vis_miss <- vis_miss(airquality)
+aq_vis_miss <- vis_miss(data_plot)
 
-aq_upset <- gg_miss_upset(airquality)
+aq_upset <- gg_miss_upset(data_plot)
 
-aq_fct <- gg_miss_fct(airquality, Month)
+aq_fct <- gg_miss_fct(data_plot, team)
 
-aq_miss_var <- gg_miss_var(airquality) + labs(title = "A")
-aq_miss_case <- gg_miss_case(airquality) + labs(title = "B")
-aq_vis_miss <- vis_miss(airquality) + labs(title = "C")
+aq_miss_var <- gg_miss_var(data_plot) + labs(title = "A")
+aq_miss_case <- gg_miss_case(data_plot) + labs(title = "B")
+aq_vis_miss <- vis_miss(data_plot) + labs(title = "C")
 

### 2020-09-11 paper/paper.Rmd
edit wording.

@@
 ```
+
+Authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
 
@@
 reporting missingness. Football related articles, published in 2019 were
-studied. A survey of 137 articles, sampled at random, was conducted to
+studied. A survey of 136 articles, sampled at random, was conducted to
 determine whether a missing data statement was included. As expected, the
@@
 
-## Practices to avoid across the analysis pathway
+## Practices to avoid
 

### 2020-09-19 paper/supplement-2.Rmd
update search method.

@@
 The databases were: SPORTDiscus, Embase and Cinahl. A
-list of the search terms and complete search statements are provided in
+list of the search terms has been provided in
 Table \@ref(tab:search-table-print) below.
@@
 
-```{r notes-to-add-for-index, echo = FALSE, eval = FALSE}
-# from sport discuss
-# "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed"
-# "Articles: 1,372"
-
-# from Embase
-# "#1 AND #2 AND [english]/lim AND 2019:py"
-# "Articles: 1,756"
-
-# from Cinahl
-# "Published Date: 20190101-20191231; English Abstract Available; Peer Reviewed"
-# "Articles: 735"
-```
-
 ```{r search-table-print}
 table_terms %>% 
-  kable(caption = "Key search terms and complete search statements and limits for the databases, SPORTDiscuss, Embase, and Cinahl") %>% 
+  kable(caption = "Key search terms and complete search statements for the databases, SPORTDiscuss, Embase, and Cinahl") %>% 
   kable_styling("striped",
@@
 intervals were calculated using the Clopper-Pearson method for the
-binomial distribution [@Clopper1934],via the 'binom' package [@binom] in R (version 4.0.2) [@rcore].
+binomial distribution [@Clopper1934], via the 'binom' package [@binom] in R (version 4.0.2) [@rcore].
 

### 2020-10-01 R/read-gs.R
coding session from 2020/10/01 on reading in the google sheet data

@@ -24,8 +24,22 @@ gs_football <- bind_rows(reviewer_1 = gs_football_r1,
   clean_names() %>% 
   mutate(study_eligibility = as.logical(as.numeric(study_eligibility)),
          exclusion_reason = replace_na(exclusion_reason, "Included")) %>% 
-  na_if("NA")
+  na_if("NA") %>% 
+  na_if("NA ") %>% 
+  na_if(" NA") %>% 
+  na_if(" ") %>% 
+  filter(study_eligibility) %>% 
+  mutate(was_data_shared = case_when(
+    was_data_shared == "At_request" ~ "0",
+    was_data_shared == "can request" ~ "0",
+    was_data_shared == "Public_data" ~ "0",
+    TRUE ~ was_data_shared
+  )) %>% 
+  mutate_at(.vars = vars(any_missing_data_statement:was_data_shared),
+            .funs = as.integer)
+  
 
+gs_football
 # data shared
 
 library(visdat)
@@ -36,12 +50,76 @@ vis_dat(gs_football_r2, sort_type = FALSE)
 vis_dat(gs_football, sort_type = FALSE)
 
 gs_football %>% 
-  filter(study_eligibility) %>% 
-  vis_dat(sort_type = FALSE)
+  select(any_missing_data_statement:was_data_shared) %>% 
+  map(unique)
+  
+gs_football %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement)  %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = n)
+
+gs_football %>% 
+  select(title,
+         reviewer,
+         any_missing_data_statement) %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = any_missing_data_statement) %>% 
+  filter(reviewer_1 != reviewer_2)
+
+pivot_var <- function(data, var){
+  data %>% 
+  select(title,
+         reviewer,
+         {{ var }}) %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = {{ var }})
+}
+
+identify_mismatch <- function(data, var){
+  pivot_var(data, {{ var }}) %>% 
+    filter(reviewer_1 != reviewer_2)
+}
+
+identify_mismatch(gs_football, any_missing_data_statement)
+identify_mismatch(gs_football, any_statement_about_mcar_mnar_mar)
+identify_mismatch(gs_football, were_missing_data_explored)
+identify_mismatch(gs_football, was_imputation_used)
[…]

### 2020-10-03 paper/paper.Rmd
edit wording in the conclusion.
update mismatch csv and main document edits

@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@PRISMA]. Football related
-articles published in 2019 were studied. Three major sports medicine
+football research, we conducted a systematic search [@PRISMA]. Articles on football topics (or involving footballers) published in 2019 were studied. Three major sports medicine
 databases (SPORTDiscus, Embase and Cinahl) were searched, using key
@@
 
-Of the 199 articles screened, 136 met the inclcusion criteria, representing
+Of the 199 articles screened, 136 met the inclusion criteria, representing
 \~10% of the articles found (see Supplement 2, Figure 1). The proportion of
@@
 we have provided recommendations for reporting and exploring missing data.
-Research should consider these recommendations, and pay greater
+Researchers should consider these recommendations, and pay greater
 attention to missing data and its influence on research results.

### 2020-10-04 paper/paper.Rmd
main text edits.

@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@PRISMA]. Articles on football topics (or involving footballers) published in 2019 were studied. Three major sports medicine
-databases (SPORTDiscus, Embase and Cinahl) were searched, using key
+football research, we conducted a systematic search [@PRISMA].
+Articles on football topics (or involving footballers) published in 2019 were studied.
+Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key
 terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
@@
 values were treated was also extracted (see Supplement 2). Two authors
-independently extracted the data (X% agreement). Findings are reported
+independently extracted the data (X% agreement). Results are reported
 as the proportion and 95% confidence interval.
@@
 Practice to avoid: *Unreported missing values.* While not always the case, even if
-not reported, missing data can be obvious. A recent paper examined the
+not reported, missing data can be obvious. For example, a recent study examined the
 relationship between match performance indicators and outcome, in
-Australian Football between the 2001 and 2016. Of the 91 team performance indicators included in the analysis,
-at least one variable (i.e., 'meters gained') was not available over the entire 2001-2016 period [@Young2019]**Check this reference**.
-**We might need to add a sentence about 'question mark' in the figures?**
+Australian Football between the 2001 and 2016, where at least one of the
+91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019]**Check this reference**.
 This was not noted anywhere in the paper, or in the supplementary
@@
 analysis. It is also difficult for a reader to evaluate how the
-missingness could bias the results. The absence of a
-missing data statement when data are missing is not unique to the study
-above [@Young2019]**Check this reference**. Authors should include a missing data statement,
+missingness could bias the results. Authors should include a missing data statement,
 irrespective of whether there is data missing, or not.
@@
 it is critical how these values are imputed, and that the imputation
-method is documented. One imputation method is imputing the mean value
-(e.g., participant or group average) [@Young2019]**Check this reference**. While imputing the mean
-preserves the study sample size (and point estimates), it also reduces
+method is documented. Authors sometimes choose to replace missing 
+observations with mean values (e.g., participant or group average) [@refs](https://doi.org/10.1016/j.jsams.2018.05.019, https://doi.org/10.1016/j.jsams.2008.01.009). 
+While imputing the mean preserves the study sample size (and point estimates), it also reduces
 the variance, can alter the relationship between the variable with
 missing observations and other variables, and can bias (underestimate)
-standard errors [@Scheffer2002].
-Smaller standard errors typically reduce *p*-values, which may lead to
+standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
 incorrect inference. Mean imputation should be avoided. An imputation
@@
 coefficients (i.e., less/more). When using imputation, authors need to
-perform a sensitivity analysis (**ref circulation article-Nick**). This is
+perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**). This is
 discussed below.
@@
 looking for a specific threshold.
-Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (such as the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
+Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
 maximisation.
@@
 when imputing data, as typically only predictors should be imputed. 
-For discussion on imputing the response value, see ... [@refs]. **(note: take a look again)**
+For discussion on imputing the response value, see ...
+[@refs].
+**(note: take a look again - this is the discusson of missing y values section - and possible reference relevant to RCT's is this: https://doi.org/10.1191%2F1740774504cn032oa)**
 
@@
 
-The current state of reporting missing data in the sports literature has room for improvement. In this section we discuss recommended practices for exploring and
+The current state of reporting missing data in football research has room for improvement. 
+In this section we discuss recommended practices for exploring and
 reporting missing data. Our intention is not to provide a 'cookbook'
 style approach to missing data, but rather, broad recommendations to help
-researchers when writing the methods and results sections of a study,
-and assist researchers when evaluating a study during the peer-review
-process.
+authors when writing the methods and results sections of a study,
+and assist researchers when evaluating a study during the peer-review process.
 
@@
 
-> *25 of the 280 values in the dataset (~8.92%) were missing.
+> *Twenty-five of the 280 values in the dataset (~8.92%) were missing.
 These were due to faults in GPS tracking as participants ran through sections of track covered by forest, a known issue with some GPS tracking.
@@
 showed that current practice of reporting missing data is poor, 
-with only about one in 20 studies reported missing data. 
-This could suggest that this aspect of data anlaysis receives little attention. 
+with only about one in 20 studies reporting missing data.
+We speculate that this result is representative of practice in the 
+wider sports science literature. The extremely low reporting of missing 
+data could suggest that this aspect of data anlaysis receives little attention. 
 To address this issue and assist researchers,

### 2020-10-04 paper/paper.Rmd
edit conclusion.

@@
 
-Our survey of 2019 articles on football related topics
+Our survey of articles on football related topics
 showed that current practice of reporting missing data is poor, 
-with only about one in 20 studies reporting missing data.
+with only about one in 20 studies reporting missingness.
 We speculate that this result is representative of practice in the 
 wider sports science literature. The extremely low reporting of missing 
-data could suggest that this aspect of data anlaysis receives little attention. 
+data could suggest that this aspect of anlaysis receives little attention. 
 To address this issue and assist researchers,

### 2020-10-04 paper/paper.Rmd
update main document and references.

@@
 
-Authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
+Authors:
+
+David N Borg(1)
+
+ORCiD: 0000-0002-0152-571X
+
+Twitter: elborgo9
+
+Robert Nguyen(2)
+
+ORCiD:
+
+Twitter:
+
+Nicholas J Tierney(3,4)
+
+ORCiD:
+
+Twitter:
 
@@
 **Data availability**
+The code to reproduce the results and plots in the paper can be accessed at: **Nick to host?**
 
@@
 to assume that this result is representative of current practice in the wider sports science literature. 
-The low proportion of papers reporting missingness could be explained by a lack of awareness and education on missing data (https://bjsm.bmj.com/content/early/2020/08/19/bjsports-2020-102607.abstract),
+The low proportion of papers reporting missingness could be explained by 
+a lack of awareness and education on missing data [@Sainani2020call], 
 similar to other areas of sciences (refs-education and
@@
 method is documented. Authors sometimes choose to replace missing 
-observations with mean values (e.g., participant or group average) [@refs](https://doi.org/10.1016/j.jsams.2018.05.019, https://doi.org/10.1016/j.jsams.2008.01.009). 
+observations with mean values (e.g., participant or group average) [@Sampson2018; @Cresswell2009]. 
 While imputing the mean preserves the study sample size (and point estimates), it also reduces
@@
 the data -- such as linear regression, or K nearest neighbours -- should
-be considered. This is discussed in the recommendations section.
+be considered.
 
@@
 coefficients (i.e., less/more). When using imputation, authors need to
-perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**). This is
-discussed below.
+perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**).
 
@@
 
-(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data, in an simulated injury dataset, from Australian Rules football matches, played in August 2020.
-The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
-Researchers might be interested in investigating the relationship between number of sprints and total sprint distance with incidence of injury.
-**should we include this last sentence?**
+(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
 
@@
 ```{r missing-overview, fig.cap = "Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
+The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
+Researchers might be interested in investigating the relationship between number of sprints and total sprint distance with incidence of injury.
 Panel A shows...
 Panel B shows ...
-Panel C shows ...", fig.height = 10}
+Panel C shows ...
+Panel D shows ...", fig.height = 10}
 

### 2020-10-08 R/agreement.R
code cleanup

@@ -1,0 +1,135 @@
+gs_football <- readr::read_rds("data/gs-football.rds")
+
+pivot_var <- function(data, var){
+  data %>% 
+    select(title,
+           reviewer,
+           {{ var }}) %>% 
+    pivot_wider(names_from = reviewer,
+                values_from = {{ var }})
+}
+
+identify_mismatch <- function(data, var){
+  pivot_var(data, {{ var }}) %>% 
+    filter(reviewer_1 != reviewer_2)
+}
+
+
+# calculate agreement
+library(irr)
+
+agree_var <- function(data, var){
+  statement_agree <- pivot_var(data, {{ var }}) %>% 
+    select(reviewer_1, reviewer_2) %>% 
+    irr::agree() 
+  
+  statement_agree$value
+}
+
+agree_var(gs_football, any_missing_data_statement)
+agree_var(gs_football, any_statement_about_mcar_mnar_mar)
+agree_var(gs_football, were_missing_data_explored)
+agree_var(gs_football, was_imputation_used)
+agree_var(gs_football, was_data_shared)
+
+# identify table of mismatches  
+gs_football %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement)  %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = n)
+
+
+# Now we bring the third and final review 
+gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")
+
+gs_mismatches %>% 
+  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
+               names_to = "reviewer",
+               values_to = "decision") %>% 
+  pivot_wider(names_from = variable,
+              values_from = decision)
+
+gs_football %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = )
+# any_missing_data_statement
+# list(
+#   any_missing_data_statement = identify_mismatch(gs_football,
+#                                                  any_missing_data_statement),
+#   any_statement_about_mcar_mnar_mar = identify_mismatch(gs_football,
+#                                                         any_statement_about_mcar_mnar_mar),
+#   
+#   were_missing_data_explored = identify_mismatch(gs_football, were_missing_data_explored),
+#   
+#   was_imputation_used = identify_mismatch(gs_football, was_imputation_used),
+#   
[…]

### 2020-10-08 R/exploratory-data-analysis.R

@@ -1,0 +1,13 @@
+
+# exploring the data
+
+library(visdat)
+library(naniar)
+vis_dat(gs_football_r1, sort_type = FALSE)
+vis_dat(gs_football_r2, sort_type = FALSE)
+vis_dat(gs_football, sort_type = FALSE)
+
+# check values are integers
+gs_football %>% 
+  select(any_missing_data_statement:was_data_shared) %>% 
+  map(unique)

### 2020-10-08 R/read-gs.R

@@ -1,8 +1,11 @@
 library(tidyverse)
 library(googlesheets4)
 library(janitor)
+
+read_football_gs <- function(email = "nicholas.tierney@monash.edu"){
 # insert your own email address
-gs4_auth(email = "nicholas.tierney@monash.edu") 
+# gs4_auth(email = "nicholas.tierney@monash.edu") 
+gs4_auth(email = email)
 #gs4_auth(email = "davidnicholas.borg@gmail.com") 
 sheet_url <- "https://docs.google.com/spreadsheets/d/1bxrBnQ8_ceX9Lz0-cRAB4VGBeW3SXluDQe-V84vDARM/edit"
 gs_football_r1 <- read_sheet(sheet_url, 
@@ -37,129 +40,5 @@ gs_football <- bind_rows(reviewer_1 = gs_football_r1,
   )) %>% 
   mutate_at(.vars = vars(any_missing_data_statement:was_data_shared),
             .funs = as.integer)
-  
-
-gs_football
-# data shared
-
-library(visdat)
-library(naniar)
-vis_dat(gs_football_r1, sort_type = FALSE)
-vis_dat(gs_football_r2, sort_type = FALSE)
-
-vis_dat(gs_football, sort_type = FALSE)
-
-gs_football %>% 
-  select(any_missing_data_statement:was_data_shared) %>% 
-  map(unique)
-  
-gs_football %>% 
-  group_by(reviewer) %>% 
-  count(any_missing_data_statement)  %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = n)
-
-gs_football %>% 
-  select(title,
-         reviewer,
-         any_missing_data_statement) %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = any_missing_data_statement) %>% 
-  filter(reviewer_1 != reviewer_2)
-
-pivot_var <- function(data, var){
-  data %>% 
-  select(title,
-         reviewer,
-         {{ var }}) %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = {{ var }})
+  gs_football
 }
-
-identify_mismatch <- function(data, var){
-  pivot_var(data, {{ var }}) %>% 
-    filter(reviewer_1 != reviewer_2)
-}
-
-identify_mismatch(gs_football, any_missing_data_statement)
-identify_mismatch(gs_football, any_statement_about_mcar_mnar_mar)
-identify_mismatch(gs_football, were_missing_data_explored)
[…]

### 2020-10-08 R/write-gs.R

@@ -1,0 +1,6 @@
+source("R/read-gs.R")
+
+gs_football <- read_football_gs()
+
+readr::write_rds(x = gs_football,
+                 path = "data/gs-football.rds")

### 2020-10-09 paper/paper.Rmd
small changes in paper

@@
 
-# Impute missing values
-data_impute <- ampute(data = data[,1:3])
-data_with_missing <- data_impute$amp
+# Ampute missing values
+data_ampute <- ampute(data = data[,1:3])
+# try and tinker with the amount of missings - so the variables have different
+# amounts of missingness
+# try and introduce some missingness mechanisms
+
+data_with_missing <- data_ampute$amp
 
@@
 
-aq_miss_var <- gg_miss_var(data_plot) + labs(title = "A")
-aq_miss_case <- gg_miss_case(data_plot) + labs(title = "B")
-aq_vis_miss <- vis_miss(data_plot) + labs(title = "C")
+gg_miss_scatter <- ggplot(data_plot,
+       aes(x = sprint_distance,
+           y = sprints)) +
+  geom_miss_point()
 
-(aq_miss_var | aq_miss_case) / (aq_vis_miss) / (aq_fct)
+aq_miss_var <- gg_miss_var(data_plot)
+
+cowplot::plot_grid(aq_vis_miss,
+                   # aq_upset,
+                   aq_fct,
+                   gg_miss_scatter,
+                   aq_miss_var, 
+                   align = "hv",
+                   labels = "AUTO")
 
@@
 Missing values can occur in the both the predictors (the "independent variables") 
-and response variable (the "dependent variables"). Care should be taken 
-when imputing data, as typically only predictors should be imputed. 
+and response variable (the "dependent variables"). 
+Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
+
+Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputed the outcome as they provide some sence of what the uncertainty is.
 For discussion on imputing the response value, see ...
 [@refs].
-**(note: take a look again - this is the discusson of missing y values section - and possible reference relevant to RCT's is this: https://doi.org/10.1191%2F1740774504cn032oa)**
+
+Missing values in the outcome can be dealt with specially for Randomised Control Trials, see - https://doi.org/10.1191%2F1740774504cn032oa.
 

### 2020-10-10 paper/paper.Rmd
format missing data overplot figure.

@@
 library(janitor)
+library(cowplot)
 ```
@@
 
-aq_vis_miss <- vis_miss(data_plot)
+aq_vis_miss <- vis_miss(data_plot) #+ 
+  #theme(text = element_text(size = 14))
 
-aq_upset <- gg_miss_upset(data_plot)
-
-aq_fct <- gg_miss_fct(data_plot, team)
+aq_fct <- gg_miss_fct(data_plot, team) + 
+  #theme(text = element_text(size = 14)) +
+  labs(x = "Team", y = "Variable")
 
@@
            y = sprints)) +
-  geom_miss_point()
+  geom_miss_point(size = 2) + 
+  theme_minimal() +
+  #theme(text = element_text(size = 14)) +
+  labs(x = "Total sprint distance (m)", y = "Number of sprints (count)", colour = "Missing")
 
@@
 
-cowplot::plot_grid(aq_vis_miss,
-                   # aq_upset,
-                   aq_fct,
-                   gg_miss_scatter,
-                   aq_miss_var, 
+# Grid plots
+cowplot::plot_grid(aq_vis_miss, aq_fct, gg_miss_scatter, aq_miss_var, 
+                   nrow = 2,
+                   ncol = 2,
                    align = "hv",
+                   axis = "tblr",
+                   label_size = 20,
                    labels = "AUTO")
-
 ```
@@
 # Supplement 2
-
-# Supplement 3

### 2020-10-15 doc/eda.Rmd
first pass at establishing drake plan

---
title: "Untitled Draft"
author: "Report Author"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Analysis

```{r}

# exploring the data

library(visdat)
library(naniar)
vis_dat(gs_football_r1, sort_type = FALSE)
vis_dat(gs_football_r2, sort_type = FALSE)
vis_dat(gs_football, sort_type = FALSE)

# check values are integers
gs_football %>% 
  select(any_missing_data_statement:was_data_shared) %>% 
  map(unique)
```

```{r}
gs_football <- readr::read_rds("data/gs-football.rds")

# calculate agreement
library(irr)

# any_missing_data_statement
# list(
#   any_missing_data_statement = identify_mismatch(gs_football,
#                                                  any_missing_data_statement),
#   any_statement_about_mcar_mnar_mar = identify_mismatch(gs_football,
#                                                         any_statement_about_mcar_mnar_mar),
#   
#   were_missing_data_explored = identify_mismatch(gs_football, were_missing_data_explored),
#   
#   was_imputation_used = identify_mismatch(gs_football, was_imputation_used),
#   
#   was_data_shared = identify_mismatch(gs_football, was_data_shared)
# ) %>%
#   bind_rows(.id = "variable") %>%
#   write_csv(path = "data/missing-data-mismatches.csv")

# Now we bring the third and final review 
gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")

gs_mismatches %>% 
  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
               names_to = "reviewer",
               values_to = "decision") %>% 
  pivot_wider(names_from = variable,
              values_from = decision)

gs_football %>% 
  pivot_wider(names_from = reviewer,
              values_from = )

# supp materials - agreement between Reviewer 1 and Reviewer 2

#the agreement was `r agree_var(gs_football, was_data_shared)`%

gs_football %>% 
  filter(study_eligibility) %>% 
  vis_miss(sort_miss = TRUE)

gs_football %>% 
  filter(study_eligibility) %>% 
  miss_var_summary()

# preliminary study
gs_football %>% 
  filter(study_eligibility) %>% 
  group_by(reviewer) %>% 
  count(any_missing_data_statement)

gs_football %>% 
  filter(study_eligibility) %>% 
  group_by(reviewer) %>% 
  count(any_missing_data_statement)

library(irr)
data(video)
video

gs_football %>% 
  filter(study_eligibility) %>% 
  filter(any_missing_data_statement == "NA")

gs_football %>% 
  filter(study_eligibility) %>% 
  select(reviewer, any_missing_data_statement) %>% 
  pivot_wider(names_from = reviewer,
              values_from = any_missing_data_statement,
              values_fn = list)

```

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>

### 2020-10-15 paper/paper.Rmd
first pass at establishing drake plan

--- 2020-10-09 paper/paper.Rmd
@@
 ---
-title: 'Missing data: Current practice in football research and recommendations for
-  improvement'
+title: 'Missing data: Current practice in football research and recommendations for improvement'
 bibliography: references.bib
@@
 library(janitor)
+library(drake)
+```
+
+```{r load-funs}
+pivot_var <- function(data, var){
+  data %>% 
+    select(title,
+           reviewer,
+           {{ var }}) %>% 
+    pivot_wider(names_from = reviewer,
+                values_from = {{ var }})
+}
+
+agree_var <- function(data, var){
+  statement_agree <- pivot_var(data, {{ var }}) %>% 
+    select(reviewer_1, reviewer_2) %>% 
+    irr::agree()
+
+  statement_agree$value
+}
+
+```
+
+```{r read-data}
+gs_football <- readr::read_rds(here::here("data",
+                           "gs-football.rds"))
+```
+
+```{r load-drake}
+# loadd()
 ```
@@
 
+```{r agree-var}
+agree_var(gs_football, any_missing_data_statement)
+agree_var(gs_football, any_statement_about_mcar_mnar_mar)
+agree_var(gs_football, were_missing_data_explored)
+agree_var(gs_football, was_imputation_used)
+agree_var(gs_football, was_data_shared)
+
+# identify table of mismatches  
+gs_football %>% 
+  group_by(reviewer) %>% 
+  count(any_missing_data_statement)  %>% 
+  pivot_wider(names_from = reviewer,
+              values_from = n)
+```
+
 Our intention was to survey, at random, 10% of the total articles found (after 
@@
 
-```{r}
+```{r afl-stats, cache = TRUE}
 

### 2020-10-15 R/agreement.R

@@ -1,135 +1,0 @@
-gs_football <- readr::read_rds("data/gs-football.rds")
-
-pivot_var <- function(data, var){
-  data %>% 
-    select(title,
-           reviewer,
-           {{ var }}) %>% 
-    pivot_wider(names_from = reviewer,
-                values_from = {{ var }})
-}
-
-identify_mismatch <- function(data, var){
-  pivot_var(data, {{ var }}) %>% 
-    filter(reviewer_1 != reviewer_2)
-}
-
-
-# calculate agreement
-library(irr)
-
-agree_var <- function(data, var){
-  statement_agree <- pivot_var(data, {{ var }}) %>% 
-    select(reviewer_1, reviewer_2) %>% 
-    irr::agree() 
-  
-  statement_agree$value
-}
-
-agree_var(gs_football, any_missing_data_statement)
-agree_var(gs_football, any_statement_about_mcar_mnar_mar)
-agree_var(gs_football, were_missing_data_explored)
-agree_var(gs_football, was_imputation_used)
-agree_var(gs_football, was_data_shared)
-
-# identify table of mismatches  
-gs_football %>% 
-  group_by(reviewer) %>% 
-  count(any_missing_data_statement)  %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = n)
-
-
-# Now we bring the third and final review 
-gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")
-
-gs_mismatches %>% 
-  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
-               names_to = "reviewer",
-               values_to = "decision") %>% 
-  pivot_wider(names_from = variable,
-              values_from = decision)
-
-gs_football %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = )
-# any_missing_data_statement
-# list(
-#   any_missing_data_statement = identify_mismatch(gs_football,
-#                                                  any_missing_data_statement),
-#   any_statement_about_mcar_mnar_mar = identify_mismatch(gs_football,
-#                                                         any_statement_about_mcar_mnar_mar),
-#   
-#   were_missing_data_explored = identify_mismatch(gs_football, were_missing_data_explored),
-#   
-#   was_imputation_used = identify_mismatch(gs_football, was_imputation_used),
-#   
[…]

### 2020-10-15 R/id-mismatch.R

@@ -1,0 +1,4 @@
+identify_mismatch <- function(data, var){
+  pivot_var(data, {{ var }}) %>% 
+    filter(reviewer_1 != reviewer_2)
+}

### 2020-10-15 R/pivot_var.R

@@ -1,0 +1,8 @@
+pivot_var <- function(data, var){
+  data %>% 
+    select(title,
+           reviewer,
+           {{ var }}) %>% 
+    pivot_wider(names_from = reviewer,
+                values_from = {{ var }})
+}

### 2020-10-15 R/read-markdown.R

@@ -1,3 +1,0 @@
-library(markdrive)
-
-gdoc_checkout("Missing data in sports and exercise science research")

### 2020-10-15 paper/paper.Rmd

--- 2020-10-10 paper/paper.Rmd
@@
 library(janitor)
+library(drake)
 library(cowplot)
+```
+
+```{r load-funs}
+pivot_var <- function(data, var){
+  data %>% 
+    select(title,
+           reviewer,
+           {{ var }}) %>% 
+    pivot_wider(names_from = reviewer,
+                values_from = {{ var }})
+}
+
+agree_var <- function(data, var){
+  statement_agree <- pivot_var(data, {{ var }}) %>% 
+    select(reviewer_1, reviewer_2) %>% 
+    irr::agree()
+
+  statement_agree$value
+}
+
+```
+
+```{r read-data}
+gs_football <- readr::read_rds(here::here("data",
+                           "gs-football.rds"))
+```
+
+```{r load-drake}
+# loadd()
 ```

### 2020-10-15 paper/paper.Rmd
start to bring in mismatches, fix spelling, fix the four panel figure

@@
 
-```{r read-data}
+```{r read-data, message = FALSE}
 gs_football <- readr::read_rds(here::here("data",
                            "gs-football.rds"))
+```
+
+```{r mismatch}
+# Now we bring the third and final review 
+gs_mismatches <- readr::read_csv(here::here("data",
+"missing-data-mismatches.csv"))
+
+gs_mismatch_pivot <- gs_mismatches %>% 
+  pivot_longer(cols = c(reviewer_1,reviewer_2, third_check),
+               names_to = "reviewer",
+               values_to = "decision") %>% 
+  pivot_wider(names_from = variable,
+              values_from = decision)
 ```
@@
 
-```{r count-words}
+```{r count-words, message = FALSE, warning = FALSE}
 wordcountaddin::text_stats()
@@
 There is no missing data because there is no data to be observed.
-Similiarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
+Similarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
 
@@
 explicit missings. Explicit values are missing, but recorded; whereas
-with implicit missinges, their presence is implied based on other
 information in the data. For example, in  Table \@ref(tab:implicit-missings), player Koenen has
@@
 # Simulate dataset
-d <- get_afltables_stats(start_date = "2020-08-01", end_date = "2020-09-01")
-
-data_afl <- d %>% select(ID, Round, Playing.for, Venue)
+# NOTE: fix this later in the drake plan
+# d <- get_afltables_stats(start_date = "2020-08-01", end_date = "2020-09-01")
+# write_rds(d, here::here("data", "afl-data.rds"))
+d <- readr::read_rds(here::here("data", "afl-data.rds"))
+afl_data <- d %>% select(ID, Round, Playing.for, Venue)
 
@@
 sprints <- rpois(n = nrow(d), lambda = 100)
-sprint_distance <- truncnorm::rtruncnorm(n = nrow(d), a = 0, b = Inf, mean = 2000, sd = 500) # Check if this is reasonable (mean 2 km per game)
+sprint_distance <- truncnorm::rtruncnorm(n = nrow(afl_data), 
+                                         a = 0, 
+                                         b = Inf, 
+                                         mean = 2000, 
+                                         sd = 500) # Check if this is reasonable (mean 2 km per game)
 
@@
 # Put the dataset together
-data_plot <- cbind(data_afl, data_with_missing)
+data_plot <- cbind(afl_data, data_with_missing)
 
@@
 data_plot <- data_plot %>% clean_names() %>% rename(team = playing_for)
+
+```
+
+```{r}
+data_plot_clean <- data_plot %>% 
+  mutate(team = recode(team, 
+                        "Gold Coast" = "GC",
+                        "Greater Western Sydney" = "GWS",
+                        "Fremantle" = "FRE",
+                        "Collingwood" = "COL",
+                        "Port Adelaide" = "PA",
+                        "Western Bulldogs" = "WB",
+                        "Richmond" = "RIC",
+                        "Brisbane Lions" = "BL",
+                        "Geelong" = "GEE",
+                        "North Melbourne" = "NM",
+                        "Adelaide" = "ADE",
+                        "Melbourne" = "MEL",
+                        "Sydney" = "SYD",
+                        "St Kilda" = "STK",
+                        "Essendon" = "ESS",
+                        "West Coast" = "WC",
+                        "Carlton" = "CAR",
+                        "Hawthorn" = "HAW")) %>% 
+  as_tibble() %>% 
+  rename(sprint_dist = sprint_distance)
 
@@
 
-aq_vis_miss <- vis_miss(data_plot) #+ 
+aq_vis_miss <- vis_miss(data_plot_clean) + labs(tag = "A")
   #theme(text = element_text(size = 14))
 
-aq_fct <- gg_miss_fct(data_plot, team) + 
+aq_fct <- gg_miss_fct(data_plot_clean, team) + 
   #theme(text = element_text(size = 14)) +
-  labs(x = "Team", y = "Variable")
+  labs(x = "Team", y = "Variable", tag = "B") +
+  theme(axis.text.x = element_text(size = 5))
 
-gg_miss_scatter <- ggplot(data_plot,
+gg_miss_scatter <- ggplot(data_plot_clean,
        aes(x = sprint_distance,
@@
   #theme(text = element_text(size = 14)) +
-  labs(x = "Total sprint distance (m)", y = "Number of sprints (count)", colour = "Missing")
+  labs(x = "Total sprint distance (m)", 
+       y = "Number of sprints (count)", 
+       colour = "",
+       tag = "C")
 
-aq_miss_var <- gg_miss_var(data_plot)
+aq_miss_var <- gg_miss_var(data_plot_clean) + labs(tag = "D")
+```
 
-# Grid plots
-cowplot::plot_grid(aq_vis_miss, aq_fct, gg_miss_scatter, aq_miss_var, 
-                   nrow = 2,
-                   ncol = 2,
-                   align = "hv",
-                   axis = "tblr",
-                   label_size = 20,
-                   labels = "AUTO")
+```{r patch-2, fig.height = 10}
+(aq_vis_miss + aq_fct) / (gg_miss_scatter + aq_miss_var) & 
+  theme(aspect.ratio = 1,
+        legend.position = "bottom")
 ```
@@
 
-Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputed the outcome as they provide some sence of what the uncertainty is.
+Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputed the outcome as they provide some sense of what the uncertainty is.
 For discussion on imputing the response value, see ...
@@
 wider sports science literature. The extremely low reporting of missing 
-data could suggest that this aspect of anlaysis receives little attention. 
+data could suggest that this aspect of analysis receives little attention. 
 To address this issue and assist researchers,

### 2020-10-15 doc/eda.Rmd
from today's meeting

@@
 gs_mismatches <- readr::read_csv("data/missing-data-mismatches.csv")
-
+library(tidyverse)
 gs_mismatches %>% 
@@
 
-gs_football %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = )
-
 # supp materials - agreement between Reviewer 1 and Reviewer 2
-
 #the agreement was `r agree_var(gs_football, was_data_shared)`%

### 2020-10-15 paper/paper.Rmd
from today's meeting

@@
 library(drake)
-library(cowplot)
+library(scales)
 ```
@@
 While there has been significant interest in modelling missing
-data [@refs], the exploration and reporting of missings have received
+data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received
 less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
@@
 To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@PRISMA].
+football research, we conducted a systematic search [@Moher2009].
 Articles on football topics (or involving footballers) published in 2019 were studied.
@@
 
-```{r agree-var}
-agree_var(gs_football, any_missing_data_statement)
-agree_var(gs_football, any_statement_about_mcar_mnar_mar)
-agree_var(gs_football, were_missing_data_explored)
-agree_var(gs_football, was_imputation_used)
-agree_var(gs_football, was_data_shared)
+```{r agreement}
 
-# identify table of mismatches  
-gs_football %>% 
-  group_by(reviewer) %>% 
-  count(any_missing_data_statement)  %>% 
-  pivot_wider(names_from = reviewer,
-              values_from = n)
+md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
+  percent(scale = 1,
+          accuracy = 0.1)
+
 ```
@@
 values were treated was also extracted (see Supplement 2). Two authors
-independently extracted the data (X% agreement). Results are reported
-as the proportion and 95% confidence interval.
+independently extracted the data (`r md_reported` agreement). Results are reported
+as the proportion and 95% confidence interval. (NOTE: need to get the final third reviewer part in)
 
@@
 
-Of the 199 articles screened, 136 met the inclusion criteria, representing
+```{r}
+# NOTE: update with the final third reviewer
+# need to update the gs_football
+library(binom)
+# !!comment to explain NAs
+successes <- sum(gs_football$any_missing_data_statement == 1, na.rm = TRUE)
+trials <- sum(!is.na(gs_football$any_missing_data_statement))
+missing_binom_conf <- binom.confint(x = successes, 
+              n = trials, 
+              conf.level = 0.95, 
+              methods = "exact")
+
+n_screened <- n_distinct(gs_football$key)
+miss_reported_pct <- percent(missing_binom_conf$mean, accuracy = 0.1)
+miss_reported_pct_lower <- percent(missing_binom_conf$lower, accuracy = 0.1)
+miss_reported_pct_upper <- percent(missing_binom_conf$upper, accuracy = 0.1)
+
+n_shared <- sum(gs_football$was_data_shared == 1)
+```
+
+```{r}
+# now of the articles that don't mention missingness
+gs_football_no_mention_sharing <- gs_football %>% 
+  filter(any_missing_data_statement == 0)
+
+n_no_mention_missings <- n_distinct(gs_football_no_mention_sharing$key)
+
+no_mention_successes <- sum(gs_football_no_mention_sharing$was_data_shared == 1, 
+                            na.rm = TRUE)
+no_mention_trials <- sum(!is.na(gs_football_no_mention_sharing$was_data_shared))
+n_mention_missing_binom_conf <- binom.confint(x = no_mention_successes, 
+                                              n = no_mention_trials, 
+                                              conf.level = 0.95, 
+                                              methods = "exact")
+
+percent_binom_conf <- function(binom_conf){
+  mean_pct <- percent(n_mention_missing_binom_conf$mean, accuracy = 0.1)
+lower_pct <- percent(n_mention_missing_binom_conf$lower, accuracy = 0.1)
+upper_pct <- percent(n_mention_missing_binom_conf$upper, accuracy = 0.1)
+
+return(c(mean_pct = mean_pct,
+         lower_pct = lower_pct,
+         upper_pct = upper_pct))
+}
+
+percent_binom_conf(n_mention_missing_binom_conf)
+
+```
+
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing
 \~10% of the articles found (see Supplement 2, Figure 1). The proportion of
-articles that reported missing data was 5.9% (95% confidence interval;
-X% to X%) or 8/136. Of the articles that did not mention missing data
-(128/136), 7.1% (3.3% to 13.1%) or 9/128 shared their dataset, making
+articles that reported missing data was `r miss_reported_pct` (95% confidence interval;
+`r miss_reported_pct_lower` to `r miss_reported_pct_upper`) or `r successes`/`r n_screened`.
+
+Of the articles that did not mention missing data `n_no_mention_missings` / `r n_screened`, 7.1% (3.3% to 13.1%) shared their dataset, making
 it difficult to determine whether there was any missingness. While the

### 2020-10-22 paper/paper.Rmd
edits

@@
 ---
-title: 'Missing data: Current practice in football research and recommendations for improvement'
+title: 'Missing Data: Current Practice in Football Research and Recommendations for Improvement'
 bibliography: references.bib
@@
 
-Twitter:
+Twitter: roberto_boberto
 
@@
 
-aq_vis_miss <- vis_miss(data_plot_clean) + labs(tag = "A")
+aq_vis_miss <- vis_miss(data_plot_clean) + labs(tag = "(A)")
   #theme(text = element_text(size = 14))
@@
   #theme(text = element_text(size = 14)) +
-  labs(x = "Team", y = "Variable", tag = "B") +
+  labs(x = "Team", y = "Variable", tag = "(B)") +
   theme(axis.text.x = element_text(size = 5))
@@
        colour = "",
-       tag = "C")
+       tag = "(C)")
 
-aq_miss_var <- gg_miss_var(data_plot_clean) + labs(tag = "D")
+aq_miss_var <- gg_miss_var(data_plot_clean) + labs(tag = "(D)")
 ```

### 2020-10-22 paper/paper.Rmd
fix from work session

--- 2020-10-15 paper/paper.Rmd
@@
 
-```{r read-data, message = FALSE}
+```{r read-data, message = FALSE, include = FALSE}
 gs_football <- readr::read_rds(here::here("data",
@@
 
-```{r mismatch}
+```{r mismatch, message = FALSE, include = FALSE}
 # Now we bring the third and final review 
@@
               values_from = decision)
+
+gs_mismatch_pivot %>% 
+  slice(1:3)
+
+gs_football
+```
+
+```{r finalise-matches}
+gs_football_final_matches <- gs_football %>% 
+  pivot_longer(cols = any_missing_data_statement:was_data_shared,
+               names_to = "variable",
+               values_to = "value") %>% 
+  pivot_wider(names_from = reviewer) %>% 
+  left_join({ 
+      select(gs_mismatches, 
+             -reviewer_1,
+             -reviewer_2)
+    }, 
+            by = c("variable", "title")) %>% 
+  # double check that we get our data back
+  # filter(reviewer_1 != reviewer_2)
+  mutate(third_check = as.integer(third_check),
+         final_check = case_when(
+    is.na(third_check) ~ reviewer_1,
+    reviewer_1 != reviewer_2 ~ third_check
+  ))  %>%  select(-c(reviewer_1, 
+            reviewer_2,
+            third_check,
+            notes)) %>% 
+  pivot_wider(names_from = variable,
+              values_from = final_check)
+  # filter(reviewer_1 != reviewer_2)
+
 ```
@@
 
-Twitter:
+Twitter: nj_tierney
 
-Affiliations:
+Affiliations: 0000-0003-1460-8722
 
@@
 Address: 
-Email:
+Email: nicholas.tierney@monash.edu
 
 **Data availability**
-The code to reproduce the results and plots in the paper can be accessed at: **Nick to host?**
+
+The code to reproduce the results and plots in the paper can be accessed at the following github repository:
 
@@
 wordcountaddin::text_stats()
+```
+
+```{r fun-pct-binom-conf}
+percent_binom_conf <- function(binom_conf){
+  mean_pct <- percent(binom_conf$mean, accuracy = 0.1)
+lower_pct <- percent(binom_conf$lower, accuracy = 0.1)
+upper_pct <- percent(binom_conf$upper, accuracy = 0.1)
+
+return(data.frame(mean_pct = mean_pct,
+                  lower_pct = lower_pct,
+                  upper_pct = upper_pct))
+}
+
+```
+
+```{r binom-ci-fun}
+# need to update the gs_football
+library(binom)
+
+football_binom_confint <- function(data, var){
+
+  var_vec <- data %>% pull({{ var }})
+  successes <- sum(var_vec, na.rm = TRUE)
+  trials <- sum(!is.na(var_vec))
+
+  binom.confint(x = successes, 
+              n = trials, 
+              conf.level = 0.95, 
+              methods = "exact")
+}
+```
+
+```{r binom-ci-reporting}
+n_screened <- nrow(gs_football_final_matches)
+n_shared <- sum(gs_football_final_matches$was_data_shared == 1)
+
+binom_ci_miss_reported <- football_binom_confint(gs_football_final_matches, 
+                                                 any_missing_data_statement) %>% 
+  percent_binom_conf()
+
+# now of the articles that don't mention missingness
+gs_football_final_matches_dont_report <- gs_football_final_matches %>% 
+  filter(any_missing_data_statement == 0)
+
+n_no_mention_missings <- nrow(gs_football_final_matches_dont_report)
+
+binom_ci_data_shared <- gs_football_final_matches_dont_report %>% 
+  football_binom_confint(var = was_data_shared) %>% 
+  percent_binom_conf()
+
 ```
@@
 reporting missingness. Football related articles, published in 2019 were
-studied. A survey of 136 articles, sampled at random, was conducted to
+studied. A survey of `r n_screened` articles, sampled at random, was conducted to
 determine whether a missing data statement was included. As expected, the
 proportion of studies in football research that included a missing data
-statement was low, at only X% (95% confidence interval; X% to X%); suggesting that 
+statement was low, at only `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that 
 missingness is seldom considered by researchers. It is suspected that this result 
@@
 values were treated was also extracted (see Supplement 2). Two authors
-independently extracted the data (`r md_reported` agreement). Results are reported
-as the proportion and 95% confidence interval. (NOTE: need to get the final third reviewer part in)
+independently extracted the data (`r md_reported` agreement). Disagreements were resolved by author DNB. Results are reported as the proportion and 95% confidence interval (CI).
 
@@
 
-```{r}
-# NOTE: update with the final third reviewer
-# need to update the gs_football
-library(binom)
-# !!comment to explain NAs
-successes <- sum(gs_football$any_missing_data_statement == 1, na.rm = TRUE)
-trials <- sum(!is.na(gs_football$any_missing_data_statement))
-missing_binom_conf <- binom.confint(x = successes, 
-              n = trials, 
-              conf.level = 0.95, 
-              methods = "exact")
-
-n_screened <- n_distinct(gs_football$key)
-miss_reported_pct <- percent(missing_binom_conf$mean, accuracy = 0.1)
-miss_reported_pct_lower <- percent(missing_binom_conf$lower, accuracy = 0.1)
-miss_reported_pct_upper <- percent(missing_binom_conf$upper, accuracy = 0.1)
-
-n_shared <- sum(gs_football$was_data_shared == 1)
-```
-
-```{r}
-# now of the articles that don't mention missingness
-gs_football_no_mention_sharing <- gs_football %>% 
-  filter(any_missing_data_statement == 0)
-
-n_no_mention_missings <- n_distinct(gs_football_no_mention_sharing$key)
-
-no_mention_successes <- sum(gs_football_no_mention_sharing$was_data_shared == 1, 
-                            na.rm = TRUE)
-no_mention_trials <- sum(!is.na(gs_football_no_mention_sharing$was_data_shared))
-n_mention_missing_binom_conf <- binom.confint(x = no_mention_successes, 
-                                              n = no_mention_trials, 
-                                              conf.level = 0.95, 
-                                              methods = "exact")
-
-percent_binom_conf <- function(binom_conf){
-  mean_pct <- percent(n_mention_missing_binom_conf$mean, accuracy = 0.1)
-lower_pct <- percent(n_mention_missing_binom_conf$lower, accuracy = 0.1)
-upper_pct <- percent(n_mention_missing_binom_conf$upper, accuracy = 0.1)
-
-return(c(mean_pct = mean_pct,
-         lower_pct = lower_pct,
-         upper_pct = upper_pct))
-}
-
-percent_binom_conf(n_mention_missing_binom_conf)
-
-```
-
 Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing
 \~10% of the articles found (see Supplement 2, Figure 1). The proportion of
-articles that reported missing data was `r miss_reported_pct` (95% confidence interval;
-`r miss_reported_pct_lower` to `r miss_reported_pct_upper`) or `r successes`/`r n_screened`.
+articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
 
-Of the articles that did not mention missing data `n_no_mention_missings` / `r n_screened`, 7.1% (3.3% to 13.1%) shared their dataset, making
+Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
+shared their dataset, making
 it difficult to determine whether there was any missingness. While the
@@
 
-Our survey of 136 articles published on football related topics found
-that only about one in 20 papers reported missing data. It is reasonable
+Our survey of `r n_screened` articles published on football related topics found
+that only one in 13 papers (`r binom_ci_miss_reported$mean_pct`) reported missing data. It is reasonable
 to assume that this result is representative of current practice in the wider sports science literature. 
@@
 
-```{r}
+```{r rename-afl-data}
 data_plot_clean <- data_plot %>% 

### 2020-10-30 paper/paper.Rmd
edit references.

--- 2020-10-22 paper/paper.Rmd
@@
 
-David N Borg(1)
+David N Borg(1,2)
 
@@
 
-Robert Nguyen(2)
+Robert Nguyen(3)
 
-ORCiD:
+ORCiD: 0000-0001-9373-3753
 
@@
 
-Nicholas J Tierney(3,4)
+Nicholas J Tierney(4,5)
 
@@
 Practice to avoid: *Unreported missing values.* While not always the case, even if
-not reported, missing data can be obvious. For example, a recent study examined the
+not reported, missing data can be obvious. For example, a study examined the
 relationship between match performance indicators and outcome, in
-Australian Football between the 2001 and 2016, where at least one of the
-91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019]**Check this reference**.
+Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
+91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019_mention_of_missing].
 This was not noted anywhere in the paper, or in the supplementary

### 2020-10-30 paper/paper.Rmd

--- 2020-10-22 paper/paper.Rmd
@@
 
-Robert Nguyen(2)
+Robert Nguyen(3)
 
-ORCiD:
+ORCiD: 0000-0001-9373-3753
 
-Twitter:
+Twitter: roberto_boberto
 
-Nicholas J Tierney(3,4)
+Nicholas J Tierney(4,5)
 

### 2020-10-30 packages.R
update with drake and capsule

@@ -12,3 +12,11 @@ library(dotenv)
 library(drake)
 library(rmarkdown)
 library(tidyverse)
+library(binom)
+library(kableExtra)
+library(knitr)
+library(gt)
+library(PRISMAstatement)
+library(DiagrammeRsvg)
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")

### 2020-10-30 paper/paper.Rmd
Nick orcid.

@@
 
-```{r load-funs}
-pivot_var <- function(data, var){
-  data %>% 
-    select(title,
-           reviewer,
-           {{ var }}) %>% 
-    pivot_wider(names_from = reviewer,
-                values_from = {{ var }})
-}
-
-agree_var <- function(data, var){
-  statement_agree <- pivot_var(data, {{ var }}) %>% 
-    select(reviewer_1, reviewer_2) %>% 
-    irr::agree()
-
-  statement_agree$value
-}
-
-```
-
-```{r read-data, message = FALSE, include = FALSE}
-gs_football <- readr::read_rds(here::here("data",
-                           "gs-football.rds"))
+```{r load-targets}
+loadd(gs_football)
+loadd(gs_mismatches)
+loadd(simulated_afl_data)
+loadd(gs_football_final_matches)
 ```
@@
 # Now we bring the third and final review 
-gs_mismatches <- readr::read_csv(here::here("data",
-"missing-data-mismatches.csv"))
-
 gs_mismatch_pivot <- gs_mismatches %>% 
@@
               values_from = decision)
-
-gs_mismatch_pivot %>% 
-  slice(1:3)
-
-gs_football
-```
-
-```{r finalise-matches}
-gs_football_final_matches <- gs_football %>% 
-  pivot_longer(cols = any_missing_data_statement:was_data_shared,
-               names_to = "variable",
-               values_to = "value") %>% 
-  pivot_wider(names_from = reviewer) %>% 
-  left_join({ 
-      select(gs_mismatches, 
-             -reviewer_1,
-             -reviewer_2)
-    }, 
-            by = c("variable", "title")) %>% 
-  # double check that we get our data back
-  # filter(reviewer_1 != reviewer_2)
-  mutate(third_check = as.integer(third_check),
-         final_check = case_when(
-    is.na(third_check) ~ reviewer_1,
-    reviewer_1 != reviewer_2 ~ third_check
-  ))  %>%  select(-c(reviewer_1, 
-            reviewer_2,
-            third_check,
-            notes)) %>% 
-  pivot_wider(names_from = variable,
-              values_from = final_check)
-  # filter(reviewer_1 != reviewer_2)
-
-```
-
-```{r load-drake}
-# loadd()
 ```
@@
 
-ORCiD:
+ORCiD: 0000-0003-1460-8722
 
@@
 
-Affiliations: 0000-0003-1460-8722
+Affiliations
 
@@
 
-```{r fun-pct-binom-conf}
-percent_binom_conf <- function(binom_conf){
-  mean_pct <- percent(binom_conf$mean, accuracy = 0.1)
-lower_pct <- percent(binom_conf$lower, accuracy = 0.1)
-upper_pct <- percent(binom_conf$upper, accuracy = 0.1)
-
-return(data.frame(mean_pct = mean_pct,
-                  lower_pct = lower_pct,
-                  upper_pct = upper_pct))
-}
-
-```
-
-```{r binom-ci-fun}
-# need to update the gs_football
-library(binom)
-
-football_binom_confint <- function(data, var){
-
-  var_vec <- data %>% pull({{ var }})
-  successes <- sum(var_vec, na.rm = TRUE)
-  trials <- sum(!is.na(var_vec))
-
-  binom.confint(x = successes, 
-              n = trials, 
-              conf.level = 0.95, 
-              methods = "exact")
-}
-```
-
 ```{r binom-ci-reporting}
@@
 
-binom_ci_miss_reported <- football_binom_confint(gs_football_final_matches, 
-                                                 any_missing_data_statement) %>% 
+binom_ci_miss_reported <- football_binom_confint(
+  gs_football_final_matches, 
+  any_missing_data_statement
+  ) %>% 
   percent_binom_conf()
@@
 
-```{r afl-stats, cache = TRUE}
-
-# Aim of this section:
-# Simulate data from a real data set, insert missing
-# 'Study aim': predicting hamstring strain injury, using number of sprints and sprint distance as x-variables
-
-# Simulate dataset
-# NOTE: fix this later in the drake plan
-# d <- get_afltables_stats(start_date = "2020-08-01", end_date = "2020-09-01")
-# write_rds(d, here::here("data", "afl-data.rds"))
-d <- readr::read_rds(here::here("data", "afl-data.rds"))
-afl_data <- d %>% select(ID, Round, Playing.for, Venue)
-
-# Simulate injury and sprint variables
-injury <- rbinom(n = nrow(d), 1, prob = 0.1)
-sprints <- rpois(n = nrow(d), lambda = 100)
-sprint_distance <- truncnorm::rtruncnorm(n = nrow(afl_data), 
-                                         a = 0, 
-                                         b = Inf, 
-                                         mean = 2000, 
-                                         sd = 500) # Check if this is reasonable (mean 2 km per game)
-
-data <- cbind(injury, sprints, sprint_distance)
-
-# Ampute missing values
-data_ampute <- ampute(data = data[,1:3])
-# try and tinker with the amount of missings - so the variables have different
-# amounts of missingness
-# try and introduce some missingness mechanisms
-
-data_with_missing <- data_ampute$amp
-
-# Put the dataset together
-data_plot <- cbind(afl_data, data_with_missing)
-
-# Clean variable names
-data_plot <- data_plot %>% clean_names() %>% rename(team = playing_for)
-
-```
-
-```{r rename-afl-data}
-data_plot_clean <- data_plot %>% 
-  mutate(team = recode(team, 
-                        "Gold Coast" = "GC",
-                        "Greater Western Sydney" = "GWS",
-                        "Fremantle" = "FRE",
-                        "Collingwood" = "COL",
-                        "Port Adelaide" = "PA",
-                        "Western Bulldogs" = "WB",
-                        "Richmond" = "RIC",
-                        "Brisbane Lions" = "BL",
-                        "Geelong" = "GEE",
-                        "North Melbourne" = "NM",
-                        "Adelaide" = "ADE",
-                        "Melbourne" = "MEL",
-                        "Sydney" = "SYD",
-                        "St Kilda" = "STK",
-                        "Essendon" = "ESS",
-                        "West Coast" = "WC",
-                        "Carlton" = "CAR",
-                        "Hawthorn" = "HAW")) %>% 
-  as_tibble() %>% 
-  rename(sprint_dist = sprint_distance)
-
-```
-
 ```{r missing-overview, fig.cap = "Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
@@
 
-aq_vis_miss <- vis_miss(data_plot_clean) + labs(tag = "(A)")
+aq_vis_miss <- vis_miss(simulated_afl_data) + labs(tag = "(A)")
   #theme(text = element_text(size = 14))
 
-aq_fct <- gg_miss_fct(data_plot_clean, team) + 
+aq_fct <- gg_miss_fct(simulated_afl_data, team) + 
   #theme(text = element_text(size = 14)) +
@@
 
-gg_miss_scatter <- ggplot(data_plot_clean,
-       aes(x = sprint_distance,
+gg_miss_scatter <- ggplot(simulated_afl_data,
+       aes(x = sprint_dist,
            y = sprints)) +
@@
 
-aq_miss_var <- gg_miss_var(data_plot_clean) + labs(tag = "(D)")
+aq_miss_var <- gg_miss_var(simulated_afl_data) + labs(tag = "(D)")
 ```

### 2020-10-30 paper/supplement-2.Rmd

@@
                       fig.align = "center")
-```
-
-```{r libraries, message = FALSE}
-library(tidyverse)
-library(kableExtra)
-library(knitr)
-library(conflicted)
-library(gt)
-library(PRISMAstatement)
-library(DiagrammeRsvg)
-conflicted::conflict_prefer("select", "dplyr")
-conflicted::conflict_prefer("filter", "dplyr")
 ```
@@
 
-```{r read-gs-data}
-
+```{r loadd-targets}
+loadd(gs_football)
+loadd(gs_mismatches)
+loadd(simulated_afl_data)
+loadd(gs_football_final_matches)
 ```

### 2020-10-30 paper/supplement-2.Rmd
add supplement 2 changes

@@
 
+```{r}
+gs_football_missings_stated  <- gs_football_final_matches %>% 
+  filter(any_missing_data_statement == 1)
+```
+
+```{r}
+gs_football_binom_statement <- gs_football_missings_stated %>% 
+  football_binom_confint(any_statement_about_mcar_mnar_mar)  %>% 
+  percent_binom_conf()
+
+gs_football_binom_explored <- gs_football_missings_stated %>% 
+  football_binom_confint(were_missing_data_explored)  %>% 
+  percent_binom_conf()
+
+gs_football_binom_imputation <- gs_football_missings_stated %>% 
+  football_binom_confint(was_imputation_used) %>% 
+  percent_binom_conf()
+
+```
+
+```{r}
+report_cis <- function(var){
+  ci_mean <- var$mean_pct
+  ci_upper <- var$lower_pct
+  ci_lower <- var$upper_pct
+
+  glue::glue("{var$mean_pct} \\
+           [{var$lower_pct} - \\
+           {var$upper_pct}]")
+}
+```
+
 ```{r extracted-info}
@@
 extractions <- tibble( # This is Table 2
-  `Missing data reported` = "XX",
-  `Type of missingness reported (i.e., MCAR, MAR or MNAR)` = "XX",
-  `Missing data explored` = "XX",
-  `Missing observations imputed` = "XX",
-  `Imputation method` = "XX",
-  `Data shared` = "XX"
+  `Type of missingness reported (i.e., MCAR, MAR or MNAR)` = report_cis(gs_football_binom_statement),
+  `Missing data explored` = report_cis(gs_football_binom_explored),
+  `Missing observations imputed` = report_cis(gs_football_binom_imputation)
 )
 
+# pivot_then
+extractions
+
+knitr::kable(extractions)
+
 ```
+
+There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data
 

### 2020-10-30 packages.R

@@ -15,6 +15,8 @@ library(tidyverse)
 library(binom)
 library(kableExtra)
 library(knitr)
+library(scales)
+library(googlesheets4)
 library(gt)
 library(PRISMAstatement)
 library(DiagrammeRsvg)

### 2020-11-05 paper/supplement-1.Rmd
edit wording.
grammar.
edit writing.

@@
 being recorded for this time period. Data is MCAR, because the missings are not
-systematically related to other measured variables. It would also require that this lunch break related issue didn't systematically repeat itself. That is, it is completely random.
+systematically related to other measured variables. For data to be MCAR, this lunch break related issue would need to occur completely at random, not systematically repeated each game.
 
@@
 
-Data MNAR is when the missings are related to something unobservable or something unobserved.
-The previous example of data MAR could be data MNAR if the surface type was not recorded in the dataset - it would then be unobserved.
+Data MNAR is when the missings are related to something unobservable or unobserved.
+The previous example of data MAR could be data MNAR if the surface type was not recorded in the dataset---it would then be unobserved.
 Data MNAR also arises when missing values are related to the values missing.
-This could take place if, for example, the scales used only go up to 110Kg, and then produce missing values.
-This is a systematic issue - it happens each time, and means that heavier players are not recorded.
+This could take place if, for example, the scales used only measure up to 110 Kg, and then produce missing values.
+This is a systematic issue---it happens each time, and means that heavier players are not recorded.
 

### 2020-11-05 paper/paper.Rmd
formatting.
mdash edits.

@@
 
-Affiliations
+Affiliations:
 
@@
 imputed. There is no universal approach to handling missing data. 
-Contextual factors--such as, the study design and objective, and
-pattern of missingness--determine how missings should be handled, on a
+Contextual factors---such as, the study design and objective, and
+pattern of missingness---determine how missings should be handled, on a
 case-by-case basis [@Sainani2015].
@@
 data are not being reported, it does suggest that this aspect of the
-analysis process receives little consideration--especially in contrast
+analysis process receives little consideration---especially in contrast
 to many of the papers that included (potentially meaningless) tests of
@@
 strategy that incorporates information from other related variables in
-the data -- such as linear regression, or K nearest neighbours -- should
+the data---such as linear regression, or K nearest neighbours---should
 be considered.
@@
 understanding the implications in using mean imputation, compared to
-K-nearest neighbour imputation--in terms of the effect on parameter estimates
+K-nearest neighbour imputation---in terms of the effect on parameter estimates
 ('significant' versus 'not significant'), and the (un)certainty of the
@@
 
-1.  Describe screening procedures (define or example -- think about screened out?).
+1.  Describe screening procedures **(define or example---think about screened out?)**.
 

### 2020-11-05 paper/supplement-2.Rmd
update plan for supplement 2

--- 2020-10-30 paper/supplement-2.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, 
+                      cache = FALSE,
                       message = FALSE,
                       fig.align = "center")
+```
+
+```{r loadd-targets}
+loadd(gs_football)
+loadd(gs_mismatches)
+loadd(simulated_afl_data)
+loadd(gs_football_final_matches)
+loadd(binom_ci_miss_reported)
+loadd(n_no_mention_missings)
+loadd(binom_ci_data_shared)
+```
+
+```{r check}
+loadd(n_screened)
+print(n_screened)
 ```
@@
 
-The search returned a total of 3,863 articles (Figure \@ref(fig:prisma)). Of the 199
-articles screened, 137 were deemed eligible, representing \~10% of total
-articles published on football related studies in 2019 (Figure 1). The
-proportion of articles that reported missing data was 5.9% \[95% confidence
-interval; X% to X%\] or 8/136. A breakdown of the extracted information
-is provided in Table 2.
+The search returned a total of 3,863 articles (Figure \@ref(fig:prisma)).
 
-```{r loadd-targets}
-loadd(gs_football)
-loadd(gs_mismatches)
-loadd(simulated_afl_data)
-loadd(gs_football_final_matches)
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1).
+The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
+
+Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
+shared their dataset, making it difficult to determine whether there was any missingness.
+While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration--especially in contrast to many of the papers that included (potentially meaningless) tests of normality (e.g., Kruskall-Wallis test).
+
+A breakdown of the extracted information is provided in Table 2.
+
+```{r filter-states}
+gs_football_missings_stated  <- gs_football_final_matches %>% 
+  filter(any_missing_data_statement == 1)
+```
+
+```{r binom-calcs}
+gs_football_binom_statement <- gs_football_missings_stated %>% 
+  football_binom_confint(any_statement_about_mcar_mnar_mar)  %>% 
+  percent_binom_conf()
+
+gs_football_binom_explored <- gs_football_missings_stated %>% 
+  football_binom_confint(were_missing_data_explored)  %>% 
+  percent_binom_conf()
+
+gs_football_binom_imputation <- gs_football_missings_stated %>% 
+  football_binom_confint(was_imputation_used) %>% 
+  percent_binom_conf()
 ```
@@
 extractions <- tibble( # This is Table 2
-  `Missing data reported` = "XX",
-  `Type of missingness reported (i.e., MCAR, MAR or MNAR)` = "XX",
-  `Missing data explored` = "XX",
-  `Missing observations imputed` = "XX",
-  `Imputation method` = "XX",
-  `Data shared` = "XX"
+  `Type (i.e., MCAR, MAR or MNAR)` = report_ci(gs_football_binom_statement),
+  `Missing data were explored` = report_ci(gs_football_binom_explored),
+  `Missing data were imputed` = report_ci(gs_football_binom_imputation)
 )
 
+# pivot_then
+extractions %>% 
+  pivot_longer(cols = everything(),
+               names_to = "Reporting of Missingness",
+               values_to = "Mean and 95% CI") %>% 
+  # turns it into a table
+  knitr::kable() %>% 
+  kable_styling("striped",
+                full_width = TRUE) 
 ```
+
+There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data.
 

### 2020-11-05 R/calc_n_no_mention_missings.R

@@ -1,0 +1,19 @@
+##' .. content for \description{} (no empty lines) ..
+##'
+##' .. content for \details{} ..
+##'
+##' @title
+##' @param gs_football_final_matches
+##' @return
+##' @author Nicholas Tierney
+##' @export
+calc_n_no_mention_missings <- function(gs_football_final_matches) {
+
+  gs_football_final_matches_dont_report <- gs_football_final_matches %>% 
+    filter(any_missing_data_statement == 0)
+  
+  n_no_mention_missings <- nrow(gs_football_final_matches_dont_report)
+  
+  n_no_mention_missings
+
+}

### 2020-11-05 R/plan.R

@@ -7,6 +7,12 @@ the_plan <-
   simulated_afl_data = generate_afl_data(),
   gs_football_final_matches = finalise_reviews(gs_mismatches, gs_football),
   
+  n_screened = nrow(gs_football_final_matches),
+  n_shared = sum(gs_football_final_matches$was_data_shared == 1),
+  binom_ci_miss_reported = calc_binom_ci_miss_reported(gs_football_final_matches),
+  n_no_mention_missings = calc_n_no_mention_missings(gs_football_final_matches),
+  binom_ci_data_shared = calc_binom_ci_data_shared(gs_football_final_matches),
+  
    paper = target(
      command = {
        rmarkdown::render(knitr_in("paper/paper.Rmd"))

### 2020-11-06 paper/supplement-2.Rmd
update figures and try to fix tables

@@
 output: 
+  bookdown::pdf_document2:
+    toc: false
   bookdown::html_document2:
     keep_md: true
-  bookdown::pdf_document2:
-    toc: false
 bibliography: "supp2.bib"
@@
 
+```{r}
+library(kableExtra)
+options(kableExtra.latex.load_packages = FALSE)
+```
+
 ```{r loadd-targets}
@@
 loadd(binom_ci_data_shared)
+loadd(write_prisma_figure)
 ```
@@
 table_terms %>% 
-  kable(caption = "Key search terms and complete search statements for the databases, SPORTDiscuss, Embase, and Cinahl") %>% 
+  kbl(caption = "Key search terms and complete search statements for the databases, SPORTDiscuss, Embase, and Cinahl",
+      booktabs = TRUE) %>% 
   kable_styling("striped",
@@
 ```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 8}
-prisma(found = 3863,
-       found_other = F,
-       no_dupes = 1954, 
-       screened = 1373, # (1954 - 1373) 
-       screen_exclusions = 581, 
-       full_text = 199,
-       full_text_exclusions = 62, 
-       qualitative = 137, 
-       quantitative = 0,
-       width = 800, 
-       height = 800,
-       font_size = 16,
-       labels = list(
-full_text_exclusions = "Full-text articles excluded (n = 62),
-with reasons:
-Commentary/letter (n = 4),
-Conference abstract (n = 21),
-Full-text not available (n = 9),
-Infographic (n = 2),
-Not in English (n = 17),
-Qualitative study (n = 9)
-"))
-
+knitr::include_graphics("figures/prisma-figure.pdf")
 ```
@@
 
-The search returned a total of 3,863 articles (Figure \@ref(fig:prisma)).
+The search returned a total of 3,863 articles.
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)).
+The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
 
-Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1).
-The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-
-Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
+Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
 shared their dataset, making it difficult to determine whether there was any missingness.
@@
 
-A breakdown of the extracted information is provided in Table 2.
+A breakdown of the extracted information is provided in Table \@ref(tab:extracted-info).
 
@@
   # turns it into a table
-  knitr::kable() %>% 
+  kbl(booktabs = TRUE,
+      caption = "Reporting of missingness, many papers did not explore missings in a meaningful way, and there was the same number of papers who imputed values as reported MCAR, MAR, or MNAR.") %>% 
   kable_styling("striped",
@@
 
-A survey of 137 articles published in 2019 on football related topics
-showed that only one in 20 reported missing data, suggesting
+A survey of `r n_screened` articles published in 2019 on football related topics
+showed that only one in nine reported missing data, suggesting
 that missing data are rarely considered by sport science researchers.

### 2020-11-06 R/plan.R

@@ -12,7 +12,9 @@ the_plan <-
   binom_ci_miss_reported = calc_binom_ci_miss_reported(gs_football_final_matches),
   n_no_mention_missings = calc_n_no_mention_missings(gs_football_final_matches),
   binom_ci_data_shared = calc_binom_ci_data_shared(gs_football_final_matches),
-  
+  prisma_figure = create_prisma_figure(),
+  write_prisma_figure = PRISMAstatement:::prisma_pdf(x = prisma_figure, 
+                                                     filename = "paper/figures/prisma-figure.pdf"),
    paper = target(
      command = {
        rmarkdown::render(knitr_in("paper/paper.Rmd"))

### 2020-11-06 paper/paper.Rmd

@@
 output:
+  bookdown::pdf_document2:
+    toc: no
   html_document:
@@
     toc_float: yes
-  pdf_document:
-    toc: yes
   word_document:
@@
 authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
+csl: sport-medicine.csl
 ---
@@
 loadd(gs_football_final_matches)
+# need to put these into the plan
+loadd(n_screened)
+loadd(n_shared)
+loadd(binom_ci_miss_reported)
+loadd(n_no_mention_missings)
+loadd(binom_ci_data_shared)
 ```
@@
 
-1: Griffith University, Menzies Health Institute Queensland, The Hopkins
-Centre, Brisbane, Australia.
+1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.
 
@@
 
-3: University of New South Wales, Department of Statistics, School of
-Mathematics and Statistics, Sydney, Australia.
+3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
 
-4: Monash University, Department of Econometrics and Business Statistics,
-Melbourne, Australia.
+4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
 
-5: Australian Centre of Excellence for Mathematical and Statistical
-Frontiers (ACEMS), Melbourne, Australia.
+5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.
 
@@
 
-Nicholas J Tierney
-Monash University, Department of Econometrics and Business Statistics,
-Melbourne, Australia.
-Address: 
-Email: nicholas.tierney@monash.edu
+Nicholas J Tierney Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
+Address: Email: [nicholas.tierney\@monash.edu](mailto:nicholas.tierney@monash.edu){.email}
 
@@
 
-**Word count**
+<!-- **Word count** -->
 
-```{r count-words, message = FALSE, warning = FALSE}
+```{r count-words, message = FALSE, warning = FALSE, eval = FALSE}
 wordcountaddin::text_stats()
@@
 
-```{r binom-ci-reporting}
-n_screened <- nrow(gs_football_final_matches)
-n_shared <- sum(gs_football_final_matches$was_data_shared == 1)
+# Abstract {.unnumbered}
 
-binom_ci_miss_reported <- football_binom_confint(
-  gs_football_final_matches, 
-  any_missing_data_statement
-  ) %>% 
-  percent_binom_conf()
-
-# now of the articles that don't mention missingness
-gs_football_final_matches_dont_report <- gs_football_final_matches %>% 
-  filter(any_missing_data_statement == 0)
-
-n_no_mention_missings <- nrow(gs_football_final_matches_dont_report)
-
-binom_ci_data_shared <- gs_football_final_matches_dont_report %>% 
-  football_binom_confint(var = was_data_shared) %>% 
-  percent_binom_conf()
-
-```
-
-# Abstract {-}
-
-Missing data are often unavoidable. The reason values go missing, along
-with decisions made of how missing data are handled (deleted or
-imputed), can have a profound effect on the validity, accuracy and usability 
-of a study's results. In this article, we aimed to: estimate
-the proportion of studies in football research that included a missing
-data statement, highlight several practices to avoid in relation to
-missing data, and provide recommendations for exploring, visualising and
-reporting missingness. Football related articles, published in 2019 were
-studied. A survey of `r n_screened` articles, sampled at random, was conducted to
-determine whether a missing data statement was included. As expected, the
-proportion of studies in football research that included a missing data
-statement was low, at only `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that 
-missingness is seldom considered by researchers. It is suspected that this result 
-is consistent with practice in the wider sports science literature. We recommend 
-researchers describe the number and percentage of missing
-values, including when there are no missing values. Exploratory analysis
-should be conducted to explore missing values, and visualisations
-describing missingness overall should be provided in the paper, or at
-least supplementary materials. Missing values should almost always be
-imputed, and imputation methods should be explored to ensure they are
-appropriately representative.
+Missing data are often unavoidable.
+The reason values go missing, along with decisions made of how missing data are handled (deleted or imputed), can have a profound effect on the validity, accuracy and usability of a study's results.
+In this article, we aimed to: estimate the proportion of studies in football research that included a missing data statement, highlight several practices to avoid in relation to missing data, and provide recommendations for exploring, visualising and reporting missingness.
+Football related articles, published in 2019 were studied.
+A survey of `r n_screened` articles, sampled at random, was conducted to determine whether a missing data statement was included.
+As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers.
+It is suspected that this result is consistent with practice in the wider sports science literature.
+We recommend researchers describe the number and percentage of missing values, including when there are no missing values.
+Exploratory analysis should be conducted to explore missing values, and visualisations describing missingness overall should be provided in the paper, or at least supplementary materials.
+Missing values should almost always be imputed, and imputation methods should be explored to ensure they are appropriately representative.
 
@@
 # Introduction {#intro}
+
+Missing data are values that should have been observed, but were not.
+Since the values are unobserved, this can mean the population of interest might not be properly sampled, inducing bias into a study's results.
+Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results [@Sainani2015; @Nakagawa2008; @Sterne2009].
+Data can go missing many ways.
+For example, accidentally skipping a survey question, equipment failure, or intentionally not recording values.
+Before analysis, missing data must be handled, with values typically deleted or imputed.
+There is no universal approach to handling missing data.
+Contextual factors--such as, the study design and objective, and pattern of missingness--determine how missings should be handled, on a case-by-case basis [@Sainani2015].
+
+For clarity, we describe an example of missingness.
+An Australian Football data collector was taking lunch during the third quarter, so no data was recorded for this time period.
+This would be considered missing data.
+Compare this to the game being cancelled due to a pandemic.
+There is no missing data because there is no data to be observed.
+Similarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
+
+Data does not go missing the same way every time.
+There are three broad categorisations describing why data can be missing: Missing Completely at Random (MCAR), Missing at Random (MAR), and Missing Not at Random (MNAR).
+These categorisations help describe whether the missingness occurs by chance (MCAR), due to some observed variable (MAR) or an unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR biased [@Newman2014].
+Football relevant examples are provided in Supplement 1, and more detailed explanations of MCAR, MAR, MNAR can be found elsewhere [@Sainani2015; @Nakagawa2008; @Sterne2009].
+
+The three categorisations of missingness types provide a useful framework to describe bias arising from missing data.
+If you strongly suspect data is MCAR, your results are less likely to be biased.
+If you suspect MAR or MNAR, there is bias in your results.
+This does not mean data MAR or MNAR make results invalid, instead it identifies potential bias, which can be used to help improve future research design.
+However, these categorisations do not necessarily guide the user in their subsequent actions in analysis, or describe specifically how values go missing.
+To identify possible mechanisms for missingness, the data must be explored.
+It can be challenging to identify the missingness mechanism in data, but it is critical, since missing values could change the outcome of a study.
+This could mean the work cannot be reproduced, and worse, may lead scientists and practitioners to the wrong outcome.
 
@@
 It is imperative that missing data, or lack thereof, are reported [@Schafer2002].
-While there has been significant interest in modelling missing
-data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received
-less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
-of articles that report missing data in football research; (2) highlight
-several practices that should be avoided; and (3) provide recommendations for
-exploring and reporting missingness.
+While there has been significant interest in modelling missing data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received less attention [@Schafer2002].
+This article aimed to: (1) estimate the proportion of articles that report missing data in football research; (2) highlight several practices that should be avoided; and (3) provide recommendations for exploring and reporting missingness.
 
@@
 
-To estimate the proportion of articles that report missing data in
-football research, we conducted a systematic search [@Moher2009].
+To estimate the proportion of articles that report missing data in football research, we conducted a systematic search [@Moher2009].
 Articles on football topics (or involving footballers) published in 2019 were studied.
-Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key
-terms and search limits (Supplement 2). Included articles were: (a) written in English; (b) had an accessible
-full-text; and (c) included quantitative data. The PRISMA figure in Supplement 2 summarises the search process,
-and further details of the search can also be found in Supplement 2.
+Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key terms and search limits (Supplement 2).
+Included articles were: (a) written in English; (b) had an accessible full-text; and (c) included quantitative data.
+The PRISMA figure in Supplement 2 summarises the search process, and further details of the search can also be found in Supplement 2.
 
 ```{r agreement}
-
 md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
@@
           accuracy = 0.1)
-
 ```
 
-Our intention was to survey, at random, 10% of the total articles found (after 
-title and abstract screening). We were primarily interested in
-whether missing data was reported or acknowledged; and if not, whether
-a dataset was shared. Where applicable, information relating to how missing
-values were treated was also extracted (see Supplement 2). Two authors
-independently extracted the data (`r md_reported` agreement). Disagreements were resolved by author DNB. Results are reported as the proportion and 95% confidence interval (CI).
+Our intention was to survey, at random, 10% of the total articles found (after title and abstract screening).
+We were primarily interested in whether missing data was reported or acknowledged; and if not, whether a dataset was shared.
+Where applicable, information relating to how missing values were treated was also extracted (see Supplement 2).
+Two authors independently extracted the data (`r md_reported` agreement).
+Disagreements were resolved by author DNB.
+Results are reported as the proportion and 95% confidence interval (CI).
 
@@
 
-Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing
-\~10% of the articles found (see Supplement 2, Figure 1). The proportion of
-articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1).
+The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
 
@@
 
-Our survey of `r n_screened` articles published on football related topics found
-that only one in 13 papers (`r binom_ci_miss_reported$mean_pct`) reported missing data. It is reasonable
-to assume that this result is representative of current practice in the wider sports science literature. 
-The low proportion of papers reporting missingness could be explained by 
-a lack of awareness and education on missing data [@Sainani2020call], 
-similar to other areas of sciences (refs-education and
-ecology examples)(2)**references needed**. Below we discuss several practices to avoid in
-relation to missingness, and provide recommendations for exploring,
-visualising and reporting missing data.
+Our survey of `r n_screened` articles published on football related topics found that only one in 13 papers (`r binom_ci_miss_reported$mean_pct`) reported missing data.
+It is reasonable to assume that this result is representative of current practice in the wider sports science literature.
+The low proportion of papers reporting missingness could be explained by a lack of awareness and education on missing data [@Sainani2020call], similar to other areas of sciences (refs-education and ecology examples)(2)**references needed**.
+Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.
 
@@
 
-Broadly, there are two types of missing values in data: implicit and
-explicit missings. Explicit values are missing, but recorded; whereas
-information in the data. For example, in  Table \@ref(tab:implicit-missings), player Koenen has
-missing values for quarters two and four. Sometimes values like these
-can be logically imputed, as it might be known that these values are
-recorded as 0 in this format, rather than NA. Other places missing data
-can arise include: through joins when merging data without corresponding
-values, surveys, an inability to collect a biological sample (e.g.,
-venous blood), equipment malfunction, failure, or not being worn.
+Broadly, there are two types of missing values in data: implicit and explicit missings.
+Explicit values are missing, but recorded; whereas information in the data.
+For example, in Table \@ref(tab:implicit-missings), player Koenen has missing values for quarters two and four.
+Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA.
+Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.
 
@@
 
-We recommend that researchers use missing data overview graphics (see
-Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
+We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
 
@@
 
-To learn more about exploring missing values, we recommend the vignettes
-in naniar [@naniar], and the methods in (@Tierney2018).
+To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in (@Tierney2018).
 
@@
 
-Imputing data might feel wrong, as we are 'making up data'. The truth
-is, we can (generally) never know what the missing values were. The goal
-of imputing data is to make the best possible inference from the data.
-We recommend data is imputed, with a few caveats. Consider removing
-variables with a high proportion of missing data (e.g., variables with
-the majority or more missing than not). We deliberately do not suggest a 'rule of
-thumb' for the amount of missingness that should be imputed, and caution against
-looking for a specific threshold.
-Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation
-maximisation.
+Imputing data might feel wrong, as we are 'making up data'.
+The truth is, we can (generally) never know what the missing values were.
+The goal of imputing data is to make the best possible inference from the data.
+We recommend data is imputed, with a few caveats.
+Consider removing variables with a high proportion of missing data (e.g., variables with the majority or more missing than not).
+We deliberately do not suggest a 'rule of thumb' for the amount of missingness that should be imputed, and caution against looking for a specific threshold.
+Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation maximisation.
 
-Missing values can occur in the both the predictors (the "independent variables") 
-and response variable (the "dependent variables"). 
+Missing values can occur in the both the predictors (the "independent variables") and response variable (the "dependent variables").
 Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
 
-Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputed the outcome as they provide some sense of what the uncertainty is.
-For discussion on imputing the response value, see ...
-[@refs].
+Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty [@Krausky2018].
+Missing values in the outcome can be dealt with specially for Randomised Control Trials, see - <https://doi.org/10.1191%2F1740774504cn032oa>.
 
-Missing values in the outcome can be dealt with specially for Randomised Control Trials, see - https://doi.org/10.1191%2F1740774504cn032oa.
+<!-- For discussion on imputing the response value, see ... [@refs]. -->
 
-Much in the way of there is no single 'best' statistical method, there
-is no perfect, one-size-fits-all approach for imputing data. The goal is
-to generate similar values that might have been otherwise recorded.
-Sometimes this means using a neighborhood approach of finding similar
-values. Or it could mean predicting responses using linear or tree
-models. Other times the most likely value might have been 0 or the
-last or first value carried forward. For detailed descriptions, and a
-summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].
+Much in the way of there is no single 'best' statistical method, there is no perfect, one-size-fits-all approach for imputing data.
+The goal is to generate similar values that might have been otherwise recorded.
+Sometimes this means using a neighborhood approach of finding similar values.
+Or it could mean predicting responses using linear or tree models.
+Other times the most likely value might have been 0 or the last or first value carried forward.
+For detailed descriptions, and a summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].
 
-The imputation methods discussed so far impute a single value for each
-missing value, and have the eponymous name, "single imputation".
-Multiple imputation is a method where multiple values are imputed for
-each missing value, creating "m" datasets, which are then specially
-combined during analysis. Multiple imputation is generally seen as the
-best method to get the most reasonable inference from the data, as it
-reflects the uncertainty in the missing values. For more information on
-using multiple imputation, we recommend Stefan van Buuren's book, @Van_Buuren2012.
+The imputation methods discussed so far impute a single value for each missing value, and have the eponymous name, "single imputation".
+Multiple imputation is a method where multiple values are imputed for each missing value, creating "m" datasets, which are then specially combined during analysis.
+Multiple imputation is generally seen as the best method to get the most reasonable inference from the data, as it reflects the uncertainty in the missing values.
+For more information on using multiple imputation, we recommend Stefan van Buuren's book, @Van_Buuren2012.
 
-Irrespective of the imputation method used, it is essential to compare results
-of different missing data handling to understand how they may bias the
-results. For example, comparing analysis results from applying listwise
-deletion, compared to mean imputation, compared to linear regression
-imputation. This can reveal bias occurring in imputation methods. An
-example of this approach is described in the Case Study in
-@Tierney2018.
+Irrespective of the imputation method used, it is essential to compare results of different missing data handling to understand how they may bias the results.
+For example, comparing analysis results from applying listwise deletion, compared to mean imputation, compared to linear regression imputation.
+This can reveal bias occurring in imputation methods.
+An example of this approach is described in the Case Study in @Tierney2018.
 
@@
 The current state of reporting missing data in football research has room for improvement.
-In this section we discuss recommended practices for exploring and
-reporting missing data. Our intention is not to provide a 'cookbook'
-style approach to missing data, but rather, broad recommendations to help
-authors when writing the methods and results sections of a study,
-and assist researchers when evaluating a study during the peer-review process.
+In this section we discuss recommended practices for exploring and reporting missing data.
+Our intention is not to provide a 'cookbook' style approach to missing data, but rather, broad recommendations to help authors when writing the methods and results sections of a study, and assist researchers when evaluating a study during the peer-review process.
 
-In the methods section of a study, we recommend the following points are
-addressed:
+In the methods section of a study, we recommend the following points are addressed:
 
@@
 
-> Missing values of ZZ were imputed with a linear regression using terms XX.
-> Missing values were imputed with the mean values
+> Missing values of ZZ were imputed with a linear regression using terms XX. Missing values were imputed with the mean values
 
-An example writeup of missing data for a results section is provided
-below:
+An example writeup of missing data for a results section is provided below:
 
-> *Twenty-five of the 280 values in the dataset (~8.92%) were missing.
+> *Twenty-five of the 280 values in the dataset (\~8.92%) were missing.
 These were due to faults in GPS tracking as participants ran through sections of track covered by forest, a known issue with some GPS tracking.
@@
 
-Our survey of articles on football related topics
-showed that current practice of reporting missing data is poor, 
-with only about one in 20 studies reporting missingness.
-We speculate that this result is representative of practice in the 
-wider sports science literature. The extremely low reporting of missing 
-data could suggest that this aspect of analysis receives little attention. 
-To address this issue and assist researchers,
-we have provided recommendations for reporting and exploring missing data.
-Researchers should consider these recommendations, and pay greater
-attention to missing data and its influence on research results.
+Our survey of articles on football related topics showed that current practice of reporting missing data is poor, with only about one in 20 studies reporting missingness.
+We speculate that this result is representative of practice in the wider sports science literature.
+The extremely low reporting of missing data could suggest that this aspect of analysis receives little attention.
+To address this issue and assist researchers, we have provided recommendations for reporting and exploring missing data.
+Researchers should consider these recommendations, and pay greater attention to missing data and its influence on research results.
 
-# References {-}
-
-# Supplement 1
-
-# Supplement 2
+# References {.unnumbered}

### 2020-11-07 paper/paper.Rmd
edit writing.

@@
     toc: yes
-authors: David N Borg(1), Robert Nguyen(2), Nicholas J Tierney(3,4)
+authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
 csl: sport-medicine.csl
@@
 
-1.  Describe screening procedures **(define or example---think about screened out?)**.
+1.  Describe screening procedures (define or example).
 
@@
 
-2.  State if missing data was explored. If there is missingness, provide a summary graphic (see figure XX)
+2.  State if missing data was explored. If there is missingness, provide a summary graphic (see (Figure \@ref(fig:missing-overview)))
 
@@
 
-> Missing values of ZZ were imputed with a linear regression using terms XX. Missing values were imputed with the mean values
+> Missing values of ZZ were imputed with a linear regression using terms XX. Missing values were imputed with the mean values.
 
@@
 
-Our survey of articles on football related topics showed that current practice of reporting missing data is poor, with only about one in 20 studies reporting missingness.
+Our survey of articles on football related topics showed that current practice of reporting missing data is poor, with only one in nine studies reporting missingness.
 We speculate that this result is representative of practice in the wider sports science literature.

### 2020-11-07 paper/paper.Rmd
edit references

@@
 output:
-  bookdown::pdf_document2:
-    toc: no
   html_document:
@@
     toc_float: yes
+  bookdown::pdf_document2:
+    toc: no
   word_document:
@@
 
-The code to reproduce the results and plots in the paper can be accessed at the following github repository:
+The code to reproduce the results and plots in the paper can be accessed at the following github repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).
 
@@
 There is no universal approach to handling missing data.
-Contextual factors--such as, the study design and objective, and pattern of missingness--determine how missings should be handled, on a case-by-case basis [@Sainani2015].
+Contextual factors---such as, the study design and objective, and pattern of missingness---determine how missings should be handled, on a case-by-case basis [@Sainani2015].
 
@@
 The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-
-Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
-shared their dataset, making
+Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
 it difficult to determine whether there was any missingness. While the
@@
 It is reasonable to assume that this result is representative of current practice in the wider sports science literature.
-The low proportion of papers reporting missingness could be explained by a lack of awareness and education on missing data [@Sainani2020call], similar to other areas of sciences (refs-education and ecology examples)(2)**references needed**.
+The low proportion of papers reporting missingness could be explained by a lack of awareness and education [@Sainani2020call], similar to other science areas [@Peugh2004; @Jelivcic2009].
 Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.
@@
 coefficients (i.e., less/more). When using imputation, authors need to
-perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**).
+perform a sensitivity analysis [@Wood2004].
 
@@
 
-Missing values can occur in the both the predictors (the "independent variables") and response variable (the "dependent variables").
+Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables").
 Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
-
-Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty [@Krausky2018].
-Missing values in the outcome can be dealt with specially for Randomised Control Trials, see - <https://doi.org/10.1191%2F1740774504cn032oa>.
-
-<!-- For discussion on imputing the response value, see ... [@refs]. -->
+Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty (see Chapter 15 in [@McElreath2020]).
+Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].
 

### 2020-11-07 paper/supplement-2.Rmd
try fixing kableExtra
edit references

@@
 output: 
+  bookdown::html_document2:
+    keep_md: true
   bookdown::pdf_document2:
     toc: false
-  bookdown::html_document2:
-    keep_md: true
 bibliography: "supp2.bib"
@@
 library(kableExtra)
-options(kableExtra.latex.load_packages = FALSE)
 ```

### 2020-11-07 paper/supplement-2.Rmd
edit writing and results

@@
 Studies containing original data, including case studies (with time-series data) 
-and n=1 trials, were considered.
-The search results were imported into EndNote (version 9.3.2).
+and n=1 trials, were considered. The search results were imported into EndNote (version 9.3.2).
 Duplicates were removed, and titles and abstracts were screened against
@@
 
+```{r agreement}
+md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
+  percent(scale = 1,
+          accuracy = 0.1)
+```
+
 Articles were surveyed to determine: (a) whether a missing data
@@
 statement shared their data. Two authors independently extracted the
-data. Agreement was X%; discrepancies were resolved by NJT. Findings are
+data. Agreement was `r md_reported`; discrepancies were resolved by DNB. Findings are
 reported as the proportion and 95% confidence interval (CI). Confidence
 intervals were calculated using the Clopper-Pearson method for the
-binomial distribution [@Clopper1934], via the 'binom' package [@binom] in R (version 4.0.2) [@rcore].
+binomial distribution [@Clopper1934], via the 'binom' package [@binom] in R [@rcore].
 
@@
 The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-
-Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`).
-shared their dataset, making it difficult to determine whether there was any missingness.
+Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness.
 While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration--especially in contrast to many of the papers that included (potentially meaningless) tests of normality (e.g., Kruskall-Wallis test).
 
-A breakdown of the extracted information is provided in Table \@ref(tab:extracted-info).
+There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data.
+A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).
 
@@
 
-There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data.
-
 # Summary

### 2020-11-07 paper/paper.Rmd
paper edits

@@
 ---
-title: 'Missing Data: Current Practice in Football Research and Recommendations for Improvement'
+title: 'Missing Data: Current Practice in Football Research and Recommendations for
+  Improvement'
 bibliography: references.bib
+csl: sport-medicine.csl
 output:
@@
     toc_float: yes
+  pdf_document:
+    toc: yes
   bookdown::pdf_document2:
@@
 authors: David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
-csl: sport-medicine.csl
 ---

### 2020-11-07 packages.R
add 'irr' package

@@ -16,6 +16,7 @@ library(binom)
 library(kableExtra)
 library(knitr)
 library(scales)
+library(irr)
 library(googlesheets4)
 library(gt)
 library(PRISMAstatement)

### 2020-11-07 packages.R
add irr

@@ -12,11 +12,11 @@ library(dotenv)
 library(drake)
 library(rmarkdown)
 library(tidyverse)
+library(irr)
 library(binom)
 library(kableExtra)
 library(knitr)
 library(scales)
-library(irr)
 library(googlesheets4)
 library(gt)
 library(PRISMAstatement)

### 2020-11-08 paper/paper.Rmd
edit paper

@@
 library(scales)
+library(tidyr)
+library(irr)
 ```
@@
 
-**Acknowledgement/conflict of interest**
+**Conflict of interest**
 
@@
 A survey of `r n_screened` articles, sampled at random, was conducted to determine whether a missing data statement was included.
-As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers.
+As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers.
 It is suspected that this result is consistent with practice in the wider sports science literature.
@@
 
+```{r agree}
+agree_var <- function(data, var){
+  statement_agree <- pivot_var(data, {{ var }}) %>% 
+    select(reviewer_1, reviewer_2) %>% 
+    irr::agree()
+
+  statement_agree$value
+}
+```
+
+```{r pivot}
+pivot_var <- function(data, var){
+  data %>% 
+    select(title,
+           reviewer,
+           {{ var }}) %>% 
+    pivot_wider(names_from = reviewer,
+                values_from = {{ var }})
+}
+```
+
 ```{r agreement}
@@
 
-Our survey of `r n_screened` articles published on football related topics found that only one in 13 papers (`r binom_ci_miss_reported$mean_pct`) reported missing data.
+Our survey of `r n_screened` articles published on football related topics found that only one in nine papers (`r binom_ci_miss_reported$mean_pct`) reported missing data.
 It is reasonable to assume that this result is representative of current practice in the wider sports science literature.
@@
 
-(Figure \@ref(fig:missing-overview)) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
+Figure \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
 
-```{r missing-overview, fig.cap = "Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
+```{r missing-overview, fig.cap = "Figure 1.
+Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
 The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
@@
 
-To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in (@Tierney2018).
+To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
 
@@
 Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
-Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty (see Chapter 15 in [@McElreath2020]).
+Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty---see Chapter 15 in [@McElreath2020].
 Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].
@@
 Multiple imputation is generally seen as the best method to get the most reasonable inference from the data, as it reflects the uncertainty in the missing values.
-For more information on using multiple imputation, we recommend Stefan van Buuren's book, @Van_Buuren2012.
+For more information on using multiple imputation, we recommend Stefan van Buuren's book [@Van_Buuren2012].
 
@@
 This can reveal bias occurring in imputation methods.
-An example of this approach is described in the Case Study in @Tierney2018.
+An example of this approach is described in the Case Study in [@Tierney2018].
 

### 2020-11-08 paper/paper.Rmd
edit paper format

@@
 output:
+  pdf_document:
+    toc: no
   html_document:
@@
     toc_float: yes
-  pdf_document:
-    toc: yes
   bookdown::pdf_document2:
@@
 
-Authors:
+**Authors**
 
-David N Borg(1,2)
+David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
 
-ORCiD: 0000-0002-0152-571X
-
-Twitter: elborgo9
-
-Robert Nguyen(3)
-
-ORCiD: 0000-0001-9373-3753
-
-Twitter: roberto_boberto
-
-Nicholas J Tierney(4,5)
-
-ORCiD: 0000-0003-1460-8722
-
-Twitter: nj_tierney
-
-Affiliations:
+**Affiliations**
 
 1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.
+2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
+3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
+4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
+5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.
 
-2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
+**Twitter handles**
 
-3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
-
-4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
-
-5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.
+elborgo9, roberto_boberto, nj_tierney
 
@@
 
-The code to reproduce the results and plots in the paper can be accessed at the following github repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).
+The code to reproduce the results and plots in the paper can be accessed at the following GitHub repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).
 
@@
 To identify possible mechanisms for missingness, the data must be explored.
-It can be challenging to identify the missingness mechanism in data, but it is critical, since missing values could change the outcome of a study.
+It can be challenging to identify the missingness mechanism in data, but it is critical, since missing values can change the outcome of a study.
 This could mean the work cannot be reproduced, and worse, may lead scientists and practitioners to the wrong outcome.
-
-Missing data are values that should have been observed, but were not. Since the
-values are unobserved, this can mean the population of interest might not be 
-properly sampled, inducing bias into a study's results. Left undetected or poorly handled, 
-missing data can undermine the validity and accuracy of research results
-[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
-missing many ways. For example, accidentally skipping a survey question,
-equipment failure, or intentionally not recording values. Before
-analysis, missing data must be handled, with values typically deleted or
-imputed. There is no universal approach to handling missing data. 
-Contextual factors---such as, the study design and objective, and
-pattern of missingness---determine how missings should be handled, on a
-case-by-case basis [@Sainani2015].
-
-For clarity, we describe an example of missingness. An Australian Football data collector 
-was taking lunch during the third quarter, so no data was recorded for this time period. This
-would be considered missing data. Compare this to the game being cancelled due to a pandemic. 
-There is no missing data because there is no data to be observed.
-Similarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
-
-Data does not go missing the same way every time. There are three broad
-categorisations describing why data can be missing: Missing Completely
-at Random (MCAR), Missing at Random (MAR), and Missing Not at Random
-(MNAR). These categorisations help describe whether the missingness
-occurs by chance (MCAR), due to some observed variable (MAR) or an
-unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR
-biased [@Newman2014]. Football relevant examples are provided in Supplement 1, and
-more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
-[@Sainani2015; @Nakagawa2008; @Sterne2009].
-
-The three categorisations of missingness types provide a useful
-framework to describe bias arising from missing data. If you strongly
-suspect data is MCAR, your results are less likely to be biased. If you suspect
-MAR or MNAR, there is bias in your results. This does not mean data MAR
-or MNAR make results invalid, instead it identifies potential bias, which can be
-used to help improve future research design. However, these categorisations
-do not necessarily guide the user in their subsequent actions in
-analysis, or describe specifically how values go missing. To identify
-possible mechanisms for missingness, the data must be explored. It can
-be challenging to identify the missingness mechanism in data, but it is
-critical, since missing values could change the outcome of a study. This
-could mean the work cannot be reproduced, and worse, may lead scientists
-and practitioners to the wrong outcome.
 
@@
 To estimate the proportion of articles that report missing data in football research, we conducted a systematic search [@Moher2009].
-Articles on football topics (or involving footballers) published in 2019 were studied.
+Articles on football topics, or involving footballers, published in 2019 were studied.
 Three major sports medicine databases (SPORTDiscus, Embase and Cinahl) were searched, using key terms and search limits (Supplement 2).
@@
 
-```{r agree}
+```{r}
 agree_var <- function(data, var){
@@
 
-```{r pivot}
+```{r}
 pivot_var <- function(data, var){
@@
 
-Our intention was to survey, at random, 10% of the total articles found (after title and abstract screening).
+Our intention was to survey, at random, 10% of the total articles found after title and abstract screening.
 We were primarily interested in whether missing data was reported or acknowledged; and if not, whether a dataset was shared.
@@
 Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (see Supplement 2, Figure 1).
-The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct`(95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
+The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
+Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
 it difficult to determine whether there was any missingness. While the
@@
 
-Practice to avoid: *Unreported missing values.* While not always the case, even if
+*Unreported missing values.* While not always the case, even if
 not reported, missing data can be obvious. For example, a study examined the
@@
 
-Practice to avoid: *Mean imputation.* The default of most statistical software for
+*Mean imputation.* The default of most statistical software for
 handling missing data in modelling is listwise deletion. Listwise
@@
 
-Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
+*Not evaluating the effect, or choice, of imputation.* The 
 method used to impute missing data has the potential to
@@
 Explicit values are missing, but recorded; whereas information in the data.
-For example, in Table \@ref(tab:implicit-missings), player Koenen has missing values for quarters two and four.
+For example, in \@ref(fig:implicit-missings), player Koenen has missing values for quarters two and four.
 Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA.
@@
 We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
+The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
 
-Figure \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
-
-```{r missing-overview, fig.cap = "Figure 1.
-Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
+```{r missing-overview, fig.cap = "**Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
 The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
@@
 We deliberately do not suggest a 'rule of thumb' for the amount of missingness that should be imputed, and caution against looking for a specific threshold.
-Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as: linear regression, k-nearest neighbours, or expectation maximisation.
+Rather, we recommend: values should generally be imputed, avoiding methods imputing the same value (e.g., the mean or median), and suggest using imputation methods such as linear regression, k-nearest neighbours, or expectation maximisation.
 

### 2020-11-08 paper/paper.Rmd
edit paper

@@
 It is reasonable to assume that this result is representative of current practice in the wider sports science literature.
-The low proportion of papers reporting missingness could be explained by a lack of awareness and education [@Sainani2020call], similar to other science areas [@Peugh2004; @Jelivcic2009].
+The low proportion of papers reporting missingness could be explained by a lack of awareness and education [@Sainani2020call], similar to other sciences [@Peugh2004; @Jelivcic2009].
 Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.
@@
 missingness could bias the results. Authors should include a missing data statement,
-irrespective of whether there is data missing, or not.
+irrespective of whether data are missing, or not.
 
@@
 method is documented. Authors sometimes choose to replace missing 
-observations with mean values (e.g., participant or group average) [@Sampson2018; @Cresswell2009]. 
+observations with mean values, such as a participant or group average [@Sampson2018; @Cresswell2009]. 
 While imputing the mean preserves the study sample size (and point estimates), it also reduces

### 2020-11-09 paper/paper.Rmd
edit formatting for tex.

@@
   Improvement'
+
 bibliography: references.bib
@@
                       fig.align = "center",
-                      echo = FALSE)
+                      echo = FALSE,
+                      dpi = 600)
 ```
@@
 
-**Authors**
-
+\begin{center}
 David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
-
+\end{center}
+\vspace{5mm}
 **Affiliations**
 
-1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.
-2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
-3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
-4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
+1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.\hfill \break
+2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.\hfill \break
+3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.\hfill \break
+4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
 5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.
 
-**Twitter handles**
+\vspace{5mm}
+**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement.
 
-elborgo9, roberto_boberto, nj_tierney
+\vspace{5mm}
+**Twitter:** elborgo9, roberto_boberto, nj_tierney
 
-**Conflict of interest**
+\vspace{5mm}
+**Corresponding Author**
+
+Nicholas J Tierney\hfill \break
+Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
+Email: [nicholas.tierney\@monash.edu](mailto:nicholas.tierney@monash.edu){.email}
+
+\vspace{5mm}
+**Conflict of Interest**
 
@@
 
-**Corresponding author**
+\vspace{5mm}
+**Funding**
 
-Nicholas J Tierney Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
-Address: Email: [nicholas.tierney\@monash.edu](mailto:nicholas.tierney@monash.edu){.email}
+No funding was awarded for this project.
 
-**Data availability**
+\vspace{5mm}
+**Data Availability**
 
-The code to reproduce the results and plots in the paper can be accessed at the following GitHub repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).
+The code to reproduce the results and plots in the paper can be accessed at the GitHub repository: [https://github.com/SciBorgo/Missing-data.git](https://github.com/SciBorgo/Missing-data.git).
 
@@
 
-# Methods and materials
+# Methods and Materials
 
@@
 
-## Practices to avoid
+## Practices to Avoid
 
@@
 
-## Common causes of missing values
+## Common Causes of Missing Values
 
@@
 Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.
+
+\vspace{5mm}
 
@@
 
-## Visualising missing values
+## Visualising Missing Values
 
-We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
-The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
-
-```{r missing-overview, fig.cap = "**Overviews of missing values in simulated injury data in Australian Football matches for August, 2020.
-The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
-Researchers might be interested in investigating the relationship between number of sprints and total sprint distance with incidence of injury.
-Panel A shows...
-Panel B shows ...
-Panel C shows ...
-Panel D shows ...", fig.height = 10}
+```{r}
 
@@
            y = sprints)) +
-  geom_miss_point(size = 2) + 
+  geom_miss_point(size = 2, alpha = 0.15) + 
   theme_minimal() +
@@
 
-```{r patch-2, fig.height = 10}
-(aq_vis_miss + aq_fct) / (gg_miss_scatter + aq_miss_var) & 
+We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
+The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
+To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
+
+```{r missing-overview,  fig.cap = "Overviews of missing values in simulated injury data from Australian Football matches played in August, 2020.
+The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
+Panel A shows the location of missing (black) and complete (grey) cases, and the percentage of missing data in each variable and overall.
+Panel B shows the percentage of missings for each team, in each variable.
+Panel C shows the relationship between two variables in the dataset, including missing values (red) which are shifted, so that they can be visualised on the same axes as the regular values.
+Panel D shows the number of missing values in each variable", fig.height = 7.75}
+
+(aq_vis_miss + aq_fct)/(gg_miss_scatter + aq_miss_var) & 
   theme(aspect.ratio = 1,
@@
 
-To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
-
-## Understanding imputation
+## Understanding Imputation
 

### 2020-11-09 paper/paper.Rmd
edit grammar.

@@
 
+We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]).
+The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in a simulated injury dataset.
+To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
+
 ```{r}
@@
 ```
-
-We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017].
-The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in an simulated injury dataset.
-To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
 

### 2020-11-09 paper/paper.Rmd
edit conclusion.

@@
 
-Our survey of articles on football related topics showed that current practice of reporting missing data is poor, with only one in nine studies reporting missingness.
+Our survey of articles in football research showed that only one in nine studies reported missingness.
 We speculate that this result is representative of practice in the wider sports science literature.
-The extremely low reporting of missing data could suggest that this aspect of analysis receives little attention.
+The low reporting of missing data could suggest that this aspect of analysis receives little attention.
 To address this issue and assist researchers, we have provided recommendations for reporting and exploring missing data.

### 2020-11-17 paper/paper.Rmd
add date.
edit writing and formatting.
edits

@@
 David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
+\vspace{5mm}
+
+17 November 2020
+\vspace{5mm}
+
+Preprint (not peer reviewed)
 \end{center}
@@
 \vspace{5mm}
-**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement.
+**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement. *SportR$\chi$iv*. https://doi.org/xxxx
 
@@
 \vspace{5mm}
-**Conflict of Interest**
+**Conflict of Interest and Funding**
 
-The authors declare no conflicts of interest.
-
-\vspace{5mm}
-**Funding**
-
-No funding was awarded for this project.
+The authors declare no conflicts of interest. No funding was awarded for this project.
 
@@
 Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
-91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019_mention_of_missing].
+91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019mentionofmissing].
 This was not noted anywhere in the paper, or in the supplementary
@@
 Other times the most likely value might have been 0 or the last or first value carried forward.
-For detailed descriptions, and a summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].
+For detailed descriptions, and a summary of these methods, we suggest [@VanBuuren2012; @Schafer2002; @Cheema2014].
 
@@
 Multiple imputation is generally seen as the best method to get the most reasonable inference from the data, as it reflects the uncertainty in the missing values.
-For more information on using multiple imputation, we recommend Stefan van Buuren's book [@Van_Buuren2012].
+For more information on using multiple imputation, we recommend Stefan van Buuren's book [@VanBuuren2012].
 

### 2020-11-17 paper/paper.Rmd
edits

@@
 Left undetected or poorly handled, missing data can undermine the validity and accuracy of research results [@Sainani2015; @Nakagawa2008; @Sterne2009].
-Data can go missing many ways.
+Data can go missing in many ways.
 For example, accidentally skipping a survey question, equipment failure, or intentionally not recording values.
@@
 data are not being reported, it does suggest that this aspect of the
-analysis process receives little consideration---especially in contrast
-to many of the papers that included (potentially meaningless) tests of
-normality (e.g., Kruskall-Wallis test).
+analysis process receives little consideration.
 
@@
 Explicit values are missing, but recorded; whereas information in the data.
-For example, in \@ref(fig:implicit-missings), player Koenen has missing values for quarters two and four.
+For example, in Figure 1, player Koenen has missing values for quarters two and four.
 Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA.
@@
 We recommend that researchers use missing data overview graphics (see Graphics section in [@Tierney2018]; for example, overview plots [@Tierney2017]).
-The panels in \@ref(fig:missing-overview) give an overall sense of the extent of missing and complete data, in a simulated injury dataset.
-To learn more about exploring missing values, we recommend the vignettes in naniar [@naniar], and the methods in [@Tierney2018].
+The panels in Figure 2 give an overall sense of the extent of missing and complete data, in a simulated injury dataset.
+To learn more about exploring missing values, we recommend the vignettes in *naniar* [@naniar], and the methods in [@Tierney2018].
 
@@
 Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables").
-Care should be taken when imputing the outcome variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
-Methods that provide multiple imputed values for the outcome, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the outcome as they describe the uncertainty---see Chapter 15 in [@McElreath2020].
+Care should be taken when imputing the response variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
+Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the response variable as they describe the uncertainty---see Chapter 15 in [@McElreath2020].
 Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].

### 2020-11-17 paper/supplement-1.Rmd
update figures and try to fix tables
edit writing and formatting.
edits

@@
 title: Supplement 1
-output: html_document
+output: pdf_document
 bibliography: supp1.bib
@@
 
-# Missing data examples relevant to football
+# Missing Data Examples Relevant to Australian Football
 
@@
 MAR and MNAR biased. Here, we provide an example of data MAR, MCAR and MNAR,
-relevant to football.
+relevant to Australian Football.
 

### 2020-11-17 paper/supplement-2.Rmd
edit paper
edits

@@
 loadd(n_screened)
-print(n_screened)
 ```
@@
   SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
+                  "AND",
     "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
   Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
+             "AND",
               "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
   Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
+             "AND",
               "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")
@@
 
-```{r agreement}
-md_reported <- agree_var(gs_football, any_missing_data_statement) %>% 
-  percent(scale = 1,
-          accuracy = 0.1)
-```
-
 Articles were surveyed to determine: (a) whether a missing data
@@
 statement shared their data. Two authors independently extracted the
-data. Agreement was `r md_reported`; discrepancies were resolved by DNB. Findings are
+data. Discrepancies were resolved by DNB. Findings are
 reported as the proportion and 95% confidence interval (CI). Confidence
 intervals were calculated using the Clopper-Pearson method for the
-binomial distribution [@Clopper1934], via the 'binom' package [@binom] in R [@rcore].
+binomial distribution [@Clopper1934] via the *binom* package [@binom] in R [@rcore].
 
 # Results
-
-The search returned a total of 3,863 articles.
-Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)).
-The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness.
-While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration--especially in contrast to many of the papers that included (potentially meaningless) tests of normality (e.g., Kruskall-Wallis test).
-
-There were `r nrow(gs_football_missings_stated)` studies that provided some statement about missing data.
-A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).
 
@@
 
+The search returned a total of 3,863 articles.
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)).
+The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
+Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness.
+While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration.
+
+There were `r nrow(gs_football_missings_stated)` studies that provided a statement about missing data.
+A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).
+
 # Summary

### 2020-11-17 R/plan.R

@@ -12,9 +12,9 @@ the_plan <-
   binom_ci_miss_reported = calc_binom_ci_miss_reported(gs_football_final_matches),
   n_no_mention_missings = calc_n_no_mention_missings(gs_football_final_matches),
   binom_ci_data_shared = calc_binom_ci_data_shared(gs_football_final_matches),
-  prisma_figure = create_prisma_figure(),
-  write_prisma_figure = PRISMAstatement:::prisma_pdf(x = prisma_figure, 
-                                                     filename = "paper/figures/prisma-figure.pdf"),
+  #prisma_figure = create_prisma_figure(),
+  #write_prisma_figure = PRISMAstatement:::prisma_pdf(x = prisma_figure, 
+  #                                                   filename = "paper/figures/prisma-figure.pdf"),
    paper = target(
      command = {
        rmarkdown::render(knitr_in("paper/paper.Rmd"))

### 2020-11-17 packages.R

@@ -21,5 +21,8 @@ library(googlesheets4)
 library(gt)
 library(PRISMAstatement)
 library(DiagrammeRsvg)
+library(here)
+library(bookdown)
+library(magick)
 conflict_prefer("filter", "dplyr")
 conflict_prefer("select", "dplyr")

### 2020-11-19 paper/paper.Rmd
edits before preprint.

@@
 
-17 November 2020
+20 November 2020
 \vspace{5mm}
@@
 
-1: Griffith University, Menzies Health Institute Queensland, The Hopkins Centre, Brisbane, Australia.\hfill \break
-2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.\hfill \break
-3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.\hfill \break
-4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.\hfill \break
+1: Griffith University, Menzies Health Institute Queensland (MHIQ), The Hopkins Centre, Brisbane, Australia.
+2: Griffith University, School of Allied Health Sciences, Brisbane, Australia.
+3: University of New South Wales, Department of Statistics, School of Mathematics and Statistics, Sydney, Australia.
+4: Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia.
 5: Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS), Melbourne, Australia.
@@
 Missing data are often unavoidable.
-The reason values go missing, along with decisions made of how missing data are handled (deleted or imputed), can have a profound effect on the validity, accuracy and usability of a study's results.
+The reason values go missing, along with decisions made of how missing data are handled (deleted or imputed), can have a profound effect on the validity and accuracy of study results.
 In this article, we aimed to: estimate the proportion of studies in football research that included a missing data statement, highlight several practices to avoid in relation to missing data, and provide recommendations for exploring, visualising and reporting missingness.
@@
 As expected, the proportion of studies in football research that included a missing data statement was low, at only `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`); suggesting that missingness is seldom considered by researchers.
-It is suspected that this result is consistent with practice in the wider sports science literature.
-We recommend researchers describe the number and percentage of missing values, including when there are no missing values.
+We recommend that researchers describe the number and percentage of missing values, including when there are no missing values.
 Exploratory analysis should be conducted to explore missing values, and visualisations describing missingness overall should be provided in the paper, or at least supplementary materials.
 Missing values should almost always be imputed, and imputation methods should be explored to ensure they are appropriately representative.
+Researchers should consider these recommendations, and pay greater attention to missing data and its influence on research results.
 
@@
 There is no missing data because there is no data to be observed.
-Similarly, if a study screens out individuals who do not meet inclusion criteria, this is not a missing value, because there is no intent for these values to be observed or measured.
+Similarly, if a study screens out individuals who do not meet inclusion criteria, these are not missing values, because there is no intent for these values to be observed or measured.
 
@@
 While there has been significant interest in modelling missing data [@Tierney2015; @Barnett2017], the exploration and reporting of missings have received less attention [@Schafer2002].
-This article aimed to: (1) estimate the proportion of articles that report missing data in football research; (2) highlight several practices that should be avoided; and (3) provide recommendations for exploring and reporting missingness.
+The current article aimed to: (1) estimate the proportion of studies that report missing data in football research; (2) highlight several practices that should be avoided; and (3) provide recommendations for exploring and reporting missingness.
 
@@
 Included articles were: (a) written in English; (b) had an accessible full-text; and (c) included quantitative data.
-The PRISMA figure in Supplement 2 summarises the search process, and further details of the search can also be found in Supplement 2.
+The PRISMA figure in Supplement 2 summarises the search process.
 
@@
 We were primarily interested in whether missing data was reported or acknowledged; and if not, whether a dataset was shared.
-Where applicable, information relating to how missing values were treated was also extracted (see Supplement 2).
+Where applicable, information relating to how missing values were treated was also extracted (Supplement 2).
 Two authors independently extracted the data (`r md_reported` agreement).
@@
 The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI: `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
+Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), only `r binom_ci_data_shared$mean_pct` (95% CI: `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making
 it difficult to determine whether there was any missingness. While the
@@
 It is reasonable to assume that this result is representative of current practice in the wider sports science literature.
-The low proportion of papers reporting missingness could be explained by a lack of awareness and education [@Sainani2020call], similar to other sciences [@Peugh2004; @Jelivcic2009].
+The low proportion of papers reporting missingness might be explained by a lack of awareness and education [@Sainani2020call], similar to other sciences [@Peugh2004; @Jelivcic2009].
 Below we discuss several practices to avoid in relation to missingness, and provide recommendations for exploring, visualising and reporting missing data.
@@
 This was not noted anywhere in the paper, or in the supplementary
-materials. It is unknown whether imputation was undertaken. Without
+materials. It is unknown whether imputation was used. Without
 these details, it is difficult for other analysts and researchers to use
@@
 deletion removes entire rows that contain missing observations from the
-analysis. At best, listwise deletion reduces the power in analysis by
+analysis. At best, listwise deletion reduces the power in the analysis, by
 reducing the sample size. At worst, it introduces bias. For example, if
@@
 the variance, can alter the relationship between the variable with
-missing observations and other variables, and can bias (underestimate)
+missing observations and other variables in the dataset, and can bias (underestimate)
 standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
@@
 Broadly, there are two types of missing values in data: implicit and explicit missings.
-Explicit values are missing, but recorded; whereas information in the data.
+Explicit values are missing, but recorded; whereas implicit missings are not recorded in the dataset but implied.
 For example, in Figure 1, player Koenen has missing values for quarters two and four.
-Sometimes values like these can be logically imputed, as it might be known that these values are recorded as 0 in this format, rather than NA.
-Other places missing data can arise include: through joins when merging data without corresponding values, surveys, an inability to collect a biological sample (e.g., venous blood), equipment malfunction, failure, or not being worn.
+Sometimes values like these can be logically imputed, as it might be known that these values should have been recorded as 0, rather than NA.
+Other places missing data can arise include: through joins when merging data without corresponding values, on surveys, an inability to collect a biological sample (e.g., venous blood), or due to equipment malfunction, failure, or not being worn.
 
@@
 
-```{r implicit-missings, fig.cap = "Two tables demonstrating explicit and implicit missing.
-The first table shows the number of goals scored for a player in a given quarter of an AFL match with the first column showing the player name, the second the quarter they played, and the third the goals they scored.
-Note that Player, 'Koenen' has no entries for Quarter 2 and 4.
+```{r implicit-missings, fig.cap = "Two tables demonstrating explicit and implicit missings.
+The first table shows the number of goals scored for a player in a given quarter of an AFL match with the first column showing the player name, the second column the quarter they played, and the third column the goals they scored.
+Note that Player, 'Koenen' has no entries for Quarters 2 and 4.
 The second table shows the same information from the first table pivoted, with each row being a player and the number of goals they scored in each quarter, with each quarter being a column.
@@
 The dataset contains observed player id, team, and match venue data, in addition to simulated data on injury incidence, and number of sprints per match, and total sprint distance per match.
-Panel A shows the location of missing (black) and complete (grey) cases, and the percentage of missing data in each variable and overall.
+Panel A shows the location of missing and present cases, and the percentage of missing data in each variable (and overall).
 Panel B shows the percentage of missings for each team, in each variable.
@@
 Missing values can occur in the both the predictors (the "independent variables") and the response variable (the "dependent variables").
-Care should be taken when imputing the response variables, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
-Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing the response variable as they describe the uncertainty---see Chapter 15 in [@McElreath2020].
+Care should be taken when imputing the response variable, and in general they should not be imputed with a single imputation, as this will bias the subsequent analysis.
+Methods that provide multiple imputed values for a response variable, such as multiple imputation, or using a Bayesian framework that generates a distribution of posterior values, can be appropriate for imputing outcome variables as they describe the uncertainty (see Chapter 15 in [@McElreath2020]).
 Missing values in the outcome can be dealt with specially for Randomised Control Trials [@Jakobsen2017; @Wood2004].
@@
 Or it could mean predicting responses using linear or tree models.
-Other times the most likely value might have been 0 or the last or first value carried forward.
+Other times the most likely value might have been 0 or the last (or first) value carried forward.
 For detailed descriptions, and a summary of these methods, we suggest [@VanBuuren2012; @Schafer2002; @Cheema2014].
@@
 The current state of reporting missing data in football research has room for improvement.
-In this section we discuss recommended practices for exploring and reporting missing data.
+In this section we recommend practices for exploring and reporting missing data.
 Our intention is not to provide a 'cookbook' style approach to missing data, but rather, broad recommendations to help authors when writing the methods and results sections of a study, and assist researchers when evaluating a study during the peer-review process.

### 2020-11-19 paper/supplement-1.Rmd
edits before preprint.

@@
 bibliography: supp1.bib
+csl: sport-medicine.csl
 ---
@@
 
-Data MAR could arise if players' body mass measurements were taken with a
-weighing scale that produced more missing values when placed on soft carpet,
+Data MAR could arise if players' body mass measurements were taken with a scale 
+that produced more missing values when placed on soft carpet,
 rather than hard tiles, in the team change room. If surface type was recorded,
@@
 Data MNAR also arises when missing values are related to the values missing.
-This could take place if, for example, the scales used only measure up to 110 Kg, and then produce missing values.
+This could take place if, for example, the scales used only measure up to 110 kg, and then produce missing values.
 This is a systematic issue---it happens each time, and means that heavier players are not recorded.

### 2020-11-19 paper/supplement-2.Rmd
edits before preprint.

@@
 output:
-  bookdown::html_document2:
+  pdf_document:
+    toc: false
+  html_document:
     keep_md: true
-  bookdown::pdf_document2:
-    toc: false
 bibliography: "supp2.bib"
+csl: sport-medicine.csl
 ---
@@
 
-# Systematic search details
+# Systematic Search Details
 
@@
 described below. The search protocol was not registered, and no funding
-was provided to support the undertaking of the search.
+was provided to support the undertaking of this work.
 
@@
 
-This review aimed to estimate the proportion of articles that reported 
+We aimed to estimate the proportion of articles that reported 
 missing data in football research.
@@
 sports medicine databases were searched on May 20, 2020, using key terms and limits. 
-The databases were: SPORTDiscus, Embase and Cinahl. A
-list of the search terms has been provided in
-Table \@ref(tab:search-table-print) below.
-
-```{r search-table}
-table_terms <- tibble(
-  SPORTDiscus = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
-                  "AND",
-    "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football"),
-  Embase = c("(exercise AND science OR (sports AND science) OR physiology OR biomechanics OR nutrition OR training OR testing OR (sports AND medicine) OR 'performance analysis' OR (('performance'/exp OR performance) AND ('analysis'/exp OR analysis)) OR 'performance'/exp OR performance OR 'sport psychology'/exp OR 'sport psychology' OR (('sport'/exp OR sport) AND ('psychology'/exp OR psychology)) OR 'coaching'/exp OR coaching) AND [2019-2019]/py",
-             "AND",
-              "(football OR soccer OR (indoor AND soccer) OR futsal OR rugby OR (rugby AND league) OR (rugby AND union) OR (gaelic AND football) OR 'australian rules footbal' OR (('australian'/exp OR australian) AND rules AND footbal) OR afl OR 'american football'/exp OR 'american football' OR (('american'/exp OR american) AND ('football'/exp OR football)) OR 'gridiron football' OR (gridiron AND ('football'/exp OR football)) OR 'touch football' OR (('touch'/exp OR touch) AND ('football'/exp OR football))) AND [2019-2019]/py"),
-  Cinahl = c("exercise science OR sports science OR physiology OR biomechanics OR nutrition OR training OR testing OR sports medicine OR performance analysis OR performance OR sport psychology OR coaching",
-             "AND",
-              "football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football")
-
-)
-```
-
-```{r search-table-print}
-table_terms %>% 
-  kbl(caption = "Key search terms and complete search statements for the databases, SPORTDiscuss, Embase, and Cinahl",
-      booktabs = TRUE) %>% 
-  kable_styling("striped",
-                full_width = TRUE)
-```
+The databases were: SPORTDiscus, Embase and Cinahl. The search terms were: exercise science 
+OR sports science OR physiology OR biomechanics OR nutrition OR training 
+OR testing OR sports medicine OR performance analysis OR performance 
+OR sport psychology OR coaching AND football OR soccer OR indoor soccer OR futsal OR rugby league OR rugby union OR gaelic football OR australian rules football OR AFL OR american football OR gridiron football OR touch football.
 
@@
 Studies containing original data, including case studies (with time-series data) 
-and n=1 trials, were considered. The search results were imported into EndNote (version 9.3.2).
-Duplicates were removed, and titles and abstracts were screened against
+and n=1 trials, were considered. The search results were imported into EndNote (version 9.3.2), 
+duplicates were removed, and titles and abstracts were screened against
 the inclusion criteria. Of the articles identified for full-text
 screening, 10% were sampled, at random, for survey. Because \~30% of the
-papers from the initial random sample were not eligible for inclusion
+papers from the initial sample were not eligible for inclusion
 (e.g., full-texts not in English, conference abstracts, qualitative
-studies), an additional 5% random sample was taken. It was expected that 
-30% of the additional 5% random sample would also be ineligible. With this loss,
+studies), an additional 5% random sample was taken. It was anticipated that 
+30% of the additional 5% sample would also be ineligible. With this loss,
 it was expected that 10% of the articles that remained after title and abstract 
-screening would undergo survey. 
-Figure \@ref(fig:prisma) summarises the search process.
+screening would undergo survey. Figure 1 summarises the search process.
 
-```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 8}
+```{r prisma, fig.cap = "Flowchart of the article search and inclusion for the systematic search. The full-text of 137 articles was reviewed.", warning = FALSE, fig.height = 10, fig.width = 6}
 knitr::include_graphics("figures/prisma-figure.pdf")
@@
 
-Articles were surveyed to determine: (a) whether a missing data
+Articles were surveyed to determine: whether a missing data
 statement was included, or if missing data were acknowledged in a
-figure/table; where data were missing, whether (b) data were considered
-MCAR, MAR or MNAR; (c) imputation was used; (d) the method used for imputation; 
-and (e) whether articles that did not include a missing data
-statement shared their data. Two authors independently extracted the
+figure/table; and if not, whether articles shared their data. The articles
+that included a missing data statement were surveyed to determine whether: 
+(a) the type of missingness was reported; (b) missing data were explored; and
+(c) if data were imputed. Two authors independently extracted the
 data. Discrepancies were resolved by DNB. Findings are
@@
 The search returned a total of 3,863 articles.
-Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure \@ref(fig:prisma)).
+Of the 199 articles screened, `r n_screened` met the inclusion criteria, representing \~10% of the articles found (Figure 1).
 The proportion of articles that reported missing data was `r binom_ci_miss_reported$mean_pct` (95% CI; `r binom_ci_miss_reported$lower_pct` to `r binom_ci_miss_reported$upper_pct`).
-Of the articles that did not mention missing data, `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness.
+Of the articles that did not mention missing data (`r n_no_mention_missings`/`r n_screened`), only `r binom_ci_data_shared$mean_pct` (95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`) shared their dataset, making it difficult to determine whether there was any missingness.
 While the absence of a missing data statement does not necessarily mean missing data are not being reported, it does suggest that this aspect of the analysis process receives little consideration.
-
 There were `r nrow(gs_football_missings_stated)` studies that provided a statement about missing data.
-A breakdown of the extracted information from these studies is provided in Table \@ref(tab:extracted-info).
+A breakdown of the information extracted from these studies is provided in Table 1.
 
@@
 A survey of `r n_screened` articles published in 2019 on football related topics
-showed that only one in nine reported missing data, suggesting
+showed that only one in nine reported missing data. This result could suggest
 that missing data are rarely considered by sport science researchers.
